### 2020-03-25 Documentation/article_4.tex
documentation UPDATE

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Arsclassica Article
% LaTeX Template
% Version 1.1 (1/8/17)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% Original author:
% Lorenzo Pantieri (http://www.lorenzopantieri.net) with extensive modifications by:
% Vel (vel@latextemplates.com)
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[
12pt, % Main document font size
a4paper, % Paper type, use 'letterpaper' for US Letter paper
oneside, % One page layout (no page indentation)
%twoside, % Two page layout (page indentation for binding and different headers)
headinclude,footinclude, % Extra spacing for the header and footer
BCOR5mm, % Binding correction
]{scrartcl}
\usepackage[english]{babel}
\usepackage{url}
\input{structure.tex} % Include the structure.tex file which specified the document structure and layout

\hyphenation{Fortran hy-phen-ation} % Specify custom hyphenation points in words with dashes where you would like hyphenation to occur, or alternatively, don't put any dashes in a word to stop hyphenation altogether

%----------------------------------------------------------------------------------------
%	TITLE AND AUTHOR(S)
%----------------------------------------------------------------------------------------

\title{\normalfont{disCOVIDer19}} % The article title

\subtitle{A path-guide inside the COVID-19 pandemia } % Uncomment to display a subtitle

\author{Fabio Caironi, Marzio De Corato, \\
Andrea Ierardi, Federico Matteucci,\\
 Gregorio Saporito} % The article author(s) - author affiliations need to be specified in the AUTHOR AFFILIATIONS block

\date{} % An optional date to appear under the author(s)

%----------------------------------------------------------------------------------------

\begin{document}

%----------------------------------------------------------------------------------------
%	HEADERS
%----------------------------------------------------------------------------------------

\renewcommand{\sectionmark}[1]{\markright{\spacedlowsmallcaps{#1}}} % The header for all pages (oneside) or for even pages (twoside)
%\renewcommand{\subsectionmark}[1]{\markright{\thesubsection~#1}} % Uncomment when using the twoside option - this modifies the header on odd pages
\lehead{\mbox{\llap{\small\thepage\kern1em\color{halfgray} \vline}\color{halfgray}\hspace{0.5em}\rightmark\hfil}} % The header style

\pagestyle{scrheadings} % Enable the headers specified in this block

%----------------------------------------------------------------------------------------
%	TABLE OF CONTENTS & LISTS OF FIGURES AND TABLES
%----------------------------------------------------------------------------------------

\maketitle % Print the title/author/date block

\setcounter{tocdepth}{2} % Set the depth of the table of contents to show sections and subsections only

\tableofcontents % Print the table of contents

\listoffigures % Print the list of figures

\listoftables % Print the list of tables

%----------------------------------------------------------------------------------------
%	ABSTRACT
%----------------------------------------------------------------------------------------

\section*{Abstract} % This section will not appear in the table of contents due to the star (\section*)

%----------------------------------------------------------------------------------------
%	AUTHOR AFFILIATIONS
%----------------------------------------------------------------------------------------

%\let\thefootnote\relax\footnotetext{* \textit{Department of Biology, University of Examples, London, United Kingdom}}

%\let\thefootnote\relax\footnotetext{\textsuperscript{1} \textit{Department of Chemistry, University of Examples, London, United Kingdom}}

%----------------------------------------------------------------------------------------

\newpage % Start the article content on the second page, remove this if you have a longer abstract that goes onto the second page

%----------------------------------------------------------------------------------------
%	INTRODUCTION
%----------------------------------------------------------------------------------------

\section{Introduction}

In December 2019 different cases of pneumonia were reported in Wuhan (China) \cite{huang2020clinical}. Their origin was later ascribed to a new virus classified as \textit{Severe acute respiratory syndrome coronavirus 2} (SARS-CoV-2). The origin of this virus is still subject of scientific debate between the scientific community, however one of the most common opinion is that this virus comes from bats, in particular the genus Rhinolophus \cite{zhou2020pneumonia}. The most compelling feature of this virus is that is its ability to spread also via coughing and sneezing \cite{ghinai2020first}, and also by touching infected surfaces \cite{chang2020protecting}. Differently with respect to the SARS-CoV the virus seems to have a lower mortality rate \cite{sorensen2006severe,weiss2020clinical}. In January the Chinese government imposed the quarantine for the city of Wuhan (almost 11M people); the quarantine was later expanded to the full province of Hubei (60M people) and then to the neighbour provinces Huanggang, Ezhou and Xianning. 
The virus then spread in Canada, Germany, Thailand and Japan and then in other different countries included Italy \cite{timeline+web}

%----------------------------------------------------------------------------------------
%	METHODS
%----------------------------------------------------------------------------------------

\section{Theoretical Background}

%------------------------------------------------

\section{Data Origin}

%------------------------------------------------

\section{Panels}

\subsection{Home}

\subsection{Data Inspection}

\subsection{Data Analysis}

\section{Conclusions}

\section{Packages used}

A statement requiring citation \cite{Figueredo:2009dg}.

%------------------------------------------------

%----------------------------------------------------------------------------------------
%	BIBLIOGRAPHY
%----------------------------------------------------------------------------------------

\renewcommand{\refname}{\spacedlowsmallcaps{References}} % For modifying the bibliography heading

\bibliographystyle{unsrt}

\bibliography{sample.bib} % The file containing the bibliography

%----------------------------------------------------------------------------------------

\end{document}

### 2020-03-25 inst/shiny/C19/server/2module/inspection.R
update selection introduction plot in inspection

@@ -1,10 +1,30 @@
 # General info plot
+dataset = reactive({
+  
+  if(input$regiontab2 == "default" && input$provincetab2 == "default") {
+    name <- input$countrytab2
+    print(name)
+    dataset <- countryTS$Italy
+    print(dataset)
+    
+  } else if(input$regiontab2 != "default") {
+    name <- input$regiontab2
+    print(name)
+    dataset <- regionTS[[name]]
+    
+  } else {
+    print("OK")
+    
+  }
+  #return(dataset)
+}
+
+)
 
 
 output$general_infos_plot <- highcharter::renderHighchart(
   
-  
-  highcharter::hchart(countryTS$Italy,"spline",title= "General info",highcharter::hcaes(x=data,y = totale_casi),  name="Total cases", color="blue", yAxis = 1,showInLegend=TRUE)
+  highcharter::hchart(dataset(),"spline",title= "General info",highcharter::hcaes(x=data,y = totale_casi),  name="Total cases", color="blue", yAxis = 1,showInLegend=TRUE)
   %>% 
     
     highcharter::hc_chart(zoomType = "xy") %>%
@@ -13,25 +33,26 @@ output$general_infos_plot <- highcharter::renderHighchart(
       list(showLastLabel = TRUE, opposite = TRUE, title = list(text  =  ''))
     ) %>%
     
-    highcharter::hc_add_series(data = countryTS$Italy, type = "spline", 
+    highcharter::hc_add_series(data = dataset(), type = "spline", 
                                yAxis = 1, highcharter::hcaes(x = data, y = terapia_intensiva),
                                name="Total Intesive care", color="red",showInLegend=TRUE)
   %>%
-    highcharter::hc_add_series(data = countryTS$Italy, type = "spline", 
+    highcharter::hc_add_series(data =dataset(), type = "spline", 
                                yAxis = 1, highcharter::hcaes(x = data, y = ricoverati_con_sintomi),
                                name="Total Hospitalized", color="orange",showInLegend=TRUE)
   %>%
-    highcharter::hc_add_series(data = countryTS$Italy, type = "spline", 
+    highcharter::hc_add_series(data =dataset(), type = "spline", 
                                yAxis = 1, highcharter::hcaes(x = data, y = deceduti),
                                name="Total Deaths", color="black",showInLegend=TRUE)
   %>%
-    highcharter::hc_add_series(data = countryTS$Italy, type = "spline", 
[…]

### 2020-03-25 inst/shiny/C19/ui/tabs/2.tab_inspection.R

@@ -39,7 +39,28 @@ shinydashboard::tabItem(tabName = "tab_2",
                                                            shinydashboard::tabBox(width = 12,
                                                                                   title = NULL,
                                                                                   tabPanel("Plot",
-                                                                                           highcharter::highchartOutput("general_infos_plot")%>%shinycssloaders::withSpinner( color="#dd4b39")
+                                                                                           
+                                                                                           
+                                                                                           shiny::fluidRow(
+                                                                                             shiny::column(2,
+                                                                                                           shiny::selectizeInput(
+                                                                                                             inputId = "countrytab2", label = "Country",
+                                                                                                             choices = countryNames, selected = "Italy"),
+                                                                                                           
+                                                                                                           shiny::selectizeInput(
+                                                                                                             inputId = "regiontab2", label = "Region",
+                                                                                                             choices = c("--- ALL ---" = "default", regNames), selected = NULL),
+                                                                                                           
+                                                                                                           shiny::selectizeInput(
+                                                                                                             inputId = "provincetab2", label = "Province",
+                                                                                                             choices = c("--- ALL ---" = "default", provNames), selected = NULL)
+                                                                                             ),
+                                                                                             shiny::column(10,
+                                                                                                           highcharter::highchartOutput("general_infos_plot")
+                                                                                                           
+                                                                                             )
+                                                                                           ),
+                                                                                           
                                                                                   ),
                                                                                   
                                                                                   tabPanel("Raw data",

### 2020-03-25 inst/shiny/C19/global.R
bug fix

@@ -169,6 +169,8 @@ dfita2 <- dfita2 %>%
 #====== MODULE 2 - INSPECTION ====== 
 
 
+
+
 # plot growth monitoring --------------------------------------------------------------------
 out_growth <- country_growth
 

### 2020-03-25 inst/shiny/C19/server/2module/inspection.R

@@ -1,56 +1,71 @@
-# General info plot
+
+# Province input updater
+#shiny::observe({
+#  wait <- waitInput()
+#  if(input$regiontab2 != "default") {
+#    ch <- c("--- ALL ---" = "default", regAndProv[regAndProv$regiontab2 == input$regiontab2, "provincetab2"])
+#  } else {
+#    ch <- c("--- ALL ---" = "default", provNames)
+#  }
+#  shiny::updateSelectizeInput(session, inputId = "provincetab2", choices = ch, selected = NULL)
+#})
+
+
+# General info reactive dataset
 dataset = reactive({
+  
   
   if(input$regiontab2 == "default" && input$provincetab2 == "default") {
     name <- input$countrytab2
-    print(name)
     dataset <- countryTS$Italy
-    print(dataset)
-    
-  } else if(input$regiontab2 != "default") {
+
+  }
+   
+
+   else if(input$regiontab2 != "default" &&  input$provincetab2 == "default") {
     name <- input$regiontab2
-    print(name)
     dataset <- regionTS[[name]]
-    
-  } else {
-    print("OK")
+   }
+  
+  
+  #DA MODIFICARE ID INPUT IN TAB2
+   else if(input$regiontab2 != "default" &&  input$provincetab2 != "default"){
+    name <- input$provincetab2
+    dataset <- provTS[[name]]
     
   }
-  #return(dataset)
+  else
+  {
+    need(FALSE, "Wrong inputs")
+  }
 }
 
 )
 
-
+# General info reactive plot
 output$general_infos_plot <- highcharter::renderHighchart(
-  
-  highcharter::hchart(dataset(),"spline",title= "General info",highcharter::hcaes(x=data,y = totale_casi),  name="Total cases", color="blue", yAxis = 1,showInLegend=TRUE)
-  %>% 
-    
+
+  highcharter::hchart(dataset(),"spline",title= "General info",highcharter::hcaes(x=data,y = totale_casi),  name="Total cases", color="blue", yAxis = 1,showInLegend=TRUE) %>% 
     highcharter::hc_chart(zoomType = "xy") %>%
     highcharter::hc_yAxis_multiples(
       list(lineWidth = 3, title = list(text  =  '')),
       list(showLastLabel = TRUE, opposite = TRUE, title = list(text  =  ''))
-    ) %>%
-    
+    )  %>%
     highcharter::hc_add_series(data = dataset(), type = "spline", 
[…]

### 2020-03-25 inst/shiny/C19/ui/tabs/2.tab_inspection.R

@@ -9,20 +9,12 @@ shinydashboard::tabItem(tabName = "tab_2",
                         shinydashboard::valueBox("Introduction", "Total cases, total deaths, total recovered, total hospitalised",
                                                  icon=icon("chart-line"),
                                                  color = "yellow", width = NULL),
-                        fluidRow(
-                                # column(4,
-                                #        
-                                #        box(title="Inputs", solidHeader = T,
-                                #            width = NULL,
-                                #            status = "danger"
-                                #            
-                                #        )),
-                                column(12,
-                                       
-                                       shinydashboard::box(title="General info", solidHeader = T,
-                                                           width = NULL,
-                                                           color = "red",
-                                                           tags$head(tags$style(HTML('
+                        
+                        
+                        shinydashboard::box(title="General info", solidHeader = T,
+                                            width = NULL,
+                                            color = "red",
+                                            tags$head(tags$style(HTML('
                                                      /* tabBox background */
                                                      
                                                      .nav-tabs-custom > .nav-tabs > li.active {
@@ -35,44 +27,44 @@ shinydashboard::tabItem(tabName = "tab_2",
                                                      border-color: #dd4b39 !important;
                                                      }
                                                      '))),
-                                                           
-                                                           shinydashboard::tabBox(width = 12,
-                                                                                  title = NULL,
-                                                                                  tabPanel("Plot",
-                                                                                           
-                                                                                           
-                                                                                           shiny::fluidRow(
-                                                                                             shiny::column(2,
-                                                                                                           shiny::selectizeInput(
-                                                                                                             inputId = "countrytab2", label = "Country",
-                                                                                                             choices = countryNames, selected = "Italy"),
-                                                                                                           
-                                                                                                           shiny::selectizeInput(
-                                                                                                             inputId = "regiontab2", label = "Region",
-                                                                                                             choices = c("--- ALL ---" = "default", regNames), selected = NULL),
-                                                                                                           
-                                                                                                           shiny::selectizeInput(
-                                                                                                             inputId = "provincetab2", label = "Province",
-                                                                                                             choices = c("--- ALL ---" = "default", provNames), selected = NULL)
-                                                                                             ),
-                                                                                             shiny::column(10,
-                                                                                                           highcharter::highchartOutput("general_infos_plot")
-                                                                                                           
-                                                                                             )
-                                                                                           ),
-                                                                                           
-                                                                                  ),
-                                                                                  
-                                                                                  tabPanel("Raw data",
[…]

### 2020-03-25 inst/shiny/C19/server/3module/analysis.R
fix recal of fix_trim in analysis

@@ -28,7 +28,7 @@ suggest_date <- function()
 {
   wait <- waitLoading()
   
-  new = const_trim(eval(t$data)[[t$name]]$totale_casi,1)
+  new = covid19:::const_trim(eval(t$data)[[t$name]]$totale_casi,1)
   index = which(eval(t$data)[[t$name]]$totale_casi %in% new)
   newdates = eval(t$data)[[t$name]]$data[index]
   

### 2020-03-25 inst/shiny/C19/server/2module/inspection.R
Update inspection.R

@@ -18,18 +18,18 @@ dataset = reactive({
   if(input$regiontab2 == "default" && input$provincetab2 == "default") {
     name <- input$countrytab2
     dataset <- countryTS$Italy
-
+    
   }
-   
-
-   else if(input$regiontab2 != "default" &&  input$provincetab2 == "default") {
+  
+  
+  else if(input$regiontab2 != "default" &&  input$provincetab2 == "default") {
     name <- input$regiontab2
     dataset <- regionTS[[name]]
-   }
+  }
   
   
   #DA MODIFICARE ID INPUT IN TAB2
-   else if(input$regiontab2 != "default" &&  input$provincetab2 != "default"){
+  else if(input$regiontab2 != "default" &&  input$provincetab2 != "default"){
     name <- input$provincetab2
     dataset <- provTS[[name]]
     
@@ -44,7 +44,7 @@ dataset = reactive({
 
 # General info reactive plot
 output$general_infos_plot <- highcharter::renderHighchart(
-
+  
   highcharter::hchart(dataset(),"spline",title= "General info",highcharter::hcaes(x=data,y = totale_casi),  name="Total cases", color="blue", yAxis = 1,showInLegend=TRUE) %>% 
     highcharter::hc_chart(zoomType = "xy") %>%
     highcharter::hc_yAxis_multiples(
@@ -65,7 +65,7 @@ output$general_infos_plot <- highcharter::renderHighchart(
                                name="Total recovered", color="green",showInLegend=TRUE)  %>%
     highcharter::hc_legend(align = "top", verticalAlign = "top",
                            layout = "vertical", x = 0, y = 100, enabled=TRUE) 
-    
+  
 )
 
 
@@ -101,142 +101,148 @@ output$rawData_input <- shiny::renderUI({
                     shiny::actionButton(inputId = "rawData_go", "Show")
       )
     )
-
+    
   )
   
 })
 
 output$rawData_sel_input <- shiny::renderUI({
-  if(input$rawData_terr != 1) {
-    if(input$rawData_terr == 2)
-      shiny::selectInput(inputId = "rawData_reg_sel", label = NULL,
-                         choices = regNames, selected = "Lombardia")
-    else if(input$rawData_terr == 3)
-      shiny::selectInput(inputId = "rawData_prov_sel", label = NULL,
-                         choices = provNames, selected = "Milano")
+  
+  if(is_ready(input$rawData_terr))
+     {
+       
[…]

### 2020-03-26 Documentation/article_4.tex
update

@@
 In January the Chinese government imposed the quarantine for the city of Wuhan (almost 11M people); the quarantine was later expanded to the full province of Hubei (60M people) and then to the neighbour provinces Huanggang, Ezhou and Xianning.
-The virus then spread in Canada, Germany, Thailand and Japan and then in other different countries included Italy \cite{timeline+web}
+The virus then spread in Canada, Germany, Thailand and Japan and then in other different countries included Italy \cite{timeline+web}.
+In Italy the first cases, two Chinese tourist from Hubei, were reported in Rome \cite{corr+roma}; then  other cases were reported in Codogno (Lodi,Lombardy).
+Later the virus spread almost in all regions of Italy with a higher density in Lombardy, Veneto, Emilia-Romagna, Piemonte and Marche.
+Starting from 22 of February, the Italian government started to impose the quarantine (red-zone) for 11 different municipalities, in particular Codogno, Casalpusterlengo, Lodi (included the neighbour municipalities) and Vo' Euganeo.
+Starting from this date different restrictive measures were imposted starting from the regions with the highest number of cases: public entrainment were almost suspended as well as schools and universities.
+Workers (in public and private sectors) were allowed, when possibile, to work at home (smart-working).
+Such measures culminated with a decree approved by the Italian government that divided Italy in three areas: the red zones in which the municipalities with highest number of cases in which all the population was subjected to quarantine, the yellow one (Lombardia, Veneto, Emilia Romagna) in which schools, universities and public events ,sports  as well cinemas were suspended and the rest of Italy in which no restrictive measures were adopted \cite{rep+dec1}; however 3 days later all schools and universities were suspended \cite{guard+01}
 

### 2020-03-26 inst/shiny/C19/global.R
add select province for plot inspection, change color of tabbox

@@ -28,7 +28,6 @@ toLog10<- function(num)
 
 #================================
 
-#library(dplyr)
 #================================
 #====== MODULE 1 - HOME ====== 
 

### 2020-03-26 inst/shiny/C19/server/2module/inspection.R

@@ -1,37 +1,94 @@
+shiny::observe({
+ # wait <- waitInput()
+  if(input$regiontab2 != "default") {
+    ch <- c("--- ALL ---" = "default", regAndProv[regAndProv$region == input$regiontab2, "province"])
+  } else {
+    ch <- c("--- ALL ---" = "default", provNames)
+  }
+  shiny::updateSelectizeInput(session, inputId = "provincetab2", choices = ch, selected = NULL)
+})
 
-# Province input updater
-#shiny::observe({
-#  wait <- waitInput()
-#  if(input$regiontab2 != "default") {
-#    ch <- c("--- ALL ---" = "default", regAndProv[regAndProv$regiontab2 == input$regiontab2, "provincetab2"])
-#  } else {
-#    ch <- c("--- ALL ---" = "default", provNames)
-#  }
-#  shiny::updateSelectizeInput(session, inputId = "provincetab2", choices = ch, selected = NULL)
-#})
 
+reac_dataset <- shiny::reactiveValues()
 
 # General info reactive dataset
-dataset = reactive({
+shiny::observe({
   
   
   if(input$regiontab2 == "default" && input$provincetab2 == "default") {
-    name <- input$countrytab2
-    dataset <- countryTS$Italy
-    
+    reac_dataset$name <- input$countrytab2
+    reac_dataset$dataset <- countryTS$Italy
+    reac_dataset$plot = highcharter::hchart(reac_dataset$dataset,"spline",title= "General info",highcharter::hcaes(x=data,y = totale_casi),  name="Total cases", color="blue", yAxis = 1,showInLegend=TRUE) %>% 
+      highcharter::hc_chart(zoomType = "xy") %>%
+      highcharter::hc_yAxis_multiples(
+        list(lineWidth = 3, title = list(text  =  '')),
+        list(showLastLabel = TRUE, opposite = TRUE, title = list(text  =  ''))
+      )  %>%
+      highcharter::hc_add_series(data = reac_dataset$dataset, type = "spline", 
+                                 yAxis = 1, highcharter::hcaes(x = data, y = terapia_intensiva),
+                                 name="Total Intesive care", color="red",showInLegend=TRUE) %>%
+      highcharter::hc_add_series(data =reac_dataset$dataset, type = "spline", 
+                                 yAxis = 1, highcharter::hcaes(x = data, y = ricoverati_con_sintomi),
[…]

### 2020-03-26 inst/shiny/C19/ui/tabs/2.tab_inspection.R

@@ -8,7 +8,7 @@ shinydashboard::tabItem(tabName = "tab_2",
                         
                         shinydashboard::valueBox("Introduction", "Total cases, total deaths, total recovered, total hospitalised",
                                                  icon=icon("chart-line"),
-                                                 color = "yellow", width = NULL),
+                                                 color = "navy", width = NULL),
                         
                         
                         shinydashboard::box(title="General info", solidHeader = T,
@@ -30,24 +30,27 @@ shinydashboard::tabItem(tabName = "tab_2",
                                             
                                             shinydashboard::tabBox(width = 12,
                                                                    title = NULL,
-                                                                   tabPanel("Plot",
+                                                                   tabPanel(h2("Plot"),
                                                                             
                                                                             
                                                                             shiny::fluidRow(
                                                                               shiny::column(2,
-                                                                                            shiny::selectizeInput(
+                                                                                            shiny::selectInput(
                                                                                               inputId = "countrytab2", label = "Country",
                                                                                               choices = countryNames, selected = "Italy"),
                                                                                             
-                                                                                            shiny::selectizeInput(
+                                                                                            shiny::selectInput(
                                                                                               inputId = "regiontab2", label = "Region",
                                                                                               choices = c("--- ALL ---" = "default", regNames), selected = NULL),
                                                                                             
-                                                                                               shiny::selectizeInput(
+                                                                                               shiny::selectInput(
                                                                                                inputId = "provincetab2", label = "Province",
                                                                                                choices = c("--- ALL ---" = "default"
-                                                                                                           #, provNames
-                                                                                                           ), selected = NULL)
+                                                                                                           , provNames
+                                                                                                           ), selected = NULL),
+                                                                                            
+                                                                                            hr(),
+                                                                                            p("For province, only total case data are available")
                                                                               ),
                                                                               shiny::column(10,
                                                                                             highcharter::highchartOutput("general_infos_plot")
@@ -57,7 +60,7 @@ shinydashboard::tabItem(tabName = "tab_2",
                                                                             
                                                                    ),
                                                                    
-                                                                   tabPanel("Raw data",
+                                                                   tabPanel(h2("Raw data"),
                                                                             shiny::uiOutput("rawData_input"),
                                                                             DT::dataTableOutput("rawData_table")
                                                                    )
[…]

### 2020-03-26 inst/shiny/C19/server/2module/inspection.R
Update inspection.R

@@ -1,5 +1,9 @@
+# ====== GENERAL INFO ==== 
+
+#Dataset and plot reactive
+reac_dataset <- shiny::reactiveValues()
+
 shiny::observe({
- # wait <- waitInput()
   if(input$regiontab2 != "default") {
     ch <- c("--- ALL ---" = "default", regAndProv[regAndProv$region == input$regiontab2, "province"])
   } else {
@@ -9,44 +13,24 @@ shiny::observe({
 })
 
 
-reac_dataset <- shiny::reactiveValues()
 
 # General info reactive dataset
 shiny::observe({
   
-  
   if(input$regiontab2 == "default" && input$provincetab2 == "default") {
     reac_dataset$name <- input$countrytab2
     reac_dataset$dataset <- countryTS$Italy
-    reac_dataset$plot = highcharter::hchart(reac_dataset$dataset,"spline",title= "General info",highcharter::hcaes(x=data,y = totale_casi),  name="Total cases", color="blue", yAxis = 1,showInLegend=TRUE) %>% 
-      highcharter::hc_chart(zoomType = "xy") %>%
-      highcharter::hc_yAxis_multiples(
-        list(lineWidth = 3, title = list(text  =  '')),
-        list(showLastLabel = TRUE, opposite = TRUE, title = list(text  =  ''))
-      )  %>%
-      highcharter::hc_add_series(data = reac_dataset$dataset, type = "spline", 
-                                 yAxis = 1, highcharter::hcaes(x = data, y = terapia_intensiva),
-                                 name="Total Intesive care", color="red",showInLegend=TRUE) %>%
-      highcharter::hc_add_series(data =reac_dataset$dataset, type = "spline", 
-                                 yAxis = 1, highcharter::hcaes(x = data, y = ricoverati_con_sintomi),
-                                 name="Total Hospitalized", color="orange",showInLegend=TRUE)   %>%
-      highcharter::hc_add_series(data =reac_dataset$dataset, type = "spline", 
-                                 yAxis = 1, highcharter::hcaes(x = data, y = deceduti),
-                                 name="Total Deaths", color="black",showInLegend=TRUE)  %>%
-      highcharter::hc_add_series(data =reac_dataset$dataset, type = "spline", 
-                                 yAxis = 1, highcharter::hcaes(x = data, y = dimessi_guariti),
-                                 name="Total recovered", color="green",showInLegend=TRUE)  %>%
[…]

### 2020-03-26 Documentation/article_4.tex
update

@@
 \usepackage{url}
+\usepackage{graphicx}
+\usepackage{subcaption}
 \input{structure.tex} % Include the structure.tex file which specified the document structure and layout
@@
 
-\section{Introduction}
+\section{Introduction} \label{introduction}
 
+\paragraph{Background} \label{Background}
 In December 2019 different cases of pneumonia were reported in Wuhan (China) \cite{huang2020clinical}.
-Their origin was later ascribed to a new virus classified as \textit{Severe acute respiratory syndrome coronavirus 2} (SARS-CoV-2).
-The origin of this virus is still subject of scientific debate between the scientific community, however one of the most common opinion is that this virus comes from bats, in particular the genus Rhinolophus \cite{zhou2020pneumonia}.
+Their origin was later ascribed to a new virus classified as \textit{Severe acute respiratory syndrome coronavirus 2} (SARS-CoV-2) whose TEM and SEM pictures are reported in the Fig.
+\ref{coronavirus_picture} The origin of this virus is still subject of scientific debate between the scientific community, however one of the most common opinion is that this virus comes from bats, in particular the genus Rhinolophus \cite{zhou2020pneumonia}.
 The most compelling feature of this virus is that is its ability to spread also via coughing and sneezing \cite{ghinai2020first}, and also by touching infected surfaces \cite{chang2020protecting}.
 Differently with respect to the SARS-CoV the virus seems to have a lower mortality rate \cite{sorensen2006severe,weiss2020clinical}.
+
+\begin{figure}[ht]
+\begin{subfigure}{.5\textwidth}
+  \centering
+  % include first image
+  \includegraphics[width=.7\linewidth]{Figures/Coronavirus1.jpg}
+  \caption{Transmission electron microscope (TEM) image of SARS-CoV-2—also known as 2019-nCoV, emerging from the surface of cells cultured}
+  \label{fig:sub-first}
+\end{subfigure}
+\begin{subfigure}{.5\textwidth}
+  \centering
+  % include second image
+  \includegraphics[width=.7\linewidth]{Figures/Coronavirus2.jpg} 
+  \caption{Scanning electron microscope of the SARS-CoV-2 emerging from the surface of cells cultured}
+  \label{fig:sub-second}
+\end{subfigure}
+\begin{subfigure}{.5\textwidth}
+  \centering
+  % include second image
+  \includegraphics[width=.7\linewidth]{Figures/Coronavirus3.jpg} 
+  \caption{TEM image of SARS-CoV-2. Not the spikes that gives the name coronavirus to the virus }
+  \label{fig:sub-second}
+\end{subfigure}
+\begin{subfigure}{.5\textwidth}
+  \centering
+  % include second image
+  \includegraphics[width=.7\linewidth]{Figures/Coronavirus4.jpg} 
+  \caption{SEM image of the virus}
+  \label{fig:sub-second}
+\end{subfigure}
+\caption{Different pictures of the virus as reported by the NIAID’s Rocky Mountain Laboratories (RML) in Hamilton, Montana \cite{coronavirus+pictures}  }
+\label{coronavirus_picture}
+\end{figure}
+
 In January the Chinese government imposed the quarantine for the city of Wuhan (almost 11M people); the quarantine was later expanded to the full province of Hubei (60M people) and then to the neighbour provinces Huanggang, Ezhou and Xianning.
@@
 Later the virus spread almost in all regions of Italy with a higher density in Lombardy, Veneto, Emilia-Romagna, Piemonte and Marche.
-Starting from 22 of February, the Italian government started to impose the quarantine (red-zone) for 11 different municipalities, in particular Codogno, Casalpusterlengo, Lodi (included the neighbour municipalities) and Vo' Euganeo.
+Starting from 22 of February, the Italian government started to impose the quarantine (red-zone) for 11 different municipalities, in particular Codogno, Casalpusterlengo, Lodi (included the neighbour municipalities) and Vo' .
 Starting from this date different restrictive measures were imposted starting from the regions with the highest number of cases: public entrainment were almost suspended as well as schools and universities.
 Workers (in public and private sectors) were allowed, when possibile, to work at home (smart-working).
-Such measures culminated with a decree approved by the Italian government that divided Italy in three areas: the red zones in which the municipalities with highest number of cases in which all the population was subjected to quarantine, the yellow one (Lombardia, Veneto, Emilia Romagna) in which schools, universities and public events ,sports  as well cinemas were suspended and the rest of Italy in which no restrictive measures were adopted \cite{rep+dec1}; however 3 days later all schools and universities were suspended \cite{guard+01}
+Such measures culminated with a decree approved by the Italian government that divided Italy in three areas: the red zones in which the municipalities with highest number of cases in which all the population was subjected to quarantine, the yellow one (Lombardia, Veneto, Emilia Romagna) in which schools, universities and public events ,sports  as well cinemas were suspended and the rest of Italy in which no restrictive measures were adopted \cite{rep+dec1}; however 3 days later all schools and universities were suspended \cite{guard+01}.
+The restrictive measures were also extend up to all Italy with a further decree \cite{11marzo}.
+
+\paragraph{Motivation} \label{Motivation}
+In order to have a better understanding about the spread and the effects on the population of the COVID19 disease, we propose a ShinyApp, called disCOVIDer19 , that gathers  the data , mostly provided by the Protezione Civile \cite{protezionecivile+git}, about the number of infections(current as well cumulative), hospitalized people as well in intensive care, deaths and recovered for Italy, its regions and provinces; more details will be discussed in sections \ref{Data Origin} and \ref{Panels}.
+Furthermore we fitted the cumulative cases for Italy and all its regions and provinces with a logistic curve, in order to get a bird-of-eye view on the trend followed as well on the effectiveness of the restrictive measures imposed by the Italian government.
+The theoretical foundations of this fit ,as well how to manage and use it, will be discussed in the section \ref{Theoretical Background}.
+Beside the logistic curve, we also considered a statistical approach, largely diffused among economists, that considers the number of cases in each day as a time series: on this basis we were able to make use of a particular tool, which will be briefly introduced in the section \ref{Theoretical Background}, that allowed us to  make a further forecast  about future cases.
+This latter becomes useful in case of large deviations from logistic distribution.
 
@@
 
-\section{Theoretical Background}
+\section{Theoretical Background} \label{Theoretical Background}
 
@@
 
-\section{Data Origin}
+\section{Data Origin} \label{Data Origin}
 
@@
 
-\section{Panels}
+\section{Panels} \label{Panels}
 
@@
 \section{Packages used}
-
-A statement requiring citation \cite{Figueredo:2009dg}.
 

### 2020-03-26 R/get_intesivecare_cap.R
Added dates to inspection - intensive care capacity/occupancy

@@ -18,23 +18,23 @@ get_intensivecare_cap <- function(regionTS) {
   colnames(readfile) = c("region","capacity")
   readfile = readfile[order(readfile$region),]
   
-  occupation =c()
- 
   sortedreg = regionTS[order(names(regionTS))]
-  for(index_region in 1:length(sortedreg)) 
-  {
-    
-    vector = as.numeric(as.character(sortedreg[[index_region]]$terapia_intensiva)) 
-    occupation[index_region] = tail(vector, n = 1)
-    
-  }
 
-  df = data.frame(readfile,occupation)
-  colnames(df)= c("region","capacity","occupation")
-  perc = c(as.integer(df$occupation/df$capacity*100))
-  df = data.frame(df,perc)
-  df = df[order(df$perc,decreasing = TRUE),]
-  rownames(df) <- 1:nrow(df)
+  newdf = data.frame()
+  for(i in 1:length(readfile$region))
+  {
+    date = sortedreg[[readfile$region[i]]]$data
+    intensive = sortedreg[[readfile$region[i]]]$terapia_intensiva
+    perc = intensive /readfile$capacity[i] * 100
+    perc = round(perc ,digits = 2)
+    v = data.frame(date, intensive, readfile$capacity[i],perc,readfile$region[i])
+    
+    names(v) = c("data","occupancy","capacity","perc","region")
+    newdf <- rbind(newdf, v)
+  }
+  colnames(newdf) = c("data","occupancy","capacity","perc","region")
   
-  return(df)
+
+  
+  return(newdf)
 }

### 2020-03-26 inst/shiny/C19/server/2module/inspection.R

@@ -66,8 +66,8 @@ shiny::observe({
       )  %>%
       highcharter::hc_legend(align = "top", verticalAlign = "top",
                              layout = "vertical", x = 0, y = 100, enabled=TRUE) %>%  highcharter::hc_title(text = paste0("General info for: ",reac_dataset$name),
-                            margin = 20, align = "left",
-                            style = list(useHTML = TRUE))
+                                                                                                           margin = 20, align = "left",
+                                                                                                           style = list(useHTML = TRUE))
     
   }
   else
@@ -127,140 +127,162 @@ output$rawData_input <- shiny::renderUI({
 output$rawData_sel_input <- shiny::renderUI({
   
   if(is_ready(input$rawData_terr))
-     {
-       
-       if(input$rawData_terr != 1) {
-         if(input$rawData_terr == 2)
-           shiny::selectInput(inputId = "rawData_reg_sel", label = NULL,
-                              choices = regNames, selected = "Lombardia")
-         else if(input$rawData_terr == 3)
-           shiny::selectInput(inputId = "rawData_prov_sel", label = NULL,
-                              choices = provNames, selected = "Milano")
-       }
+  {
+    
+    if(input$rawData_terr != 1) {
+      if(input$rawData_terr == 2)
+        shiny::selectInput(inputId = "rawData_reg_sel", label = NULL,
+                           choices = regNames, selected = "Lombardia")
+      else if(input$rawData_terr == 3)
+        shiny::selectInput(inputId = "rawData_prov_sel", label = NULL,
+                           choices = provNames, selected = "Milano")
+    }
   }
 })
-  
-  
-  # shiny::observeEvent(input$rawData_go, {
-  #   output$rawData_table <- shiny::renderTable({
-  #     
-  #     switch(input$rawData_terr,
-  #                "1" = countryTS$Italy %>% 
-  #                  dplyr::select(-stato, -data_seriale) %>%
-  #                  dplyr::filter(data >= input$rawData_date[1] &  data <= input$rawData_date[2]),
-  #                "2" = regionTS[[input$rawData_reg_sel]] %>%
-  #                  dplyr::select(-stato,-lat,-long,-denominazione_regione,-codice_regione,-data_seriale) %>%
[…]

### 2020-03-26 inst/shiny/C19/ui/tabs/2.tab_inspection.R

@@ -53,7 +53,7 @@ shinydashboard::tabItem(tabName = "tab_2",
                                                                                             p("For province, only total case data are available")
                                                                               ),
                                                                               shiny::column(10,
-                                                                                            highcharter::highchartOutput("general_infos_plot")
+                                                                                            highcharter::highchartOutput("general_infos_plot") %>% shinycssloaders::withSpinner( color="#dd4b39")
                                                                                             
                                                                               )
                                                                             )
@@ -76,13 +76,23 @@ shinydashboard::tabItem(tabName = "tab_2",
                                                  color = "navy", width = NULL),
                         
                         fluidRow(
-                          column(6,
+                          shinydashboard::box(title="Intensive care information", solidHeader = T,
+                                              width = NULL,
+                                              color = "red",
+                                              status = "danger",
+                          column(2,
+                               
+                                                     selectInput("occupancy_date",
+                                                                 label = "Select date",
+                                                                 choices = seq(init_date,by=1,fin_date),
+                                                                 selected = fin_date),
+                                                     
+                                                     
                                  
                                  
                                  
-                                 shinydashboard::box(title="Intensive care", solidHeader = T,
-                                                     width = NULL,
-                                                     color = "red",
+                                 ),
+                          column(10,
                                                      
                                                      tags$head(tags$style(HTML('
                          /* tabBox background */
@@ -101,16 +111,19 @@ shinydashboard::tabItem(tabName = "tab_2",
                                                      
                                                      shinydashboard::tabBox(width = 12,
                                                                             title = NULL,
+                                                                            selected = "tab1",
                                                                             # The id lets us use input$tabset1 on the server to find the current tab
-                                                                            id = "tabset2",
-                                                                            tabPanel("% occupation/capacity",
+                                                                            id = "tabs",
+                                                                            tabPanel(value = "tab1",
+                                                                                     title = "% occupation/capacity",
                                                                                      plotly::plotlyOutput("intensivecare_cap_perc",
                                                                                                           width = "100%",
                                                                                                           height = "415px"
                                                                                      ) %>% 
                                                                                        shinycssloaders::withSpinner( color="#dd4b39")
                                                                             ),
-                                                                            tabPanel("capacity vs. occupation ",
+                                                                            tabPanel(value = "tab2",
+                                                                                     title = "capacity vs. occupation ",
[…]

### 2020-03-26 inst/shiny/C19/ui/tabs/2.tab_inspection.R
update spinners

@@ -190,7 +190,7 @@ shinydashboard::tabItem(tabName = "tab_2",
                                                      
                                                      
                                                      highcharter::highchartOutput("plot_test", width = "100%",
-                                                                                  height = "400px"),
+                                                                                  height = "400px")%>% shinycssloaders::withSpinner( color="#dd4b39"),
                                                      height=NULL, status="danger")
                                  
                                  
@@ -201,7 +201,7 @@ shinydashboard::tabItem(tabName = "tab_2",
                                  shinydashboard::box(title="Tests Tracking", status="danger", 
                                                      solidHeader = TRUE, 
                                                      width=12,
-                                                     highcharter::highchartOutput("tamp_plot")
+                                                     highcharter::highchartOutput("tamp_plot")%>%shinycssloaders::withSpinner( color="#dd4b39")
                                  )
                           )
                         )

### 2020-03-26 inst/shiny/C19/ui/tabs/1.tab_home.R
Update 1.tab_home.R

@@ -99,7 +99,8 @@ shinydashboard::tabItem(tabName = "tab_1",
                    HTML("We are a team of data science students from Universita' degli Studi di Milano. 
                             This project is aimed at tracking and modelling the Covid-19 spread in Italy.
                              The data is synchronised with the civil protection database. Each time the app is run it checks for
-                            updates in the civil protection database <br><br><hr>"),
+                            updates in the civil protection database <br><br><br><hr>"),
+                   br(),
                    back_content = tagList(
                      
                      
@@ -159,12 +160,7 @@ shinydashboard::tabItem(tabName = "tab_1",
                                   <img src="https://i.stack.imgur.com/tskMh.png" alt="github"> Github </a> &nbsp; 
                                                       '), 
                                   productPrice = NULL
-                                ),
-                               shinydashboardPlus::productListItem(
-                                  src = "https://media-exp1.licdn.com/dms/image/C4D03AQFqmoSKGgjkVQ/profile-displayphoto-shrink_800_800/0?e=1590019200&v=beta&t=itYf9YpkVI0j3N7CDn4wv0KSoZJJO06NnvrHBzf0c8Q", 
-                                  productTitle = HTML('<b><a href="https://www.linkedin.com/in/tommaso-pessina-754960181/">Tommaso Pessina</a></b>'), 
-                                  productPrice = NULL
-                                )
+                               )
                               )
                        )
                      )

### 2020-03-26 R/get_intesivecare_cap.R
update dataset intensive care

@@ -14,7 +14,8 @@ get_intensivecare_cap <- function(regionTS) {
   
  # readfile = intensivecare_beds
   #readfile <- read.csv("https://raw.githubusercontent.com/marzione00/COVID_19_HACK/master/inst/extdata/Posti_TI.csv",header = TRUE)
-  readfile  <- read.csv(system.file("extdata", "Posti_TI.csv", package="covid19"),header=TRUE)
+  #readfile  <- read.csv(system.file("extdata", "Posti_TI.csv", package="covid19"),header=TRUE)
+  readfile = intensivecare_capacity 
   colnames(readfile) = c("region","capacity")
   readfile = readfile[order(readfile$region),]
   

### 2020-03-26 R/intensivecare_cap_data.R

@@ -1,0 +1,7 @@
+#' Dataset intensive care capacity
+#' 
+#' Dataset of intensive care capacity per Region in Italy
+#' 
+#' @format tibble
+#'
+"intensivecare_cap"

### 2020-03-26 R/get_intesivecare_cap.R
fixed intensive care Rda

@@ -12,9 +12,6 @@
 #' @export
 get_intensivecare_cap <- function(regionTS) {
   
- # readfile = intensivecare_beds
-  #readfile <- read.csv("https://raw.githubusercontent.com/marzione00/COVID_19_HACK/master/inst/extdata/Posti_TI.csv",header = TRUE)
-  #readfile  <- read.csv(system.file("extdata", "Posti_TI.csv", package="covid19"),header=TRUE)
   readfile = intensivecare_capacity 
   colnames(readfile) = c("region","capacity")
   readfile = readfile[order(readfile$region),]

### 2020-03-26 R/get_intesivecare_cap.R
Update get_intesivecare_cap.R

@@ -12,7 +12,7 @@
 #' @export
 get_intensivecare_cap <- function(regionTS) {
   
-  readfile = intensivecare_capacity 
+  readfile = intensivecare_cap 
   colnames(readfile) = c("region","capacity")
   readfile = readfile[order(readfile$region),]
   

### 2020-03-26 R/get_intesivecare_cap.R
Update get_intesivecare_cap.R

@@ -12,8 +12,11 @@
 #' @export
 get_intensivecare_cap <- function(regionTS) {
   
-  readfile = intensivecare_cap 
+  readfile = as.data.frame(intensivecare_cap) 
   colnames(readfile) = c("region","capacity")
+  readfile$capacity = as.numeric(readfile$capacity)
+  readfile$region = as.numeric(readfile$region)
+  
   readfile = readfile[order(readfile$region),]
   
   sortedreg = regionTS[order(names(regionTS))]

### 2020-03-26 R/get_intesivecare_cap.R
fix dataset

@@ -12,11 +12,10 @@
 #' @export
 get_intensivecare_cap <- function(regionTS) {
   
-  readfile = as.data.frame(intensivecare_cap) 
+  readfile = as.data.frame(intensivecare_cap)
   colnames(readfile) = c("region","capacity")
-  readfile$capacity = as.numeric(readfile$capacity)
-  readfile$region = as.numeric(readfile$region)
-  
+
+#  readfile$region = split(as.character(readfile$region)," ")
   readfile = readfile[order(readfile$region),]
   
   sortedreg = regionTS[order(names(regionTS))]
@@ -24,7 +23,8 @@ get_intensivecare_cap <- function(regionTS) {
   newdf = data.frame()
   for(i in 1:length(readfile$region))
   {
-    date = sortedreg[[readfile$region[i]]]$data
+    reg = readfile$region[i]
+    date = sortedreg[[reg]]$data
     intensive = sortedreg[[readfile$region[i]]]$terapia_intensiva
     perc = intensive /readfile$capacity[i] * 100
     perc = round(perc ,digits = 2)
@@ -35,7 +35,6 @@ get_intensivecare_cap <- function(regionTS) {
   }
   colnames(newdf) = c("data","occupancy","capacity","perc","region")
   
-
   
   return(newdf)
 }

### 2020-03-26 inst/shiny/C19/server/3module/analysis.R
uploaded analysis.R

@@ -1,3 +1,11 @@
+## REACTIVES OF THIS CODE
+reac_ARIMA <- shiny::reactiveValues(arimaOK = FALSE)
+    
+    # THIS IS THE REACTIVE CONTAINER OF TERRITORY SELECTION VARIABLES.
+t <- shiny::reactiveValues()
+    
+    ## Reactive values for global storage 
+reac_general <- shiny::reactiveValues()
 
 ## CHECKS FOR ERROR PREVENTING ##
 is_ready <- function(x) {
@@ -7,7 +15,7 @@ is_ready <- function(x) {
   return(TRUE)
 }
 
-# Use this function waitLoading() to validate the content of input$region and prevent errors
+# Use this function waitLoading() to validate the content of t$r and prevent errors
 # (just use it at the beginning of each chunk of code)
 waitLoading <- shiny::reactive({
   shiny::validate(
@@ -23,11 +31,9 @@ waitInput <- shiny::reactive({
   return(input$region)
 })
 
-
-suggest_date <- function()
+# DEPENDS ON REACTIVE t
+suggest_dates <- function()
 {
-  wait <- waitLoading()
-  
   new = covid19:::const_trim(eval(t$data)[[t$name]]$totale_casi,1)
   index = which(eval(t$data)[[t$name]]$totale_casi %in% new)
   newdates = eval(t$data)[[t$name]]$data[index]
@@ -35,15 +41,17 @@ suggest_date <- function()
   return(c(newdates[1], newdates[length(newdates)]))
 }
 
-suggest_lags<- function()
+# DEPENDS ON REACTIVES t AND reac_ARIMA! (but not on reac_ARIMA$arima)
+suggest_lags <- function()
 {
-  wait <- waitLoading()
-  if(is_ready(reac_ARIMA$sample_cases_trim)) {
-    sugg = toString(forecast::auto.arima(log(reac_ARIMA$sample_cases_trim)))
+  checkExp({
+    sugg <- toString(forecast::auto.arima(log(reac_ARIMA$sample_cases_trim)))
     matches <- regmatches(sugg, gregexpr("[[:digit:]]+", sugg))
     return(as.numeric(unlist(matches)))
-  }
+  }, "There is no suitable ARIMA model")
 }
+
+
 
 output$terrInput <- shiny::renderUI({
   shiny::fluidPage(
@@ -56,17 +64,17 @@ output$terrInput <- shiny::renderUI({
       shiny::column(4,
                     shiny::selectizeInput(
                       inputId = "region", label = "Region",
[…]

### 2020-03-26 inst/shiny/C19/ui/tabs/3.tab_analysis.R
uploaded tab_analysis.R

@@ -28,7 +28,7 @@ shinydashboard::tabItem(
     "Logistic model",
     "Curve fitting on confirmed cases of infection's time series",
     icon = icon("analytics"),
-    color = "yellow",
+    color = "navy",
     width = NULL
   ),
   
@@ -101,7 +101,7 @@ shinydashboard::tabItem(
     "ARIMA model",
     "ARIMA(p,i,q) on confirmed cases of infection's time series",
     icon = icon("analytics"),
-    color = "yellow",
+    color = "navy",
     width = NULL
   ),
   
@@ -114,7 +114,8 @@ shinydashboard::tabItem(
         status = "danger",
         solidHeader = TRUE,
         title = "Input",
-        shiny::uiOutput("arimaInput")
+        shiny::uiOutput("arimaDatesInput"),
+        shiny::uiOutput("arimaLagsInput")
       )
     ),
     column(

### 2020-03-27 Documentation/article_4.tex
update

@@
 \section{Theoretical Background} \label{Theoretical Background}
+Different models were proposed in the literature for modelling the spread of a disease \cite{keeling2011modeling}, here we are going to consider a relative simple model taken from the growth dynamics of populations.
+
+\paragraph{Population growth dynamics}
+The dynamics of population was founded by Thomas Malthus in 1817 \cite{malthus1817essay} with his well known equation:
+
+\begin{equation}
+x_{n+1}=x_{n}(1+r)
+\end{equation}
+
+Where $x_{n}$ stands for the population at time and $r$ the growth rate. Thus the following discrete progression will be obtained:
+
+\begin{equation}
+P_{n+1}=P_{0}(1+r)^{n+1}
+\end{equation}
+
+in which $P_{0}$ stands for the initial population. This discrete model can be reshaped in to a continuous one in the following way:
+
+\begin{equation}
+\dot{P}(t)=rP(t)
+\end{equation}
+
+this is a relative simple differential equation with separable variable. Its integration gives:
+
+\begin{equation}
+P(t)=P_{0}(t)
+\end{equation}
+
+Malthus, of course, did not believed that the population could grow ad infinitum with an exponential growth, but since he estimated that the resources growth follows a linear path, he argued that, at the intersection of the two curves, a consistent part of the population will not have access to the resources.
+Thus  he expected that the exponential growth realizes only in the first part of the growth.
+Basing on these arguments, a Belgian mathematician Pierre-François Verhulst in 1838, proposed a model \cite{verhulst1838notice} in which the key point is the maximum number of people that the resources allows to live, usually called the carrying capacity here indicated as $K$ ($P$ and $r$ have the same meaning of the previous model):
+
+\begin{equation}
+\dot{P}=rP\left(1-\dfrac{P}{K} \right)
+\end{equation}
+
+that have the following analytic solution
+
+\begin{equation}
+P=\dfrac{KP_{0}e^{rt}}{K+P_{0}(e^{rt}-1)}
+\end{equation}
 

### 2020-03-27 Documentation/article_4.tex
update

@@
 
+In Fig.
+\ref{Malthusian_growth_vs_logistic_growth} this solution is compared to the Malthusian one: as one can point out the logistic growth has a saturation effect near the asymptote of carrying capacity.
+This represent the limit capacity that Malthus proposed to indicate as the intersection between the linear and exponential growth.
+It is worth nothing that near the origin the two growth are identical, then the logistic curve becomes linear up to the saturation region.
+
+\begin{figure}
+ \includegraphics[width=0.8\linewidth]{Figures/Malthusian_growth_vs_logistic_growth-2.jpg} 
+ \caption{The Malthusian growth (red line) compared to the logistic growth (blue line). The carrying capacity is marked with a black dashed line. Image taken from \cite{malthus_vs_logistic}}
+ \label{Malthusian_growth_vs_logistic_growth}
+\end{figure}
+
+\paragraph{Application to the epidemiology}
+
+The dynamics of the two models described before can be used to shape the spread of an infection \cite{serfling1952historical,ma2014estimating} in this case the population $P$ is replaced with the total number of infected people, $P_{0}$ the number of first infected people, $r$ the spreading rate and the carrying capacity $K$ with the maximum number of people that can be infected.
+The latter is the key point for the modelling: in principle this parameter would be equal to the population number, in practice due to the restrictive measure a large part of population can be removed from this computation: this lowers the predicted number of infected people.
+Thus we can argue that the effectiveness of the restrictive measures can be inferred from the linear behaviour of the cumulative curve (except, of course, if the numer of infected people is close to the population size).
+It is worth nothing that if the quarantine fails for also one/few infected people, the carrying capacity is increased up to the people that can be infected by these subjects.
+If this contribution is not negligible a new logistic growth may be found, with of course an exponential starting behaviour.
+Thus the prediction with the logistic curve should be taken as an evaluation of the effectiveness of restrictive measures and as the best scenario that can may realize.
+For this reason it is advisable to see the logistic curve as a local (in time) estimator.
+In principle one can model a differential set of equation were the carrying capacity is time dependent.
+This possibility may be considered  by the authors in a future version of this App.
+Furthermore it is worth nothing that other similar models such as Gopertz, Richards or Bertalanffy can be considered \cite{ma2014estimating} : the authors may consider to include them in  a future version of this App.
+
 %------------------------------------------------

### 2020-03-27 inst/shiny/C19/server/3module/analysis.R
try fix analysis

@@ -1,10 +1,10 @@
 ## REACTIVES OF THIS CODE
 reac_ARIMA <- shiny::reactiveValues(arimaOK = FALSE)
-    
-    # THIS IS THE REACTIVE CONTAINER OF TERRITORY SELECTION VARIABLES.
+
+# THIS IS THE REACTIVE CONTAINER OF TERRITORY SELECTION VARIABLES.
 t <- shiny::reactiveValues()
-    
-    ## Reactive values for global storage 
+
+## Reactive values for global storage 
 reac_general <- shiny::reactiveValues()
 
 ## CHECKS FOR ERROR PREVENTING ##
@@ -34,6 +34,7 @@ waitInput <- shiny::reactive({
 # DEPENDS ON REACTIVE t
 suggest_dates <- function()
 {
+  
   new = covid19:::const_trim(eval(t$data)[[t$name]]$totale_casi,1)
   index = which(eval(t$data)[[t$name]]$totale_casi %in% new)
   newdates = eval(t$data)[[t$name]]$data[index]
@@ -42,44 +43,25 @@ suggest_dates <- function()
 }
 
 # DEPENDS ON REACTIVES t AND reac_ARIMA! (but not on reac_ARIMA$arima)
-suggest_lags <- function()
+suggest_lags <- function(dates)
 {
-  checkExp({
-    sugg <- toString(forecast::auto.arima(log(reac_ARIMA$sample_cases_trim)))
-    matches <- regmatches(sugg, gregexpr("[[:digit:]]+", sugg))
-    return(as.numeric(unlist(matches)))
-  }, "There is no suitable ARIMA model")
+  
+  logic_interval <- eval(t$data)[[t$name]]$data >=  dates[1]  &
+    eval(t$data)[[t$name]]$data <= dates[2]
+  
+  sample_date <- eval(t$data)[[t$name]]$data
+  sample_cases <- eval(t$data)[[t$name]]$totale_casi
+  
+  sample_date_trim <- sample_date[logic_interval]
+  sample_cases_trim <- sample_cases[logic_interval]+1
+  
+  sugg <- toString(forecast::auto.arima(log(sample_cases_trim)))
+  matches <- regmatches(sugg, gregexpr("[[:digit:]]+", sugg))
+  return(as.numeric(unlist(matches)))
+  
 }
 
 
-
-output$terrInput <- shiny::renderUI({
-  shiny::fluidPage(
-    shiny::fluidRow(
-      shiny::column(4,
[…]

### 2020-03-27 inst/shiny/C19/ui/tabs/3.tab_analysis.R

@@ -15,7 +15,30 @@ shinydashboard::tabItem(
                         status= "danger",
                         solidHeader = FALSE,
                         title = "Territory selection",
-                        shiny::uiOutput("terrInput"),
+                        shiny::fluidPage(
+                          shiny::fluidRow(
+                            shiny::column(4,
+                                          shiny::selectizeInput(
+                                            inputId = "country", label = "Country",
+                                            choices = countryNames, selected = "Italy")
+                            ),
+                            shiny::column(4,
+                                          shiny::selectizeInput(
+                                            inputId = "region", label = "Region",
+                                            choices = c("--- ALL ---" = "", regNames), selected = NULL)
+                            ),
+                            shiny::column(4,
+                                          shiny::selectizeInput(
+                                            inputId = "province", label = "Province",
+                                            choices = c("--- ALL ---" = "", provNames), selected = NULL)
+                            )
+                          ),
+                          shiny::fluidRow(
+                            column(12, align = "center",
+                                   shiny::actionButton(inputId = "terr_go", label = "Submit")
+                            )
+                          )
+                        ),
                         shiny::htmlOutput("selected_info")
     )
   ),
@@ -40,7 +63,22 @@ shinydashboard::tabItem(
         status = "danger",
         solidHeader = TRUE,
         title = "Input",
-        shiny::uiOutput("fitInput")
+        fluidRow(
+          column(12,
+                 
+                 shiny::sliderInput(inputId = "fitInterval", label = "Choose fitting interval",
+                                    min = init_date, max = fin_date, timeFormat = "%d %b",
+                                    step = 1, value = c(init_date, fin_date)),
+                 shiny::checkboxInput(inputId = "swab_std", label = "Standardise positive cases by total swabs"),
+                 shiny::checkboxGroupInput(inputId = "plot_type", label = "Plot type",
+                                           choices = list("Cumulative cases" = 1, "New cases" = 2),
+                                           selected = 1),
+                 hr(),
+                 h3("Residuals"),
+                 shiny::selectInput(inputId = "plot_res_type", "Residuals plot type",choices =  c("Residuals","Residuals_standardized","Autocorrelation","Sqrt of abs of res vs fitted"),selected = "Residuals")
+                 
+          )
+        )
       )
     ),
     column(
[…]

### 2020-03-27 inst/shiny/C19/server/3module/analysis.R
analysis render UI moved in tab3 ui

@@ -52,35 +52,6 @@ suggest_lags <- function()
 }
 
 
-
-output$terrInput <- shiny::renderUI({
-  shiny::fluidPage(
-    shiny::fluidRow(
-      shiny::column(4,
-                    shiny::selectizeInput(
-                      inputId = "country", label = "Country",
-                      choices = countryNames, selected = "Italy")
-      ),
-      shiny::column(4,
-                    shiny::selectizeInput(
-                      inputId = "region", label = "Region",
-                      choices = c("--- ALL ---" = "", regNames), selected = NULL)
-      ),
-      shiny::column(4,
-                    shiny::selectizeInput(
-                      inputId = "province", label = "Province",
-                      choices = c("--- ALL ---" = "", provNames), selected = NULL)
-      )
-    ),
-    shiny::fluidRow(
-      column(12, align = "center",
-             shiny::actionButton(inputId = "terr_go", label = "Submit")
-      )
-    )
-  )
-  
-})
-
 # Province input updater
 shiny::observe({
   wait <- waitInput()
@@ -96,7 +67,7 @@ shiny::observe({
 
 
 # Assignment of territory variables when pressing action button
-shiny::observeEvent(input$terr_go, {
+shiny::observe( {
   if(input$region == "" && input$province == "") {
     t$name <- input$country
     t$data <- expression(countryTS)
@@ -116,29 +87,11 @@ shiny::observeEvent(input$terr_go, {
     t$r <- TRUE
     t$p <- FALSE
   }
+  
+  reac_general$sugg_dates = suggest_dates()
+  updateSliderInput(session,"arima_interval",timeFormat = "%d %b", step = 1, min =  init_date,max =  fin_date,value = c(reac_general$sugg_dates[1], reac_general$sugg_dates[2]))
+  
 })
-
-
-
-output$fitInput <- shiny::renderUI({
-  fluidRow(
-    column(12,
-           
-           shiny::sliderInput(inputId = "fitInterval", label = "Choose fitting interval",
-                              min = init_date, max = fin_date, timeFormat = "%d %b",
-                              step = 1, value = c(init_date, fin_date)),
-           shiny::checkboxInput(inputId = "swab_std", label = "Standardise positive cases by total swabs"),
-           shiny::checkboxGroupInput(inputId = "plot_type", label = "Plot type",
[…]

### 2020-03-27 inst/shiny/C19/ui/tabs/3.tab_analysis.R

@@ -15,7 +15,26 @@ shinydashboard::tabItem(
                         status= "danger",
                         solidHeader = FALSE,
                         title = "Territory selection",
-                        shiny::uiOutput("terrInput"),
+                        shiny::fluidPage(
+                          shiny::fluidRow(
+                            shiny::column(4,
+                                          shiny::selectizeInput(
+                                            inputId = "country", label = "Country",
+                                            choices = countryNames, selected = "Italy")
+                            ),
+                            shiny::column(4,
+                                          shiny::selectizeInput(
+                                            inputId = "region", label = "Region",
+                                            choices = c("--- ALL ---" = "", regNames), selected = NULL)
+                            ),
+                            shiny::column(4,
+                                          shiny::selectizeInput(
+                                            inputId = "province", label = "Province",
+                                            choices = c("--- ALL ---" = "", provNames), selected = NULL)
+                            )
+                          ),
+                         
+                        ),
                         shiny::htmlOutput("selected_info")
     )
   ),
@@ -40,7 +59,22 @@ shinydashboard::tabItem(
         status = "danger",
         solidHeader = TRUE,
         title = "Input",
-        shiny::uiOutput("fitInput")
+        fluidRow(
+          column(12,
+                 
+                 shiny::sliderInput(inputId = "fitInterval", label = "Choose fitting interval",
+                                    min = init_date, max = fin_date, timeFormat = "%d %b",
+                                    step = 1, value = c(init_date, fin_date)),
+                 shiny::checkboxInput(inputId = "swab_std", label = "Standardise positive cases by total swabs"),
+                 shiny::checkboxGroupInput(inputId = "plot_type", label = "Plot type",
+                                           choices = list("Cumulative cases" = 1, "New cases" = 2),
+                                           selected = 1),
+                 hr(),
+                 h3("Residuals"),
+                 shiny::selectInput(inputId = "plot_res_type", "Residuals plot type",choices =  c("Residuals","Residuals_standardized","Autocorrelation","Sqrt of abs of res vs fitted"),selected = "Residuals")
+                 
+          )
+        )
       )
     ),
     column(
@@ -50,7 +84,7 @@ shinydashboard::tabItem(
         status = "danger",
         title = "Plot",
         plotly::plotlyOutput("coolplot1") %>% shinycssloaders::withSpinner(color =
[…]

### 2020-03-27 Documentation/article_4.tex
update

@@
 \begin{figure}
+ \centering
  \includegraphics[width=0.8\linewidth]{Figures/Malthusian_growth_vs_logistic_growth-2.jpg} 
@@
 
+\clearpage
+
 \section{Panels} \label{Panels}
@@
 \subsection{Home}
+
+The home panel provides an overview of the coronavirus spread in Italy.
+The data shown is synchronised with the civil protection database.
+Each time the app is run it checks for updates whose occurrence is indicated in the "most recent updates" section.
+The home panel consists of 2 main sections: the choropleth map and the summary statistics.
+
+The choropleth map (Fig \ref{Home_fig1}) is an interactive heat map with breakouts by region and by province tracking the number of Covid-19 cases.
+
+\begin{figure}[h]
+ \centering
+ \includegraphics[width=0.5\linewidth]{Figures/Home_figure_1.png} 
+ \caption{}
+ \label{Home_fig1}
+\end{figure}
+
+Beyond the raw absolute number of total cases by province/region we propose two other indicators which introduce some form of normalisation to improve the comparability of the different geographical areas.
+These indicators are percentage and density (Fig.
+\ref{Home_fig2}).
+
+\begin{figure}[h]
+ \centering
+ \includegraphics[width=0.5\linewidth]{Figures/Home_figure_2.png} 
+ \caption{}
+ \label{Home_fig2}
+\end{figure}
+
+The percentage indicator is calculated dividing the total number of Covid cases by the population of the respective region/province (the latter retrieved from Istat).
+\begin{equation}
+percentage~cases = \frac{total~cases}{local~population} \dot 100
+\end{equation}
+This indicator ensures that the number of cases in more populated areas are not over-indexed compared to less populated ones. Similarly, it ensures less populated areas are not under-indexed.
+
+The density  indicator accounts for the territorial extension (in $km^2$) of regions/provinces expressed in parts per thousand (\textperthousand) and is calculated as follows:
+\begin{equation}
+density =  \dfrac{total~cases}{territorial~extension}\dot 1000
+\end{equation}
+This indicator ensures that we are not underestimating small regions/provinces with fewer cases than larger ones but with more concentrated cases per $Km^2$.
+
+The summary statistics section (Fig.
+\ref{Home_fig3}) provides some barometer level statistics regarding the current total number of Covid cases in Italy with three further breakouts: intensive care, hospitalised, and home isolation.
+
+\begin{figure}[h]
+ \centering
+ \includegraphics[width=0.5\linewidth]{Figures/Home_figure_3.png} 
+ \caption{}
+ \label{Home_fig3}
+\end{figure}
 

### 2020-03-28 Documentation/article_4.tex
update

@@
 
+The Data Inspection panel provides a data visualisation of the general information and a deeper inspection on the hospital occupancy, growth monitoring and test tracking for the coronavirus in Italy.
+It is divided in main two section: introduction and deeper inspection.
+In first chart \ref{fig:Inspection_general_info} of the introduction are presented the general information about the total cases, total recovered, total hospitalised and intensive care occupancy.
+It is possible to select the entire Italian country or a particular region or province.
+Moreover, it is possible to visualise and filter the raw data for country, region and province in the panel Raw Data \ref{fig:Inspection_rawdata} .In the deeper inspection there are 4 charts.
+The first two charts are in two different panels of a box and visualise information about intensive care occupancy in the Italian hospital in different regions, while the others represent the growth monitoring of total cases and test tracking .
+In the first chart \ref{fig:Inspection_perc_occupancy} is represented the percentage hospital occupancy in the selected day divided by capacity with respect to the initial intensive care capacity at the start of the pandemic.
+\begin{equation}
+percentage~occupancy = \dfrac{occupancy}{capacity} \times 100
+\end{equation}
+
+The second chart \ref{fig:Inspection_occupancy2} visualises a bar chart of the hospital occupancy and capacity of intensive care in different region of Italy at the selected day.The third chart \ref{fig:Inspection_growth_monitoring} represents the percentage growth and growth change of total cases day by day.
+The fifth chart \ref{fig:Inspection_test_tracking} visualises the daily cases with respect to the daily tests.
+
+\begin{figure}[h]
+\begin{subfigure}{.8\textwidth}
+  \centering
+  % include first image
+  \includegraphics[width=.8\linewidth]{Figures/Inspection_perc_occupancy.png}
+  \caption{}
+  \label{fig:Inspection_perc_occupancy}
+\end{subfigure} \hspace{0.3\textwidth}
+\begin{subfigure}{.8\textwidth}
+  \centering
+  % include second image
+  \includegraphics[width=.8\linewidth]{Figures/Inspection_rawdata.png} 
+  \caption{}
+  \label{fig:Inspection_rawdata}
+\end{subfigure} \hspace{0.3\textwidth}
+\begin{subfigure}{.8\textwidth}
+  \centering
+  % include second image
+  \includegraphics[width=.8\linewidth]{Figures/Inspection_test_tracking.png} 
+  \caption{}
+  \label{fig:Inspection_test_tracking}
+\end{subfigure} \hspace{0.3\textwidth}
+\begin{subfigure}{.8\textwidth}
+  \centering
+  % include second image
+  \includegraphics[width=.8\linewidth]{Figures/Inspection_Cattura.png} 
+  \caption{SEM image of the virus}
+  \label{fig:Inspection_Cattura}
+\end{subfigure} \hspace{0.3\textwidth}
+\caption{•}
+\end{figure}[h]
+
+\begin{figure}[h]
+\begin{subfigure}{.8\textwidth}
+  \centering
+  % include second image
+  \includegraphics[width=.8\linewidth]{Figures/Inspection_general_info.png} 
+  \caption{SEM image of the virus}
+  \label{fig:Inspection_general_info}
+\end{subfigure} \hspace{0.3\textwidth}
+\begin{subfigure}{.8\textwidth}
+  \centering
+  % include second image
+  \includegraphics[width=.8\linewidth]{Figures/Inspection_growth_monitoring.png} 
+  \caption{SEM image of the virus}
+  \label{fig:Inspection_growth_monitoring}
+\end{subfigure} \hspace{0.3\textwidth}
+\begin{subfigure}{.8\textwidth}
+  \centering
+  % include second image
+  \includegraphics[width=.8\linewidth]{Figures/Inspection_occupancy2.png} 
+  \caption{SEM image of the virus}
+  \label{fig:Inspection_occupancy2}
+\end{subfigure}\hspace{0.3\textwidth}
+\caption{ }
+\label{fig:Inspection_panels_2}
+\end{figure}
+
 \subsection{Data Analysis}

### 2020-03-28 Documentation/article_4.tex
upgrade

--- 2020-03-27 Documentation/article_4.tex
@@
 \usepackage{subcaption}
+\usepackage{float}
 \input{structure.tex} % Include the structure.tex file which specified the document structure and layout
@@
 The data shown is synchronised with the civil protection database.
-Each time the app is run it checks for updates whose occurrence is indicated in the "most recent updates" section.
+Each time the App is run it checks for updates whose occurrence is indicated in the "most recent updates" section.
 The home panel consists of 2 main sections: the choropleth map and the summary statistics.
@@
  \includegraphics[width=0.5\linewidth]{Figures/Home_figure_1.png} 
- \caption{}
+ \caption{Choropleth map by province showing absolute number of Covid-19 cases}
  \label{Home_fig1}
@@
  \includegraphics[width=0.5\linewidth]{Figures/Home_figure_2.png} 
- \caption{}
+ \caption{Input selector where the user can choose the three different breakouts absolute, percentage, and density}
  \label{Home_fig2}
@@
 
-The percentage indicator is calculated dividing the total number of Covid cases by the population of the respective region/province (the latter retrieved from Istat).
+The percentage indicator is calculated dividing the total number of Covid cases by the population of the respective region/province (the latter retrieved from Istat \cite{Istat+res0,Istat+res1}).
 \begin{equation}
@@
 The summary statistics section (Fig.
-\ref{Home_fig3}) provides some barometer level statistics regarding the current total number of Covid cases in Italy with three further breakouts: intensive care, hospitalised, and home isolation.
+\ref{Home_fig3}) provides some barometer level statistics regarding the current total number of COVID19 cases in Italy with three further breakouts: intensive care, hospitalised, and home isolation.
 
@@
  \includegraphics[width=0.5\linewidth]{Figures/Home_figure_3.png} 
- \caption{}
+ \caption{Boxes with syncronised barometer level statistic about the Coronavirus}
  \label{Home_fig3}
@@
 \subsection{Data Inspection}
+
+The Data Inspection panel provides a data visualisation of the general information and a deeper overview about the hospital occupancy, growth monitoring and test tracking for the COVID19 in Italy.
+It is divided in main two section: introduction and deeper inspection.
+In first chart (Fig.
+\ref{fig:Inspection_general_info}) of the introduction are presented the general information about the total cases, total recovered, total hospitalised and intensive care occupancy.
+It is possible to select the entire Italian country or a particular region or province.
+Moreover, it is possible to visualise and filter the raw data for country, region and province in the panel Raw Data (Fig.
+\ref{fig:Inspection_rawdata}).
+In the deeper inspection panel there are 4 charts: the first two charts are in two different panels of a box and visualise information about intensive care occupancy in the Italian hospital in different regions, while the others represent the growth monitoring of total cases and test tracking.
+It is worth nothing that the occupation may be higher than their total number, due to the fact that the places in intensive therapy that we considered may be upgraded (see the section \ref{Data Origin}).In the first chart \ref{fig:Inspection_perc_occupancy} is represented the percentage hospital occupancy in the selected day divided by capacity with respect to the initial intensive care capacity at the start of the pandemic.
+\begin{equation}
+percentage~occupancy = \dfrac{occupancy}{capacity} \times 100
+\end{equation}
+The second chart (Fig.
+\ref{fig:Inspection_occupancy2}) visualises a bar chart of the hospital occupancy and capacity of intensive care in different region of Italy at the selected day.The third chart (Fig.
+\ref{fig:Inspection_growth_monitoring}) represents the percentage growth and growth change of total cases day by day.
+The fifth chart (Fig.
+\ref{fig:Inspection_test_tracking}) visualises the daily cases with respect to the daily tests.
+
+\begin{figure}[H]
+\begin{subfigure}{.5\textwidth}
+  \centering
+  % include first image
+  \includegraphics[width=1\linewidth]{Figures/Inspection_perc_occupancy.png}
+  \caption{}
+  \label{fig:Inspection_perc_occupancy}
+\end{subfigure} 
+\begin{subfigure}{.5\textwidth}
+  \centering
+  % include second image
+  \includegraphics[width=1\linewidth]{Figures/Inspection_rawdata.png} 
+  \caption{}
+  \label{fig:Inspection_rawdata}
+\end{subfigure} 
+\begin{subfigure}{.5\textwidth}
+  \centering
+  % include second image
+  \includegraphics[width=1\linewidth]{Figures/Inspection_test_tracking.png} 
+  \caption{}
+  \label{fig:Inspection_test_tracking}
+\end{subfigure} 
+\begin{subfigure}{.5\textwidth}
+  \centering
+  % include second image
+  \includegraphics[width=1\linewidth]{Figures/Inspection_Cattura.png} 
+  \caption{SEM image of the virus}
+  \label{fig:Inspection_Cattura}
+\end{subfigure} 
+\caption{•}
+\end{figure}
+
+\begin{figure}[H]
+\begin{subfigure}{0.5\textwidth}
+  \centering
+  % include second image
+  \includegraphics[width=1\linewidth]{Figures/Inspection_general_info.png} 
+  \caption{SEM image of the virus}
+  \label{fig:Inspection_general_info}
+\end{subfigure} 
+\begin{subfigure}{0.5\textwidth}
+  \centering
+  % include second image
+  \includegraphics[width=1\linewidth]{Figures/Inspection_growth_monitoring.png} 
+  \caption{SEM image of the virus}
+  \label{fig:Inspection_growth_monitoring}
+\end{subfigure} 
+\begin{subfigure}{0.5\textwidth}
+  \centering
+  % include second image
+  \includegraphics[width=1\linewidth]{Figures/Inspection_occupancy2.png} 
+  \caption{SEM image of the virus}
+  \label{fig:Inspection_occupancy2}
+\end{subfigure}\hspace{0.3\textwidth}
+\caption{ }
+\label{fig:Inspection_panels_2}
+\end{figure}
 

### 2020-03-28 R/const_trim.R
added @export clause

@@ -12,6 +12,7 @@
 #' \dontrun{
 #' 	const_trim(c(0,0,0,1,1.3,1.7,2,3,4.5,6,9,12),1)
 #' }
+#' @export
 const_trim <- function(v,threshold) {
 	if(is.null(v) || length(v) == 0 || !is.numeric(v))
 		return(NULL)

### 2020-03-28 inst/shiny/C19/ui/tabs/4.tab_conclusion.R
Update 4.tab_conclusion.R

@@ -2,10 +2,86 @@ shinydashboard::tabItem(tabName = "tab_4",
         
  
         shinydashboard::valueBox(
-          "Conclusion",
           "Documentation",
+          "Authors, data source and documentation",
           icon = icon("chart-bar"),
           color = "red",
           width = NULL
         ),
+        
+        
+        shinydashboard::valueBox(
+          "Authors",
+          "Information",
+          icon = icon("chart-bar"),
+          color = "navy",
+          width = NULL
+        ),
+        
+        
+        shiny::fluidRow(
+          shiny::column(6,
+        shinydashboard::box(width = 12,
+                            status= "danger",
+                            solidHeader = FALSE,
+                            title = "Authors",
+                           
+                            shinydashboardPlus::userList(
+                              shinydashboardPlus::userListItem(
+                              src="https://avatars1.githubusercontent.com/u/36515398?s=460&u=152624d5b1720f1e66ffd3fbd43c6e13704c3e42&v=4",
+                              user_name = "Andrea Ierardi",
+                              description ="MSc. Data Science and Economics -Universita' degli studi di Milano        Graduated in Computer Science Universita' del Piemonte orientale(UPO)"
+                              ),
+                              shinydashboardPlus::userListItem(
+                                src="https://avatars3.githubusercontent.com/u/59971032?s=400&u=38395af942e20f0c9a395b2b30b12763a39ca8ac&v=4",
+                                user_name = "Gregorio Saporito",
+                                description ="Description ...."
+                              ),
+                                shinydashboardPlus::userListItem(
+                                  src="https://avatars3.githubusercontent.com/u/60003802?s=460&u=6463f4db28ebb1eaf463f117093a803af1b9cabc&v=4",
+                                  user_name = "Marzio De Corato",
+                                  description =paste0("M.Sc. Data Science and Economics student (UNIMI)       Ph.D in Physics and Nanoscience (UNIMO), M.Sc. in Physics (UNIMIB) (theoretical-computational solid state physics)")
+                                )),
+                            br(),
+                          
+                                          shinydashboardPlus::userList(
+                                  shinydashboardPlus::userListItem(
+                                    src="https://media-exp1.licdn.com/dms/image/C4D03AQFNxDX9vdxEmQ/profile-displayphoto-shrink_800_800/0?e=1590019200&v=beta&t=ftXpX3Wg-dB2t8sBoSYspTKC3z5cK2ehYhWk5obIBh4",
[…]

### 2020-03-28 inst/shiny/C19/server/2module/inspection.R
update css and bootstrap

@@ -159,26 +159,36 @@ output$rawData_sel_input <- shiny::renderUI({
 #   })
 # })
 
+is_ready <- function(x) {
+  if(( is.null(x) || length(x) == 0 ))
+    return(FALSE)
+  
+  return(TRUE)
+}
+
 shiny::observeEvent(input$rawData_go, {
   output$rawData_table <- DT::renderDataTable({
     
     
-    
-    DT::datatable( 
-      switch(input$rawData_terr,
-             "1" = countryTS$Italy %>% 
-               dplyr::select(-stato, -data_seriale) %>%
-               dplyr::filter(data >= input$rawData_date[1] &  data <= input$rawData_date[2]),
-             "2" = regionTS[[input$rawData_reg_sel]] %>%
-               dplyr::select(-stato,-lat,-long,-denominazione_regione,-codice_regione,-data_seriale) %>%
-               dplyr::filter(data >= input$rawData_date[1] &  data <= input$rawData_date[2]),
-             "3" = provTS[[input$rawData_prov_sel]] %>%
-               dplyr::select(-stato,-codice_provincia,-denominazione_provincia,-sigla_provincia,
-                             -lat,-long,-denominazione_regione,-codice_regione,-data_seriale) %>%
-               dplyr::filter(data >= input$rawData_date[1] &  data <= input$rawData_date[2])         
-      ), options = list(
-        searching = FALSE,
-        pageLength = 10,scrollX = T))
+    if(is_ready(input$rawData_terr))
+    {
+      DT::datatable( 
+        switch(input$rawData_terr,
+               "1" = countryTS$Italy %>% 
+                 dplyr::select(-stato, -data_seriale) %>%
+                 dplyr::filter(data >= input$rawData_date[1] &  data <= input$rawData_date[2]),
+               "2" = regionTS[[input$rawData_reg_sel]] %>%
+                 dplyr::select(-stato,-lat,-long,-denominazione_regione,-codice_regione,-data_seriale) %>%
+                 dplyr::filter(data >= input$rawData_date[1] &  data <= input$rawData_date[2]),
+               "3" = provTS[[input$rawData_prov_sel]] %>%
+                 dplyr::select(-stato,-codice_provincia,-denominazione_provincia,-sigla_provincia,
+                               -lat,-long,-denominazione_regione,-codice_regione,-data_seriale) %>%
+                 dplyr::filter(data >= input$rawData_date[1] &  data <= input$rawData_date[2])         
[…]

### 2020-03-28 inst/shiny/C19/ui/global/css.R

@@ -3,8 +3,14 @@ shinyjs::useShinyjs()
 
 # css
 tags$head(
+ 
+  tags$link(rel = "stylesheet", href = "https://stackpath.bootstrapcdn.com/bootstrap/4.4.1/css/bootstrap.min.css", integrity="sha384-Vkoo8x4CGsO3+Hhxv8T/Q5PaXtkKtu6ug5TOeNV6gBiFeWPGFN9MuhOf23Q9Ifjh", crossorigin="anonymous"),
   tags$link(rel = "stylesheet", type = "text/css", href = "custom.css"),
   tags$link(rel = "shortcut icon",
-            href = "coronavirus_black.png")
+            href = "coronavirus_black.png"),
+# tags$script(src="https://code.jquery.com/jquery-3.4.1.slim.min.js" , integrity="sha384-J6qa4849blE2+poT4WnyKhv5vZF5SrPo0iEjwBvKU7imGFAV0wwj1yYfoRSJoZ+n", crossorigin="anonymous"),
+  
+  tags$script(src="https://cdn.jsdelivr.net/npm/popper.js@1.16.0/dist/umd/popper.min.js", integrity="sha384-Q6E9RHvbIyZFJoft+2mJbHaEWldlvI9IOYy5n3zV9zzTtmI3UksdQRVvoxMfooAo" ,crossorigin="anonymous"),
+  tags$script(src="https://stackpath.bootstrapcdn.com/bootstrap/4.4.1/js/bootstrap.min.js", integrity="sha384-wfSDF2E50Y2D1uUdj0O3uMBJnjuUD4Ih7YwaYd1iqfktj0Uod8GCExl3Og8ifwB6",crossorigin="anonymous")
+  
 )
-

### 2020-03-28 inst/shiny/C19/ui/tabs/2.tab_inspection.R

@@ -75,11 +75,12 @@ shinydashboard::tabItem(tabName = "tab_2",
                                                  icon=icon("chart-line"),
                                                  color = "navy", width = NULL),
                         
-                        fluidRow(
+                      
                           shinydashboard::box(title="Intensive care information", solidHeader = T,
                                               width = NULL,
                                               color = "red",
                                               status = "danger",
+                                              fluidRow(
                           column(2,
                                
                                                      selectInput("occupancy_date",

### 2020-03-28 inst/shiny/C19/ui/tabs/4.tab_conclusion.R

@@ -20,7 +20,7 @@ shinydashboard::tabItem(tabName = "tab_4",
         
         
         shiny::fluidRow(
-          shiny::column(6,
+          shiny::column(8,
         shinydashboard::box(width = 12,
                             status= "danger",
                             solidHeader = FALSE,
@@ -30,38 +30,64 @@ shinydashboard::tabItem(tabName = "tab_4",
                               shinydashboardPlus::userListItem(
                               src="https://avatars1.githubusercontent.com/u/36515398?s=460&u=152624d5b1720f1e66ffd3fbd43c6e13704c3e42&v=4",
                               user_name = "Andrea Ierardi",
-                              description ="MSc. Data Science and Economics -Universita' degli studi di Milano        Graduated in Computer Science Universita' del Piemonte orientale(UPO)"
+                              description =HTML('<br> MSc. Data Science and Economics -Universita degli studi di Milano <br>  BC. in Computer Science Universita del Piemonte orientale(UPO) <br><br> 
+                                                       <a href="https://www.linkedin.com/in/andreaierardi/" rel="nofollow noreferrer">
+                                                       <img src="https://i.stack.imgur.com/gVE0j.png" alt="linkedin"> LinkedIn</a> &nbsp; 
+                                                       <a href="https://github.com/Andreaierardi" rel="nofollow noreferrer">
+                                                      <img src="https://i.stack.imgur.com/tskMh.png" alt="github"> Github</a> <br> 
+                                                      <a href="https://ierardiandrea.com/" rel="nofollow noreferrer">
+                                                      <i class="fas fa-cloud"></i> Website  </a> ')
                               ),
+                              
                               shinydashboardPlus::userListItem(
                                 src="https://avatars3.githubusercontent.com/u/59971032?s=400&u=38395af942e20f0c9a395b2b30b12763a39ca8ac&v=4",
                                 user_name = "Gregorio Saporito",
-                                description ="Description ...."
-                              ),
+                                description =HTML('<br><a href="https://www.linkedin.com/in/greg-saporito/" rel="nofollow noreferrer">
+                                                       <img src="https://i.stack.imgur.com/gVE0j.png" alt="linkedin"> LinkedIn</a>  &nbsp;
+                                                       <a href="https://github.com/gregorio-saporito" rel="nofollow noreferrer"> 
+                                                       <img src="https://i.stack.imgur.com/tskMh.png" alt="github"> Github </a> &nbsp; '
+                              )),
                                 shinydashboardPlus::userListItem(
[…]

### 2020-03-28 inst/shiny/C19/ui.R
update visual for mobile

@@ -2,7 +2,10 @@
 ## ui.R ##
 
 
-ui <- shinydashboardPlus::dashboardPagePlus(skin = "red", title = "Covid-19",
+ui <- fluidPage(
+  
+  HTML('<meta name="viewport" content="width=1024">'),
+   shinydashboardPlus::dashboardPagePlus(skin = "red", title = "Covid-19",
                         
                        shinydashboard::dashboardHeader(title = img(src = "coronavirus_white.png")),
                         
@@ -45,3 +48,4 @@ ui <- shinydashboardPlus::dashboardPagePlus(skin = "red", title = "Covid-19",
                         )
 )
 
+)

### 2020-03-28 inst/shiny/C19/ui/global/css.R

@@ -3,14 +3,15 @@ shinyjs::useShinyjs()
 
 # css
 tags$head(
- 
-  tags$link(rel = "stylesheet", href = "https://stackpath.bootstrapcdn.com/bootstrap/4.4.1/css/bootstrap.min.css", integrity="sha384-Vkoo8x4CGsO3+Hhxv8T/Q5PaXtkKtu6ug5TOeNV6gBiFeWPGFN9MuhOf23Q9Ifjh", crossorigin="anonymous"),
   tags$link(rel = "stylesheet", type = "text/css", href = "custom.css"),
-  tags$link(rel = "shortcut icon",
-            href = "coronavirus_black.png"),
-# tags$script(src="https://code.jquery.com/jquery-3.4.1.slim.min.js" , integrity="sha384-J6qa4849blE2+poT4WnyKhv5vZF5SrPo0iEjwBvKU7imGFAV0wwj1yYfoRSJoZ+n", crossorigin="anonymous"),
   
-  tags$script(src="https://cdn.jsdelivr.net/npm/popper.js@1.16.0/dist/umd/popper.min.js", integrity="sha384-Q6E9RHvbIyZFJoft+2mJbHaEWldlvI9IOYy5n3zV9zzTtmI3UksdQRVvoxMfooAo" ,crossorigin="anonymous"),
-  tags$script(src="https://stackpath.bootstrapcdn.com/bootstrap/4.4.1/js/bootstrap.min.js", integrity="sha384-wfSDF2E50Y2D1uUdj0O3uMBJnjuUD4Ih7YwaYd1iqfktj0Uod8GCExl3Og8ifwB6",crossorigin="anonymous")
-  
+  #FOR FUTURE IMPLEMENTATION OF BOOTSTRAP
+ # tags$link(rel = "stylesheet", href = "https://stackpath.bootstrapcdn.com/bootstrap/4.4.1/css/bootstrap.min.css", integrity="sha384-Vkoo8x4CGsO3+Hhxv8T/Q5PaXtkKtu6ug5TOeNV6gBiFeWPGFN9MuhOf23Q9Ifjh", crossorigin="anonymous"),
+ # 
+ # tags$link(rel = "shortcut icon",
+ #           href = "coronavirus_black.png"),
[…]

### 2020-03-28 inst/shiny/C19/ui/tabs/2.tab_inspection.R

@@ -19,6 +19,7 @@ shinydashboard::tabItem(tabName = "tab_2",
                                                      
                                                      .nav-tabs-custom > .nav-tabs > li.active {
                                                      border-top-color: red !important;
+                                                     float:left;
                                                      }
                                                      
                                                      .btn-default {

### 2020-03-28 inst/shiny/C19/ui/tabs/4.tab_conclusion.R

@@ -108,6 +108,18 @@ shinydashboard::tabItem(tabName = "tab_4",
           width = NULL
         ),
         
+        shiny::fluidRow(
+          shiny::column(8,
+        shinydashboard::box(width = 12,
+                            status= "danger",
+                            solidHeader = FALSE,
+                            title = "Documentation",
+                            br(),
+                            h3("To be completed.."),
+                            hr(),
                             
+        tags$iframe(style="height:600px; width:100%", src="C:/Users/AndreDany/Desktop/GitHub/COVID_19_HACK/inst/shiny/C19/documentation/documentation_29032020.pdf")
+))
+        )
   
 )

### 2020-03-28 SEIR_2.py
Delete SEIR_2.py

@@ -1,129 +1,0 @@
-# -*- coding: utf-8 -*-
-"""
-Created on Wed Mar 25 16:07:41 2020
-
-@author: Io
-"""
-
-import numpy as np
-import pandas as pd
-
-from sklearn.linear_model import LinearRegression
-from scipy.integrate import odeint
-
-IT = np.array([0, 0, 0.05, 0.1, 0.15, 0.2, 0.3, 0.1, 0.1]) # Incubation time 
-TT = np.array([0.2, 0.5, 0.3]) # Test time
-IBST = np.flip(np.array([0.7, 0.2, 0.1])) # IBST(tau) = P(infective at day of symptoms - tau)
-
-
-m = max(len(IT), len(TT), len(IBST))
-N = 60500000
-
-def import_italy():
-    
-    url = 'https://raw.githubusercontent.com/pcm-dpc/COVID-19/master/dati-andamento-nazionale/dpc-covid19-ita-andamento-nazionale.csv'
-    df = pd.read_csv(url)
-    
-    P = df.totale_casi.values   # tested positive cases
-    E = (P-P)[:-m]
-    I = (P-P)[:-m]
-    
-    for t in range(len(P)-m):
-        IT.resize(len(P)-t, refcheck=False)
-        TT.resize(len(P)-t, refcheck=False)
-        IBST.resize(len(P)-t, refcheck=False)
-        
-        E[t] = ((IT+TT)/2).dot(P[t:])
-        I[t] = ((IBST+TT)/2).dot(P[t:])
-    
-    R = df.dimessi_guariti.values + df.deceduti.values
-    R = R[:-m]
-    S = N-E-I-R
-    
-    refined_data = [S, E, I, R]
-    
-    return refined_data, P[:-m]
-    
-def SEIR_model(U, t, param):
-        S, E, I, R = U
-        alpha, beta, gamma, epsilon = param
-        return np.array([
-                -beta*S*I + alpha*R,    #susceptible
-                beta*S*I - epsilon*E,   #exposed
-                epsilon*E - gamma*I,    #infected
-                gamma*I - alpha*R])     #removed
-    
-def der(U):
-    dU = np.empty_like(U)
-    i = 0
-    
-    for X in U:
-        dU[i] = np.array([X[1]-X[0] if t == 0
-            else X[-1]-X[-2] if t == len(X)-1
-            else (X[t+1]-X[t-1])/2 
-            for t in range(len(X))])
-        i += 1
-        
[…]

### 2020-03-28 SEIR_lib.py
Delete SEIR_lib.py

@@ -1,283 +1,0 @@
-"""
-Estimating a SEIR model on Italian data:
-
-S_der = -beta*S*I + alpha*R
-E_der = beta*S*I - epsilon*E
-I_der = epsilon*E - gamma*I
-R_der = gamma*I - alpha*R
-
-Restricted SEIR: 
-    alpha = immunity_loss_time = 0
-    beta = R_0*gamma
-    gamma = 1/T_inf = 1/2.5   
-    epsilon = 1/T_inc = 1/5.5
-    
-
-
-Incubation Time (T_inc = 5.5) data from : 
-https://www.jwatch.org/na51083/2020/03/13/covid-19-incubation-period-update
-
-Infectious Period (T_inf ~ 2.5) data from : 
-http://gabgoh.github.io/COVID/index.html    
-
-Data pulled from civil protection:
-https://raw.githubusercontent.com/pcm-dpc/COVID-19
-
-MODELLING ASSUMPTIONS per function:
-    import_[...] :  Incubation Time is 6 days (from mean of 5.5)
-                    Italian population is 60500000
-                    Data is T_inc days old
-                    
-    der :           X_der[T] = dX/dt (T) is approximable with (X[T+1]-X[T-1])/2 in inner points
-    
-PROBLEMS WITH CURRENT MODEL:
-    The parameters (alpha, beta, gamma, epsilon), though estimated in a sensible manner,
-        produce a model that does not fit the data at all
-    Dataset is from 24/02 only 
-    
-    
-    
-    
-    
-"""
-
-import numpy as np
-import pandas as pd
-import matplotlib.pyplot as plt
-
-from sklearn.linear_model import LinearRegression
-from scipy.integrate import odeint
-
-
-# Global Parameters
-
-T_inc = 5.5 # Incubation time
-T_inc_int = int(np.ceil(T_inc)) # Approx to int
-T_inf = 2.5 # Infectiveness time
-First_measures = -3 # First quarantine measures, 21/02/2020
-Local_measures = 1 # Strong quarantine in Lombardia & some other provinces, 25/02/2020
-National_measures = 13 # Strong quarantine in all of Italy, 8/03/2020
-
-
-"""Import data from civil protection, returns data series S, E, I, R"""
-
-def import_italy():   
-    # E and R are observed, I is estimated from E[+T_inc], S = N-E-I-R
[…]

### 2020-03-28 SEIR_lib_R0.py
Delete SEIR_lib_R0.py

@@ -1,66 +1,0 @@
-# -*- coding: utf-8 -*-
-"""
-Created on Thu Mar 19 12:39:41 2020
-
-@author: Io
-"""
-
-import SEIR_lib
-import pandas as pd 
-import matplotlib.pyplot as plt
-
-#%% Import
-
-total, national = SEIR_lib.import_italy()
-S, E, I, R = total
-S_n, I_n, E_n, R_n = national   
-
-N = (S + E + I + R)[0]
-S_der = SEIR_lib.der(S)
-E_der = SEIR_lib.der(E)
-I_der = SEIR_lib.der(I)
-R_der = SEIR_lib.der(R)
-
-S_n_der = pd.Series(float)
-S_n_der.pre = SEIR_lib.der(S_n.pre)
-S_n_der.post = SEIR_lib.der(S_n.post)
-I_n_der = pd.Series(float)
-I_n_der.pre = SEIR_lib.der(I_n.pre)
-I_n_der.post = SEIR_lib.der(I_n.post)
-E_n_der = pd.Series(float)
-E_n_der.pre = SEIR_lib.der(E_n.pre)
-E_n_der.post = SEIR_lib.der(E_n.post)
-R_n_der = pd.Series(float)
-R_n_der.pre = SEIR_lib.der(R_n.pre)
-R_n_der.post = SEIR_lib.der(R_n.post)
-
-#%% Fitting with restricted parameters
-
-parameters, R_sq = SEIR_lib.estimate_restricted_parameters(S, E, I, E_der)
-alpha, beta, gamma, epsilon = parameters
-R_0 = SEIR_lib.estimate_R_0(N, beta, gamma)
-
-parameters_pre, R_sq_pre = SEIR_lib.estimate_restricted_parameters(S_n.pre, E_n.pre, I_n.pre, E_n_der.pre)
-alpha_pre, beta_pre, gamma_pre, epsilon_pre = parameters_pre
-R_0_pre = SEIR_lib.estimate_R_0(N, beta_pre, gamma_pre)
-
-parameters_post, R_sq_post = SEIR_lib.estimate_restricted_parameters(S_n.post, E_n.post, I_n.post, E_n_der.post)
-alpha_post, beta_post, gamma_post, epsilon_post = parameters_post
-R_0_post = SEIR_lib.estimate_R_0(N, beta_post, gamma_post)
-
-print('R_0 Total           : ', R_0.round(2), 'R_sq : ', R_sq.round(2), sep='\t')
[…]

### 2020-03-28 SEIR_lib_test.py
Delete SEIR_lib_test.py

@@ -1,82 +1,0 @@
-"""
-Showcase for the SEIR_lib module
-
-"""
-
-import SEIR_lib 
-import matplotlib.pyplot as plt
-
-#%% Import
-
-total,_ = SEIR_lib.import_italy()
-S, E, I, R = total
-
-N = (S + E + I + R)[0]
-S_der = SEIR_lib.der(S)
-E_der = SEIR_lib.der(E)
-I_der = SEIR_lib.der(I)
-R_der = SEIR_lib.der(R)
-
-#%% Fitting with unrestricted parameters
-
-beta, epsilon, R_sq_be = SEIR_lib.estimate_beta_epsilon(S, I, E, E_der)
-alpha, gamma, R_sq_ag = SEIR_lib.estimate_alpha_gamma(I, R, S_der, E_der, I_der)
-R_0 = SEIR_lib.estimate_R_0(N, beta, gamma)
-
-print('alpha : ', alpha)
-print('beta : ', beta)
-print('gamma : ', gamma)
-print('epsilon : ', epsilon)
-print('R_sq_be : ', R_sq_be)
-print('R_sq_ag : ', R_sq_ag)
-print('R_0 : ', R_0, '\n')
-
-Y0 = [S[0], E[0], I[0], R[0]] # initial conditions
-Y = I    # actual data points of infected
-parameters = (alpha, beta, gamma, epsilon)
-time_steps = range(50*len(S))   # in order to forecast in long distance 
-time_data = range(len(S))   # the actual range of time for which data is available
-
-SEIR_lib.plot_SEIR(Y0, parameters, time_steps, time_data, Y)
-
-#%% Fitting with restricted parameters
-
-parameters_r, R_sq_r = SEIR_lib.estimate_restricted_parameters(S, E, I, E_der)
-alpha_r, beta_r, gamma_r, epsilon_r = parameters_r
-R_0_r = SEIR_lib.estimate_R_0(N, beta_r, gamma_r)
-
-print('beta_r : ', beta_r)
-print('gamma_r : ', gamma_r)
-print('epsilon_r : ', epsilon_r)
-print('R_sq_r : ', R_sq_r)
-print('R_0_r : ', R_0_r)
-
-Y0 = [S[0], E[0], I[0], R[0]] # initial conditions
-Y = I    # actual data points of infected
-parameters_r = tuple(parameters_r)
-time_steps = range(50*len(S))   # in order to forecast in long distance 
[…]

### 2020-03-28 SEIR_2_test.py
Add files via upload
SEIR_2 is a python module containing the core of data gathering and SEIR fitting. No extended documentation is yet available, this is a toy version. 
SEIR_2_test is a showcase fro SEIR_2
SEIR testing parameter estimation with known model does what the name says: the methods used in SEIR_2 are tested to recover known parameters of SEIR-generated data points

@@ -1,0 +1,99 @@
+# -*- coding: utf-8 -*-
+"""
+Created on Wed Mar 25 16:23:10 2020
+
+@author: Io
+"""
+
+from SEIR_2 import *
+import matplotlib.pyplot as plt
+from scipy.interpolate import *
+
+U, P = import_italy()
+
+k=1
+
+t = range(len(P))
+t_k = range(k*len(P))
+
+S, E, I, R = U
+
+# R0 = 2.2
+# alpha = 0
+# beta = R0/N
+# gamma = 0.05
+# epsilon = 1
+alpha, beta, gamma, epsilon = None, None, None, None
+
+
+sol, param, Rsq = solve_SEIR(U, k=k, alpha=alpha, beta=beta, gamma=gamma, epsilon=epsilon)
+
+S_, E_, I_, R_ = sol
+
+plt.plot(t_k, E_, 'r', label='SEIR Exposed')
+plt.plot(t_k, I_, 'g', label='SEIR Infected')
+plt.plot(t_k, R_, 'y', label='SEIR Removed')
+
+plt.plot(t, P, 'bo', label='Positive')    
+plt.plot(t, E, 'ro', label='Exposed')
+plt.plot(t, I, 'go', label='Infected')
+plt.plot(t, R, 'yo', label='Removed')
+plt.legend()
+
+# plt.show()
+
+dict = {0:'alpha', 1:'beta', 2:'gamma', 3:'epsilon', 4:'Rsq'}
+for i in range(len(param)+1):
+    if i<len(param):
+        print(dict.get(i), param[i], sep = '\t:\t')
+    else: 
+        print(dict.get(i), Rsq, sep = '\t:\t')
+        
+#%%    LOGISTIC, SPLINE: data interpolation. more resolution
+        
+# param_l, cov_l = logistic_fit(U)
+# sup, g_rate, cen = param_l
+
+# # dict_l = {0:'Superior', 1:'Growth Rate', 2:'Flex Point'}
+# # for i in range(len(param_l)):
+# #     print(dict_l.get(i), param_l[i], sep = '\t:\t')
+
+# z = [logistic(x, sup, g_rate, cen) for x in t_triple]
+
+# plt.plot(t_triple, z)
+# plt.plot(t, I, 'o')
+
+
+# #%% SEIR WITH DATA FROM LOGISTIC and interpolation
+
[…]

### 2020-03-28 inst/shiny/C19/server/3module/analysis.R
update new box

@@ -525,4 +525,13 @@ output$arima_shell_output <- shiny::renderPrint({
   
 })
 
+output$arima_shell_resid <- shiny::renderPrint({
+  
+  wait <- waitLoading()
+  if(reac_ARIMA$arimaOK) {
+    checkExp(forecast::checkresiduals(reac_ARIMA$arima) , "There is not a suitable ARIMA model")
+  }
+  
+})
+
 

### 2020-03-28 inst/shiny/C19/ui/tabs/3.tab_analysis.R

@@ -241,6 +241,17 @@ shinydashboard::tabItem(
                                                                                            "#dd4b39"),
         width = 12
       )
+    ),
+    column(
+      6,
+      shinydashboard::box(
+        color = "red",
+        status = "danger",
+        title = "Arima residuals output",
+        shiny::verbatimTextOutput("arima_shell_resid") %>% shinycssloaders::withSpinner(color =
+                                                                                           "#dd4b39"),
+        width = 12
+      )
     )
   )
 )

### 2020-03-28 inst/shiny/C19/server/3module/analysis.R
edit gray box

@@ -529,7 +529,9 @@ output$arima_shell_resid <- shiny::renderPrint({
   
   wait <- waitLoading()
   if(reac_ARIMA$arimaOK) {
-    checkExp(forecast::checkresiduals(reac_ARIMA$arima) , "There is not a suitable ARIMA model")
+   
+   checkExp(forecast::checkresiduals(reac_ARIMA$arima,plot=FALSE), "There is not a suitable ARIMA model")
+    
   }
   
 })

### 2020-03-28 inst/shiny/C19/ui/tabs/3.tab_analysis.R

@@ -231,6 +231,7 @@ shinydashboard::tabItem(
   
   
   fluidRow(
+   
     column(
       6,
       shinydashboard::box(
@@ -242,12 +243,21 @@ shinydashboard::tabItem(
         width = 12
       )
     ),
+    
+    #background-color: #f5f5f5; */
+
     column(
       6,
       shinydashboard::box(
         color = "red",
         status = "danger",
         title = "Arima residuals output",
+        tags$head(tags$style(HTML('
+        pre
+                                  {
+                                  background-color: #ffffff;
+                                  }
+                                  '))),
         shiny::verbatimTextOutput("arima_shell_resid") %>% shinycssloaders::withSpinner(color =
                                                                                            "#dd4b39"),
         width = 12

### 2020-03-28 inst/shiny/C19/server/global/sidebar.R
fix documentation sidebar

@@ -4,6 +4,6 @@ output$menu <-   shinydashboard::renderMenu({
             shinydashboard::menuItem("Home", tabName = "tab_1", icon=icon("home")),
             shinydashboard::menuItem("Inspection", tabName = "tab_2", icon=icon("search")),
             shinydashboard::menuItem("Analysis", tabName = "tab_3", icon=icon("chart-line")),
-            shinydashboard::menuItem("Conclusion", tabName = "tab_4", icon=icon("calendar-check"))
+            shinydashboard::menuItem("Documentation", tabName = "tab_4", icon=icon("calendar-check"))
   )
 })

### 2020-03-28 inst/shiny/C19/ui/tabs/4.tab_conclusion.R

@@ -117,8 +117,11 @@ shinydashboard::tabItem(tabName = "tab_4",
                                                             hr(),
                                                             
                                                             HTML('<ul>
-                                                            <li> <a href="https://github.com/pcm-dpc/COVID-19" rel="nofollow noreferrer">  Protezione Civile </a> </li>
-                                                            <li> <a href="https://github.com/pcm-dpc/COVID-19" rel="nofollow noreferrer">  Istat </a> </li>
+                                                            <li> <h4> <a href="https://github.com/pcm-dpc/COVID-19" rel="nofollow noreferrer">  Protezione Civile - General information  </a> </h4> </li>
+                                                            
+                                                            <li><h4>  <a href="http://dati.istat.it/Index.aspx?DataSetCode=DCIS_POPRES1">  Istat - Population  </a> </h4> </li>                                                                                                                           
+                                                            <li><h4>  <a href="https://www.istat.it/it/files//2015/04/Superfici-delle-unit%C3%A0-amministrative-Dati-comunali-e-provinciali.zip">  Istat - Territory </a> </h4> </li>
+                                                            <li><h4>  <a href="https://www.corriere.it/cronache/20_marzo_16/coronavirus-quanti-posti-terapia-intensiva-ci-sono-italia-quanti-ne-arriveranno-0fbafa76-678a-11ea-93a4-da8ab3a8afb1.shtml" rel="nofollow noreferrer">   Corriere.it - Hospital intensive care capacity </a>  </h4></li>
 
                                                             </ul>
                                                                  ')

### 2020-03-28 Documentation/article_4.tex
update

@@
 \usepackage{float}
+\usepackage{epigraph}
+\usepackage{mathcomp}
 \input{structure.tex} % Include the structure.tex file which specified the document structure and layout
-
+\sloppy
 \hyphenation{Fortran hy-phen-ation} % Specify custom hyphenation points in words with dashes where you would like hyphenation to occur, or alternatively, don't put any dashes in a word to stop hyphenation altogether
@@
 \maketitle % Print the title/author/date block
+
+\newpage
+
+\epigraph{ "The laws of history are as absolute as the laws of physics, and if the probabilities of error are greater, it is only because history does not deal with as many humans as physics does atoms, so that individual variations
+count for more" (I.Asimov, Foundation and Empire) }
+
+\newpage
 
@@
 
-\paragraph{Population growth dynamics}
+\subsection{Population growth dynamics}
 The dynamics of population was founded by Thomas Malthus in 1817 \cite{malthus1817essay} with his well known equation:
@@
 P=\dfrac{KP_{0}e^{rt}}{K+P_{0}(e^{rt}-1)}
+\label{logistic_equation}
 \end{equation}
@@
 
-\paragraph{Application to the epidemiology}
+\subsection{Application to the epidemiology}
 
@@
 Furthermore it is worth nothing that other similar models such as Gopertz, Richards or Bertalanffy can be considered \cite{ma2014estimating} : the authors may consider to include them in  a future version of this App.
+
+\subsection{Time series approach}
+
+\subsection{An ab-initio approach: the SEIR model}
 
@@
 \begin{equation}
-percentage~cases = \frac{total~cases}{local~population} \dot 100
+percentage~cases = \frac{total~cases}{local~population} \times 100
 \end{equation}
@@
 
-The density  indicator accounts for the territorial extension (in $km^2$) of regions/provinces expressed in parts per thousand (\textperthousand) and is calculated as follows:
+The density  indicator accounts for the territorial extension (in $km^2$) of regions/provinces expressed in parts per thousand (  \textperthousand ) and is calculated as follows:
 \begin{equation}
-density =  \dfrac{total~cases}{territorial~extension}\dot 1000
+density =  \dfrac{total~cases}{territorial~extension}\times 1000
 \end{equation}
@@
 
-\subsection{Data Inspection}
+\subsection{Data Inspection}\label{Data Inspection}
 
@@
 
+Panel \textit{Analysis} explores different mathematical models for shaping the infections time series, namely the ones introduced in \ref{Theoretical Background}.
+As previously argued, each of these models predictive power hinges on uncertain actual scenarios, such as the effectiveness of restrictive measures, the presence of several independent and spread outbreaks, and so on.
+Package \cite{growthcurver} has been used for growth curve fitting.
+This package provides function \textit{SummarizeGrowth}, which inputs the time series of sample data and outputs the model represented by the growth metrics.
+Such parameters include those of the logistic equation that best fit the data \ref{logistic_equation}.
+The reasons why this package was chosen are chiefly two: at the one hand, it performs non-linear curve fitting with the Levenberg-Marquardt algorithm \cite{more1978levenberg}, which happens to be one of the most robust non-linear least squares  (NLS) algorithms.
+For this step, function \cite{nlsLM} is used.
+On the other hand, the package is endowed with two methods for background correction.
+
+At the top of analysis panel the user is allowed to pick one dataset among national, by region and by province.
+Hence, any following analysis will use the selected dataset as input (the user can browse the raw dataset in panel inspection, see \ref{Data Inspection} ).
+The first section is devoted to the logistic model.
+An input box \ref{fig:logistic_input} lists several dynamic inputs:
+\begin{itemize}
+
+\item The date range that the calculator will use for curve fitting.
+The option to choose initial and final dates different from those at, resepectively, the beginning and the end of the dataset, is useful in two ways: first, it allows to cut initial dates which possibly correspond to zero or constant infections data, as a consequence of a delay in the outbreak in that territory; second, it allows to compute the availabe prediction n days ahead, by cutting the last n dates, and compare it to the real data.
+\item A check-box that, if selected, will use for computations and rendering the data of the share of infected among the tested, instead of the the infected themselves.
+That is, it will "standardise positive cases by total swabs".
+This may be useful to capture the "selection bias", but only in the scenario in which tests are randomly assigned to the population, or at least are assigned with a logic common to all dates and territories.
+This option is not currently available for provinces.
+\item Plot type selection boxes: should cumulative cases and fitted curve be plotted or new cases, or both (The user can always show and hide plots' components by clicking on their label in the legend).
+\item Residuals plot type.
+Four options for graphically rendering of the residuals are available: Residuals vs.
+fitted values; Standardized residuals vs.
+fitted values; Autocorrelation; Square root of absolute residuals vs.
+fitted values.
+Output is displayed within three boxes: Inside the \textit{Plot} box \ref{fig:logistic_plot1} the sample points, the fitted logistic curve and a confidence band at 95 \% level for the logistic curve are overlayed.
+The user can view the sample and fitted values by hovering over or touching (for touch-screen devices) them.
+In addition, if \textit{new cases} is selected, a plot of sample cases differences (which correspond to new cases, in fact) is shown over the logistic estimated distribution \ref{fig:logistic_plot2}.
+The latter is simply obtained by deriving the right hand side of formula \ref{logistic_equation}
+\end{itemize}
+\begin{equation}
+\dot{N}= \frac{(\frac{K-N_0}{N_0})Kre^{-rt}}{[1+(\frac{K-N_0}{N_0})e^{-rt}]^2}
+\end{equation}
+
+\begin{figure}[H]
+  \centering
+\begin{subfigure}{0.6\textwidth}
+  % include second image
+  \includegraphics[width=1\linewidth]{Figures/logistic_input.png} 
+  \caption{SEM image of the virus}
+  \label{fig:logistic_input}
+\end{subfigure} 
+\hspace{5.5cm}
+\begin{subfigure}{0.6\textwidth}
+  % include second image
+  \includegraphics[width=1\linewidth]{Figures/logistic_plot1.png} 
+  \caption{SEM image of the virus}
+  \label{fig:logistic_plot1}
+\end{subfigure} 
+\hspace{5.5cm}
+\begin{subfigure}{0.6\textwidth}
+  % include second image
+  \includegraphics[width=1\linewidth]{Figures/logistic_plot2.png} 
+  \caption{SEM image of the virus}
+  \label{fig:logistic_plot2}
+\end{subfigure}
+\caption{ }
+\label{fig:logistic_plots_set}
+\end{figure}
+
+\clearpage
+
 \section{Conclusions}

### 2020-03-29 Documentation/article_4.tex
upgrade

--- 2020-03-28 Documentation/article_4.tex
@@
 \usepackage{float}
+\usepackage{epigraph}
+\usepackage{mathcomp}
+\usepackage{textcomp}
 \input{structure.tex} % Include the structure.tex file which specified the document structure and layout
-
+\sloppy
 \hyphenation{Fortran hy-phen-ation} % Specify custom hyphenation points in words with dashes where you would like hyphenation to occur, or alternatively, don't put any dashes in a word to stop hyphenation altogether
@@
 Later the virus spread almost in all regions of Italy with a higher density in Lombardy, Veneto, Emilia-Romagna, Piemonte and Marche.
-Starting from 22 of February, the Italian government started to impose the quarantine (red-zone) for 11 different municipalities, in particular Codogno, Casalpusterlengo, Lodi (included the neighbour municipalities) and Vo' .
+Starting from 22 of February, the Italian government started to impose the quarantine (red-zone) for eleven different municipalities, in particular Codogno, Casalpusterlengo, Lodi (included the neighbour municipalities) and Vo' .
 Starting from this date different restrictive measures were imposted starting from the regions with the highest number of cases: public entrainment were almost suspended as well as schools and universities.
@@
 
-The home panel provides an overview of the coronavirus spread in Italy.
-The data shown is synchronised with the civil protection database.
-Each time the App is run it checks for updates whose occurrence is indicated in the "most recent updates" section.
-The home panel consists of 2 main sections: the choropleth map and the summary statistics.
-
-The choropleth map (Fig \ref{Home_fig1}) is an interactive heat map with breakouts by region and by province tracking the number of Covid-19 cases.
+The home panel provides an overview of the COVID19 spread in Italy.
+The data shown are synchronised with the civil protection database.
+Whenever the App is started,  a check for the updates is performed and their occurrence is indicated in the \textit{ most recent updates }section.
+The home panel consists of two main sections: the choropleth map and the summary statistics.The choropleth map (Fig \ref{Home_fig1}) is an interactive heat map with breakouts by region and by province tracking the number of Covid-19 cases.
 
@@
 
-The percentage indicator is calculated dividing the total number of Covid cases by the population of the respective region/province (the latter retrieved from Istat \cite{Istat+res0,Istat+res1}).
+The percentage indicator is calculated dividing the total number of COVID19 cases by the population of the respective region/province (the latter retrieved from Istat \cite{Istat+res0,Istat+res1}).
 \begin{equation}
-percentage~cases = \frac{total~cases}{local~population} \dot 100
+percentage~cases = \frac{total~cases}{local~population} \times 100
 \end{equation}
-This indicator ensures that the number of cases in more populated areas are not over-indexed compared to less populated ones. Similarly, it ensures less populated areas are not under-indexed.
-
-The density  indicator accounts for the territorial extension (in $km^2$) of regions/provinces expressed in parts per thousand (\textperthousand) and is calculated as follows:
+This indicator ensures that the number of cases in more populated areas are not over-indexed compared to less populated ones.
+Similarly, it ensures less populated areas are not under-indexed.
+The density  indicator accounts for the territorial extension (in $km^2$) of regions/provinces expressed in parts per thousand (  \textperthousand ) and is calculated as follows:
 \begin{equation}
-density =  \dfrac{total~cases}{territorial~extension}\dot 1000
+density =  \dfrac{total~cases}{territorial~extension}\times 1000
 \end{equation}
 This indicator ensures that we are not underestimating small regions/provinces with fewer cases than larger ones but with more concentrated cases per $Km^2$.
-
-The summary statistics section (Fig.
+Finally, the summary statistics section (Fig.
 \ref{Home_fig3}) provides some barometer level statistics regarding the current total number of COVID19 cases in Italy with three further breakouts: intensive care, hospitalised, and home isolation.
@@
 
-\subsection{Data Inspection}
+\subsection{Data Inspection}\label{Data Inspection}
 
-The Data Inspection panel provides a data visualisation of the general information and a deeper overview about the hospital occupancy, growth monitoring and test tracking for the COVID19 in Italy.
+The \textit{Data Inspection} panel provides a data visualisation of the general information and a deeper overview about the hospital occupancy, growth monitoring and test tracking for the COVID19 in Italy.
 It is divided in main two section: introduction and deeper inspection.
@@
 \ref{fig:Inspection_rawdata}).
-In the deeper inspection panel there are 4 charts: the first two charts are in two different panels of a box and visualise information about intensive care occupancy in the Italian hospital in different regions, while the others represent the growth monitoring of total cases and test tracking.
-It is worth nothing that the occupation may be higher than their total number, due to the fact that the places in intensive therapy that we considered may be upgraded (see the section \ref{Data Origin}).In the first chart \ref{fig:Inspection_perc_occupancy} is represented the percentage hospital occupancy in the selected day divided by capacity with respect to the initial intensive care capacity at the start of the pandemic.
+In the deeper inspection panel there are four charts: the first two charts are in two different panels of a box and visualise information about intensive care occupancy in the Italian hospital in different regions, while the others represent the growth monitoring of total cases and test tracking.
+It is worth nothing that the occupation may be higher than their total number, due to the fact that the places in intensive therapy that we considered may be upgraded (see the section \ref{Data Origin}).
+In the first chart \ref{fig:Inspection_perc_occupancy} is represented the percentage hospital occupancy in the selected day divided by capacity with respect to the initial intensive care capacity at the start of the pandemic.
 \begin{equation}
@@
 
+Panel \textit{Analysis} explores different mathematical models for shaping the infection\textquotesingle s time series, namely the ones introduced in Section \ref{Theoretical Background}.
+As previously argued, each of these model\textquotesingle s predictive power hinges on uncertain actual scenarios, such as the effectiveness of restrictive measures, the presence of several independent and spread outbreaks, and so on.
+
+Package  \cite{growthcurver} has been used for growth curve fitting.
+This package provides function  \textit{SummarizeGrowth}, which inputs the time series of sample data and outputs the model represented by the growth metrics.
+Such parameters include those of the logistic equation that best fit the data, namely $K$, $P_0$ and $r$, as in Eq.\ref{logistic_equation}.
+The reasons why this package was chosen are chiefly two.
+At the one hand, it performs non-linear curve fitting with the Levenberg-Marquardt algorithm \cite{more1978levenberg} , which happens to be one of the most robust nls algorithms.
+For this step, function \cite{nlsLM}  is used.
+On the other hand, the package is endowed with methods for background correction.
+
+%----> add here packages and documentation for ARIMA <----
+
+At the top of analysis panel the user is allowed to pick one dataset among national, by region and by province.
+Hence, any following analysis will use the selected dataset as input (the user can browse the raw dataset in panel inspection, see paragraph \ref{Data Inspection} ).
+
+The first section is devoted to the logistic model. An input box (Fig.\ref{fig:logistic_input}) lists several dynamic inputs:
+\begin{enumerate}
+\item A date range that the calculator will use for curve fitting.
+The option to choose initial and final dates different from those at, respectively, the beginning and the end of the dataset, is useful in two ways:
+\begin{itemize} 
+\item It allows to cut initial dates which possibly correspond to zero or constant infections data, as a consequence of a delay in the outbreak in that territory.
+Actually, an automatic process will do so after the choice of territory, adjusting the initial date to the first day whose new infections exceed a given threshold (currently set to 1).
+\item It allows to compute the available prediction n days ahead, by cutting the last n dates, and compare it to the real data.
+\end{itemize}
+\item A check-box \emph{standardise positive cases by total swabs}.
+If selected, the share of infected among the tested, instead of the infected themselves, will be used for computations and rendering.
+This may be useful to capture the selection bias, but only in the scenario in which tests are randomly assigned to the population, or at least are assigned with a logic common to all dates and territories.
+This option is not currently available for provinces.
+\item Plot type selection boxes.
+The available plot types are cumulative cases and new cases.
+(The user can always show and hide plot\textquotesingle  s components by clicking on their label in the legend).
+\item Residuals plot type.
+Four options for graphically rendering the residuals: Residuals vs.
+fitted values; Standardized residuals vs.
+fitted values; Autocorrelation; Square root of absolute residuals vs.
+fitted values.
+\end{enumerate}
+
+The output is displayed within three boxes: 
+\begin{enumerate}
+\item Inside the \ref{fig:logistic_plot1} box the sample points, the fitted logistic curve and a confidence band at 95 \% level for the logistic curve are overlayed.
+The user can view the sample and fitted values by hovering over or touching (for touch-screen devices) them \ref{fig:logistic_plot1}.
+In addition, if \emph{new cases} is selected, a plot of sample cases differences (which correspond to new cases, in fact) is shown over the logistic estimated distribution \ref{fig:logistic_plot2}..
+The latter is simply obtained by deriving the right hand side of formula $(4.3.1)$:
+\begin{equation}
+N'(t) = \frac{(\frac{K-N_0}{N_0})Kre^{-rt}}{[1+(\frac{K-N_0}{N_0})e^{-rt}]^2}
+\end{equation}
+The user is suggested to take full advantage of plotly features.
+Besides showing points labels when passing over it and enabling plot selection by clicking on label items, plotly charts are endowed with a sequence of tools shown at the top-right of the box \ref{fig:logistic_plot3}.
+\item Summary output box \ref{fig:logistic_smry1} contains principal information about the selected model and a list of goodness-of-fit tests on the residuals. 
+\item Residuals box contains the plots of residuals, rendered in the user-chosen fashion. (see Fig.\ref{fig:logistic_res1}).
+\end{enumerate}
+
+\begin{figure}[H]
+  \centering
+\begin{subfigure}{0.6\textwidth}
+  % include second image
+  \includegraphics[width=1\linewidth]{Figures/logistic_input.png} 
+  \caption{Set of useful tools for navigating into plots, by plotly.}
+  \label{fig:logistic_input}
+\end{subfigure} 
+\hspace{5.5cm}
+\begin{subfigure}{0.6\textwidth}
+  % include second image
+  \includegraphics[width=1\linewidth]{Figures/logistic_plot1.png} 
+  \caption{Summary output box of logistic section.}
+  \label{fig:logistic_plot1}
+\end{subfigure} 
+\hspace{5.5cm}
+\begin{subfigure}{0.6\textwidth}
+  % include second image
+  \includegraphics[width=1\linewidth]{Figures/logistic_plot2.png} 
+  \caption{One plot of nls residuals: square root of absolute residuals vs. fitted values.}
+  \label{fig:logistic_plot2}
+\end{subfigure}
+\caption{ }
+\label{fig:logistic_plots_set}
+\end{figure}
+
+\begin{figure}[H]
+  \centering
+\begin{subfigure}{0.6\textwidth}
+  % include second image
+  \includegraphics[width=1\linewidth]{Figures/logistic_plot3.png} 
+  \caption{Set of useful tools for navigating into plots, by plotly.}
+  \label{fig:logistic_plot3}
+\end{subfigure} 
+\hspace{5.5cm}
+\begin{subfigure}{0.6\textwidth}
+  % include second image
+  \includegraphics[width=1\linewidth]{Figures/logistic_res1.png} 
+  \caption{Summary output box of logistic section.}
+  \label{fig:logistic_res1}
+\end{subfigure} 
+\hspace{5.5cm}
+\begin{subfigure}{0.6\textwidth}
+  % include second image
+  \includegraphics[width=1\linewidth]{Figures/logistic_smry1.png} 
+  \caption{One plot of nls residuals: square root of absolute residuals vs. fitted values.}
+  \label{fig:logistic_smry1}
+\end{subfigure}
+\caption{ }
+\label{fig:logistic_plots_set2}
+\end{figure}
+
+\clearpage
+
 \section{Conclusions}

### 2020-03-29 R/prova_eval.R
removing prova eval

@@ -1,6 +1,0 @@
-provaa <- function(bool) {
-  a <- 40
-  b <- expression(a)
-  if(TRUE)
-    paste("Risultato:", ifelse(bool, eval(b), ""), " ciao")
-}

### 2020-03-29 inst/shiny/C19/server/3module/analysis.R
upload exefit with ::: notation

@@ -137,7 +137,7 @@ output$coolplot1 <- plotly::renderPlotly({
   sample_cases_rem <- sample_cases[!logic_interval]
   sample_diff_rem <- sample_diff[!logic_interval]
   
-  fit_data <- exe_fit(sample_cases = sample_cases_trim,
+  fit_data <- covid19:::exe_fit(sample_cases = sample_cases_trim,
                       sample_date = sample_serial_date_trim,
                       days = days)
   

### 2020-03-29 inst/shiny/C19/ui/tabs/4.tab_conclusion.R

@@ -45,7 +45,7 @@ shinydashboard::tabItem(tabName = "tab_4",
                                                               shinydashboardPlus::userListItem(
                                                                 src="https://avatars1.githubusercontent.com/u/36515398?s=460&u=152624d5b1720f1e66ffd3fbd43c6e13704c3e42&v=4",
                                                                 user_name = "Andrea Ierardi",
-                                                                description =HTML('<br> MSc. Data Science and Economics - Universita degli studi di Milano <br>  BC. in Computer Science Universita del Piemonte orientale(UPO) <br><br> 
+                                                                description =HTML('<br> MSc. Data Science and Economics - Universita degli studi di Milano <br>  BC. in Computer Science <br> Universita del Piemonte orientale(UPO) <br><br> 
                                                        <a href="https://www.linkedin.com/in/andreaierardi/" rel="nofollow noreferrer">
                                                        <img src="https://i.stack.imgur.com/gVE0j.png" alt="linkedin"> LinkedIn</a> &nbsp; 
                                                        <a href="https://github.com/Andreaierardi" rel="nofollow noreferrer">

### 2020-04-02 inst/shiny/C19/global.R
fix library in description, implement difference in general info tab2 - inspection

@@ -5,7 +5,7 @@ regionTS = covid19:::get_regionTS()
 provTS = covid19:::get_provTS()
 country_growth = covid19:::get_country_growth()
 intensivecare_capacity = covid19:::get_intensivecare_cap(regionTS)
-age_cases = get_agecases(as.character(sort(names(regionTS))))
+age_cases = covid19:::get_agecases(as.character(sort(names(regionTS))))
 
 N <- nrow(countryTS)
 

### 2020-04-02 inst/shiny/C19/server/2module/inspection.R

@@ -12,7 +12,15 @@ shiny::observe({
 })
 
 
-
+shiny::observe(
+  {
+    if(input$difference == 2)
+    {
+      reac_dataset$dataset = diff(reac_dataset$dataset)
+      reac_dataset$name = paste0(reac_dataset$name, " difference")
+    }
+  }
+)
 # General info reactive dataset
 shiny::observe({
   

### 2020-04-02 inst/shiny/C19/ui/tabs/2.tab_inspection.R

@@ -49,6 +49,10 @@ shinydashboard::tabItem(tabName = "tab_2",
                                                                                                            , provNames
                                                                                                            ), selected = NULL),
                                                                                             
+                                                                                            radioButtons("difference", label = "Select cumulative or difference plot",
+                                                                                                         choices = list("Choice 1" = 1, "Choice 2" = 2), 
+                                                                                                         selected = 1),
+                                                                                            
                                                                                             hr(),
                                                                                             p("For province, only total case data are available")
                                                                               ),

### 2020-04-02 inst/shiny/C19/server/2module/inspection.R
fix inspection

@@ -12,19 +12,12 @@ shiny::observe({
 })
 
 
-shiny::observe(
-  {
-    if(input$difference == 2)
-    {
-      reac_dataset$dataset = diff(reac_dataset$dataset)
-      reac_dataset$name = paste0(reac_dataset$name, " difference")
-    }
-  }
-)
+
 # General info reactive dataset
 shiny::observe({
   
   if(input$regiontab2 == "default" && input$provincetab2 == "default") {
+    
     reac_dataset$name <- input$countrytab2
     reac_dataset$dataset <- countryTS$Italy
     
@@ -38,6 +31,22 @@ shiny::observe({
   
   if(input$provincetab2 == "default")
   {
+    
+    if(is_ready(reac_dataset$totale_casi))
+    {
+      if(input$difference == 2)
+      {
+        reac_dataset$dataset$totale_casi = diff( reac_dataset$dataset$totale_casi)
+        reac_dataset$dataset$terapia_intensiva = diff( reac_dataset$dataset$terapia_intensiva)
+        reac_dataset$dataset$totale_ospedalizzati = diff( reac_dataset$dataset$totale_ospedalizzati)
+        reac_dataset$dataset$deceduti = diff( reac_dataset$dataset$deceduti)
+        reac_dataset$dataset$dimessi_guariti = diff( reac_dataset$dataset$dimessi_guariti)
+        
+        
+      }
+    }
+    
+    
     reac_dataset$plot = highcharter::hchart(reac_dataset$dataset,"spline",title= "General info",highcharter::hcaes(x=data,y = totale_casi),  name="Total cases", color="blue", yAxis = 1,showInLegend=TRUE) %>% 
       highcharter::hc_chart(zoomType = "xy") %>%
       highcharter::hc_yAxis_multiples(
@@ -48,8 +57,8 @@ shiny::observe({
                                  yAxis = 1, highcharter::hcaes(x = data, y = terapia_intensiva),
                                  name="Total Intesive care", color="red",showInLegend=TRUE) %>%
       highcharter::hc_add_series(data =reac_dataset$dataset, type = "spline", 
-                                 yAxis = 1, highcharter::hcaes(x = data, y = ricoverati_con_sintomi),
-                                 name="Total Hospitalized", color="orange",showInLegend=TRUE)   %>%
+                                 yAxis = 1, highcharter::hcaes(x = data, y = totale_ospedalizzati),
[…]

### 2020-04-02 inst/shiny/C19/ui/tabs/2.tab_inspection.R

@@ -50,7 +50,7 @@ shinydashboard::tabItem(tabName = "tab_2",
                                                                                                            ), selected = NULL),
                                                                                             
                                                                                             radioButtons("difference", label = "Select cumulative or difference plot",
-                                                                                                         choices = list("Choice 1" = 1, "Choice 2" = 2), 
+                                                                                                         choices = list("Cumulative" = 1, "Difference" = 2), 
                                                                                                          selected = 1),
                                                                                             
                                                                                             hr(),

### 2020-04-02 Documentation/article_4.tex
Update

@@
 
-\paragraph{Background} \label{Background}
+\subsection{Background} \label{Background}
 In December 2019 different cases of pneumonia were reported in Wuhan (China) \cite{huang2020clinical}.
@@
 The restrictive measures were also extend up to all Italy with a further decree \cite{11marzo}.
+
+\subsection{Historical Background}
+
+In the last centuries different cases of pandemic diseases were recorded: from our point of view the dynamics of them it is useful to understand, ad perhaps to model,  that further peaks may succeed the first one also with a highest death rate.
+
+\paragraph{The  deadliest pandemics in human history: the Spanish flu}
+
+ The first one, the biggest one known for humans (in terms of spread as well for deaths \cite{potter2001history} ), is the Spanish Flu: this name is quite misleading since it was not originated in Spain, but the Spanish newspaper were the first one to talk about because large part of the involved countries were involved in the Great War (1914-1918).
+The origin of this disease is still under research: some scholars focused on the  United States \cite{crosby_2003,barry2004site}on France \cite{shanks2016no} and on Asia \cite{langford2005did}.
+The disease was ascribed to the virus A/H1N1 a subtype of influenza A Virus.
+The spread of the disease was strongly amplified by the fact that many people lived very close in to a very difficult hygienic conditions.
+As one can see from Fig.
+\ref{SpanishFLU} there are basically three waves: the initial spread, the second wave and the third wave.
+The second wave, that is the most lethal one, coincides with the end of the war and with the coming back of the troops: the close contact inside the trains, and the diffusion of the troops inside their home villages/cities increased abruptly the contagion rate and then the death rate.
+Furthermore this effect was amplified by the fact that a more deadly mutation of the virus diffused  \cite{barry2008cross}.
+As pointed out by different scholars such effect was enhanced since, the infected people, for which the severity of the ill was higher were transferred by train  into the hospitals and so  the virus spread with a higher rate \cite{wever2014death}.As it is possible to note from Fig.
+\ref{Mortality_spanish_flu} this second wave was largely more deadly for young people with respect the first wave.
+Finally, as can be noted in the plot \ref{SpanishFLU}, a third wave also manifested in the beginning of 1919: in this case the mortality was higher with respect to the first one but lower to the second one.
+In the end this pandemic infected 500 million people and a large part of them , from 17 to 50 million, died \cite{spreeuwenberg2018reassessing}.
+It is worth nothing that, just to give a comparison, the causalities in the Great War were from 8.8 to 10.7 million among soldiers and 11 million among civilian \cite{haythornthwaite1996world}
+
+\begin{figure}[h]
+ \centering
+ \includegraphics[width=1\linewidth]{Figures/SpanishFLU.png} 
+ \caption{Pneumonia and Influenza mortality from 1910-1922 in London: note the tree peaks due to the spanish flu.
+In the in box the variation of the diffusion rate R$_{0}$ is reported.
+Image taken from \cite{he2011mechanistic} }
+ \label{SpanishFLU}
+\end{figure}
+
+\begin{figure}[h]
+ \centering
+ \includegraphics[width=0.6\linewidth]{Figures/Mortality_spanish_flu.png} 
+ \caption{Morthality range distribution for the two waves of Spanish flu: note that the second one was more severe for the young people. Image taken from \cite{taubenberger20061918}}
+ \label{Mortality_spanish_flu}
+\end{figure}
 

### 2020-04-02 inst/shiny/C19/server/2module/inspection.R
Update inspection.R

@@ -15,56 +15,72 @@ shiny::observe({
 
 # General info reactive dataset
 shiny::observe({
-  
+
   if(input$regiontab2 == "default" && input$provincetab2 == "default") {
     
-    reac_dataset$name <- input$countrytab2
-    reac_dataset$dataset <- countryTS$Italy
+    if(input$difference == 1)
+    {
+      reac_dataset$name <- input$countrytab2
+      reac_dataset$dataset <- countryTS$Italy 
+    }
+    else
+    {
+      reac_dataset$name <- paste0("diff ", input$countrytab2)
+      
+      reac_dataset$dataset$totale_casi = diff(c(NA,  countryTS$Italy$totale_casi))
+      reac_dataset$dataset$terapia_intensiva = diff( c(NA, countryTS$Italy$terapia_intensiva))
+      reac_dataset$dataset$totale_ospedalizzati = diff(c(NA,  countryTS$Italy$totale_ospedalizzati))
+      reac_dataset$dataset$deceduti = diff(c(NA,  countryTS$Italy$deceduti))
+      reac_dataset$dataset$dimessi_guariti = diff(c(NA,  countryTS$Italy$dimessi_guariti))
+    }
+    
     
   }
   
   
   else if(input$regiontab2 != "default" &&  input$provincetab2 == "default") {
-    reac_dataset$name <- input$regiontab2
-    reac_dataset$dataset <- regionTS[[input$regiontab2]]
+    
+    if(input$difference == 1)
+    {
+      reac_dataset$name <- input$regiontab2
+      
+      reac_dataset$dataset <- regionTS[[input$regiontab2]]
+    }
+    
+    else
+    {
+      reac_dataset$name <- paste0("diff ",input$regiontab2)
+      
+      reac_dataset$dataset$totale_casi = diff(c(NA, regionTS[[input$regiontab2]]$totale_casi))
+      reac_dataset$dataset$terapia_intensiva = diff( c(NA,regionTS[[input$regiontab2]]$terapia_intensiva))
+      reac_dataset$dataset$totale_ospedalizzati = diff(c(NA, regionTS[[input$regiontab2]]$totale_ospedalizzati))
+      reac_dataset$dataset$deceduti = diff(c(NA, regionTS[[input$regiontab2]]$deceduti))
+      reac_dataset$dataset$dimessi_guariti = diff(c(NA, regionTS[[input$regiontab2]]$dimessi_guariti))
+    }
   }
   
[…]

### 2020-04-02 inst/shiny/C19/server/2module/inspection.R
add barplot for difference

@@ -16,6 +16,19 @@ shiny::observe({
 # General info reactive dataset
 shiny::observe({
 
+  if(input$difference ==1)
+  {
+    reac_dataset$plot_type = "spline"
+    reac_dataset$pointWidth = 0
+    reac_dataset$yAxis = 1
+  }
+  else 
+  {
+    reac_dataset$plot_type = "column"
+    reac_dataset$pointWidth = 30
+    reac_dataset$yAxis = 0
+  }
+  
   if(input$regiontab2 == "default" && input$provincetab2 == "default") {
     
     if(input$difference == 1)
@@ -63,26 +76,28 @@ shiny::observe({
   {
     
     
-    reac_dataset$plot = highcharter::hchart(reac_dataset$dataset,type = "column",title= "General info",highcharter::hcaes(x=data,y = totale_casi),  name="Total cases", color="blue", yAxis = 1,showInLegend=TRUE) %>% 
+    reac_dataset$plot = highcharter::hchart(reac_dataset$dataset,type =reac_dataset$plot_type,title= "General info",highcharter::hcaes(x=data,y = totale_casi),  name="Total cases", color="blue", yAxis = reac_dataset$yAxis,pointWidth= reac_dataset$pointWidth ,showInLegend=TRUE) %>% 
       highcharter::hc_chart(zoomType = "xy") %>%
       highcharter::hc_yAxis_multiples(
         list(lineWidth = 3, title = list(text  =  '')),
         list(showLastLabel = TRUE, opposite = TRUE, title = list(text  =  ''))
       )  %>%
-      highcharter::hc_add_series(data = reac_dataset$dataset, type = "column", 
-                                 yAxis = 1, highcharter::hcaes(x = data, y = terapia_intensiva),
-                                 name="Total Intesive care", color="red",showInLegend=TRUE) %>%
-      highcharter::hc_add_series(data =reac_dataset$dataset, type = "column", 
-                                 yAxis = 1, highcharter::hcaes(x = data, y = totale_ospedalizzati),
+      highcharter::hc_add_series(data =reac_dataset$dataset, type = reac_dataset$plot_type, 
+                                 yAxis = reac_dataset$yAxis,pointWidth= reac_dataset$pointWidth,  highcharter::hcaes(x = data, y = totale_ospedalizzati),
                                  name="Total symptomatic", color="orange",showInLegend=TRUE)   %>%
[…]

### 2020-04-02 SEIR.R
Add files via upload
seir rewritten in R. documentation for seir v1 still stuff missing. problems with optimise function to minimise least square

@@ -1,0 +1,159 @@
+#R SEIR
+library(deSolve)
+library(covid19)
+library(ggplot2)
+
+#Data loading & global variables
+lomb <- get_regionTS()$Lombardia
+poplomb <- italy_pop$region[italy_pop$region$territorio == "Lombardia", "valore"]
+P <- lomb$totale_casi
+R <- lomb$totale_casi - lomb$totale_positivi
+
+IT <- c(0,0,0,0,0,1)
+TT <- c(1)
+IBST <- c(1)
+m = max(length(IT), length(TT), length(IBST))
+ITl = length(IT)
+TTl = length(TT)
+IBSTl = length(IBST)
+N = 10500000
+latent_time = IT%*%(1:length(IT)-1)
+
+
+
+#%% Functions
+refine_data <- function(P, R) {
+	
+	n_obs = length(P)
+	
+	S = integer(n_obs-m)
+	E = integer(n_obs-m)
+	I = integer(n_obs-m)
+	R = R[1:(n_obs-m)]
+	
+	for(t in (1:(n_obs-m))){
+		IT_temp = c(IT, integer(n_obs-ITl-t+1))
+		TT_temp = c(TT, integer(n_obs-TTl-t+1))
+		IBST_temp = c(IBST, integer(n_obs-IBSTl-t+1))
+		
+		E[t] <- ((IT_temp+TT_temp)/2) %*% P[t:n_obs]
+		I[t] <- ((IBST_temp+TT_temp)/2) %*% P[t:n_obs]
+	}
+	
+	S = N-E-I-R
+	
+	return(data.frame(S, E, I, R))
+	
+}
+
+SEIR_model <- function(t, state, parameters){
+	with(as.list(c(state, parameters)), {
+		dS <- alpha*R-beta*S*I
+		dE <- beta*S*I-sigma*E
+		dI <- sigma*E-gamma*I
+		dR <- gamma*I-alpha*R
+	
+		list(c(dS, dE, dI, dR))
+	})
+}
+
+solve_SEIR <- function(U0, eval_time=eval_time, parameters=parameters){
+	state <- c( S = U0[1], E=U0[2], I=U0[3], R=U0[4])
+	
+	
+	out <- ode(y=state, times=eval_time, func=SEIR_model, parms=parameters)
+	out <- data.frame(out)
+	return(out)
+}
+
[…]

### 2020-04-02 inst/shiny/C19/global.R
growth calculation in gllobal for map

@@ -50,8 +50,14 @@ names(pc_data) <- tolower(names(pc_data))
 
 pc_df <- purrr::map_df(names(pc_data), function(x){
   casi <- tail(pc_data[[x]],1)$totale_casi
-  dplyr::data_frame(name=x,cases=casi)
+  casi_vecchi <- tail(pc_data[[x]],2)$totale_casi[[1]]
+  dplyr::data_frame(name=x,cases=casi, cases_old=casi_vecchi)
 })
+
+pc_df <- pc_df %>%
+  dplyr::ungroup() %>% 
+  dplyr::mutate(growth=round(((cases-cases_old)/cases_old)*100,2) ) %>% 
+  dplyr::select(-cases_old)
 
 pc_df$name
 
@@ -74,7 +80,8 @@ pc_df <- pc_df %>%
  dplyr::mutate(name=ifelse(name%in%c("trento","bolzano","p.a. trento","p.a. bolzano"),
                      "trentino-alto adige/sudtirol",name)) %>%
   dplyr::group_by(name) %>%
-  dplyr::summarise(cases=sum(cases), 
+  dplyr::summarise(cases=sum(cases),
+            growth=sum(growth),
             pop=sum(pop),
             ext=sum(ext)) %>%
  dplyr::mutate(name=ifelse(name=="emilia romagna","emilia-romagna",name))
@@ -99,15 +106,22 @@ dfita1 <- dfita1 %>%
 
 
 
+
 # --- province ---
 
 clean_prov <- purrr::map_df(names(provTS), function(x) {
   tail(provTS[[x]],1)$totale_casi
   dplyr::data_frame(
     name=x,
-    cases=tail(provTS[[x]],1)$totale_casi
+    cases=tail(provTS[[x]],1)$totale_casi,
+    cases_old=tail(provTS[[x]],2)$totale_casi[[1]]
   )
 })
+
+clean_prov <- clean_prov %>%
+  dplyr::ungroup() %>% 
+  dplyr::mutate(growth=round(((cases-cases_old)/cases_old)*100,2) ) %>% 
+  dplyr::select(-cases_old)
 
 
 url <- "http://code.highcharts.com/mapdata/countries/it/it-all.geo.json"

### 2020-04-02 inst/shiny/C19/server/1module/home.R
integrated growth in shiny

@@ -15,7 +15,7 @@ output$map_region <- output$map_region_modal <- highcharter::renderHighchart(
           
           highcharter::hc_add_series_map(map = map, df = dfita1,
                                          joinBy = "id", value = input$map_value, name="percentage (cases/pop * 100)")
-      } else {
+      } else if(input$map_value=="density") {
         highcharter::highchart(type = "map") %>% 
           highcharter::hc_chart(zoomType = "xy") %>%
           
@@ -23,6 +23,14 @@ output$map_region <- output$map_region_modal <- highcharter::renderHighchart(
                                          joinBy = "id", value = input$map_value, name="density (cases/km^2 * 1000)") %>%
           highcharter::hc_colorAxis(
             stops = highcharter::color_stops(4,c("#d8ebb5","#639a67","#2b580c","#003000")))
+      } else {
+        highcharter::highchart(type = "map") %>% 
+          highcharter::hc_chart(zoomType = "xy") %>%
+          
+          highcharter::hc_add_series_map(map = map, df = dfita1,
+                                         joinBy = "id", value = input$map_value, name="growth") %>%
+          highcharter::hc_colorAxis(
+            stops = highcharter::color_stops(4,c("#E6E6FA","#D8BFD8","#BA55D3","#800080")))
       }
   )
 
@@ -48,7 +56,7 @@ output$map_province <- output$map_province_modal <- highcharter::renderHighchart
       highcharter::hc_chart(zoomType = "xy") %>%
       highcharter::hc_add_series_map(map = ita, df = dfita2, 
                                      joinBy = "hasc", value = input$map_value, name="percentage (cases/pop * 100)")
-  } else {
+  } else if(input$map_value=="density") {
     highcharter::highchart(type = "map") %>% 
       highcharter::hc_chart(zoomType = "xy") %>%
       
@@ -56,6 +64,14 @@ output$map_province <- output$map_province_modal <- highcharter::renderHighchart
                                      joinBy = "hasc", value = input$map_value, name="density (cases/km^2 * 1000)") %>%
       highcharter::hc_colorAxis(
         stops = highcharter::color_stops(4,c("#d8ebb5","#639a67","#2b580c","#003000")))
+  } else {
+    highcharter::highchart(type = "map") %>% 
+      highcharter::hc_chart(zoomType = "xy") %>%
+      
+      highcharter::hc_add_series_map(map = ita, df = dfita2, 
[…]

### 2020-04-02 inst/shiny/C19/ui/tabs/1.tab_home.R

@@ -38,7 +38,7 @@ shinydashboard::tabItem(tabName = "home",
                                                              
                                                              selectInput(inputId = "map_value",
                                                                          label = NULL,
-                                                                         choices = c("absolute", "percentage", "density"),
+                                                                         choices = c("absolute", "percentage", "density", "growth"),
                                                                          width="150px"),
                                                              
                                                              #   shinyWidgets::pickerInput(inputId = "map_value",

### 2020-04-02 R/get_temperatures.R
updated get_temperatures

@@ -23,6 +23,9 @@ get_temperatures <- function() {
   provnames <- italy_ext$province$territorio
   temperatures <- data.frame()
   
+  init_date <- as.Date("2020-02-24")
+  fin_date <- Sys.Date()
+  
   for(prov in provnames) {
     skip <- FALSE
     url1 <- paste("https://www.ilmeteo.it/portale/archivio-meteo/",
@@ -31,30 +34,42 @@ get_temperatures <- function() {
     url2 <- paste("https://www.ilmeteo.it/portale/archivio-meteo/",
                   prov,
                   "/2020/Marzo?format=csv", sep = "")
+    url3 <- paste("https://www.ilmeteo.it/portale/archivio-meteo/",
+                  prov,
+                  "/2020/Aprile?format=csv", sep = "")
     tryCatch( {tab1 <- read.csv(url1, sep = ";")
-               tab2 <- read.csv(url2, sep = ";")}, 
+               tab2 <- read.csv(url2, sep = ";")
+               tab3 <- read.csv(url3, sep = ";")}, 
               error = function(e){ skip <<- TRUE }
               )
-    if(skip || nrow(tab1) == 0 || nrow(tab2) == 0) { next }
+    if(skip || nrow(tab1) == 0 || nrow(tab2) == 0 || nrow(tab3) == 0) { next }
     
     tab1 <- tab1 %>%
       dplyr::rename(loc = 1, date = 2, temp = 3, hum = 7, press = 12) %>%
       dplyr::select(loc,date,temp, hum, press) %>%
       dplyr::mutate(date = as.Date(date, format = "%d/%m/%Y")) %>%
       dplyr::mutate(loc = prov) %>%
-      dplyr::filter(date >= as.Date("2020-02-24"))
+      dplyr::filter(date >= init_date && date <= fin_date)
     
     tab2 <- tab2 %>%
       dplyr::rename(loc = 1, date = 2, temp = 3, hum = 7, press = 12) %>%
       dplyr::select(loc,date,temp, hum, press) %>%
       dplyr::mutate(date = as.Date(date, format = "%d/%m/%Y")) %>%
-      dplyr::mutate(loc = prov)
+      dplyr::mutate(loc = prov) %>%
+      dplyr::filter(date >= init_date && date <= fin_date)
     
-    temperatures <- rbind(temperatures, tab1, tab2)
+    tab3 <- tab3 %>%
[…]

### 2020-04-02 Documentation/article_4.tex
revisione 1 by Fabio

--- 2020-03-29 Documentation/article_4.tex
@@
 \ref{coronavirus_picture} The origin of this virus is still subject of scientific debate between the scientific community, however one of the most common opinion is that this virus comes from bats, in particular the genus Rhinolophus \cite{zhou2020pneumonia}.
-The most compelling feature of this virus is that is its ability to spread also via coughing and sneezing \cite{ghinai2020first}, and also by touching infected surfaces \cite{chang2020protecting}.
+The most compelling feature of this virus is its ability to spread via coughing and sneezing \cite{ghinai2020first}, and also by touching infected surfaces \cite{chang2020protecting}.
 Differently with respect to the SARS-CoV the virus seems to have a lower mortality rate \cite{sorensen2006severe,weiss2020clinical}.
@@
 In January the Chinese government imposed the quarantine for the city of Wuhan (almost 11M people); the quarantine was later expanded to the full province of Hubei (60M people) and then to the neighbour provinces Huanggang, Ezhou and Xianning.
-The virus then spread in Canada, Germany, Thailand and Japan and then in other different countries included Italy \cite{timeline+web}.
-In Italy the first cases, two Chinese tourist from Hubei, were reported in Rome \cite{corr+roma}; then  other cases were reported in Codogno (Lodi,Lombardy).
+The virus then spread in Canada, Germany, Thailand and Japan and then in other different countries including Italy \cite{timeline+web}.
+In Italy the first cases, two Chinese tourist from Hubei, were reported in Rome \cite{corr+roma}; then  other cases were reported in Codogno (Lodi, Lombardy).
 Later the virus spread almost in all regions of Italy with a higher density in Lombardy, Veneto, Emilia-Romagna, Piemonte and Marche.
-Starting from 22 of February, the Italian government started to impose the quarantine (red-zone) for eleven different municipalities, in particular Codogno, Casalpusterlengo, Lodi (included the neighbour municipalities) and Vo' .
-Starting from this date different restrictive measures were imposted starting from the regions with the highest number of cases: public entrainment were almost suspended as well as schools and universities.
+Starting from 22nd of February, the Italian government started to impose the quarantine (red-zone) for eleven different municipalities, in particular Codogno, Casalpusterlengo, Lodi (included the neighbour municipalities) and Vo\textquotesingle.
+From this date onward, different restrictive measures were imposed starting from regions with the highest number of cases: public entrainment were almost suspended as well as schools and universities.
 Workers (in public and private sectors) were allowed, when possibile, to work at home (smart-working).
-Such measures culminated with a decree approved by the Italian government that divided Italy in three areas: the red zones in which the municipalities with highest number of cases in which all the population was subjected to quarantine, the yellow one (Lombardia, Veneto, Emilia Romagna) in which schools, universities and public events ,sports  as well cinemas were suspended and the rest of Italy in which no restrictive measures were adopted \cite{rep+dec1}; however 3 days later all schools and universities were suspended \cite{guard+01}.
-The restrictive measures were also extend up to all Italy with a further decree \cite{11marzo}.
+Such measures culminated with a decree approved by the Italian government that divided Italy in three areas: the red zones, in which the municipalities with highest number of cases in which all the population was subjected to quarantine; the yellow zones (Lombardia, Veneto, Emilia Romagna) where schools, universities and public events, sports as well cinemas were suspended; the rest of Italy, in which no restrictive measures were adopted \cite{rep+dec1}; however 3 days later all schools and universities were suspended \cite{guard+01}.
+The restrictive measures were also extended up to all Italy with a further decree \cite{11marzo}.
+
+\subsection{Historical Background}
+
+In the last centuries several cases of pandemic diseases were recorded: from our point of view, their dynamic is useful to understand and perhaps to model that further peaks may succeed the first one even with a highest death rate.
+
+\paragraph{The  deadliest pandemics in human history: the Spanish flu}
+
+ The first one, which happens to be the biggest human pandemic (in terms of spread just as much as deaths \cite{potter2001history} ), is the Spanish Flu: this name is quite misleading since it was not originated in Spain, though the Spanish press was the first one to talk about it.
+Indeed, at that time (1918-1920) almost every other country put censorships ........
+.
+The origin of this disease is still under research: some scholars focused on the United States \cite{crosby_2003,barry2004site}, on France \cite{shanks2016no} and on Asia \cite{langford2005did}.
+The disease was ascribed to the virus A/H1N1, a subtype of influenza A Virus.
+The spread of the disease was strongly amplified by the fact that many people lived in very bad hygienic conditions.
+As one can see from Fig.\ref{SpanishFLU} there were basically three waves: the initial spread, the second wave and the third wave.
+The second wave, that was the most lethal one, coincides with the end of the war and with the coming back of the troops: the close contact inside the trains, and the diffusion of the troops inside their home villages/cities increased abruptly the contagion rate and then the death rate.
+Furthermore, this effect was amplified by the fact that a more deadly mutation of the virus became widespread \cite{barry2008cross}.
+As pointed out by different scholars, such effect was enhanced by the fact thst sick people, which infection potential was higher, were transferred by train to the hospitals.
+Immediately, the virus spread with a higher rate \cite{wever2014death}.
+As it is possible to note from Fig.
+\ref{Mortality_spanish_flu} this second wave was largely more deadly for young people with respect the first wave.
+Finally, looking at plot \ref{SpanishFLU}, a third wave also manifested itself in the beginning of 1919: here, the mortality was higher with respect to the first one, but lower to the second one.
+In the end, this pandemic infected 500 million people and killed something between 17 and 50 million \cite{spreeuwenberg2018reassessing}.
+It is worth nothing that, just to compare, the casualties in the Great War were from 8.8 to 10.7 million among soldiers and 11 million among civilian \cite{haythornthwaite1996world}
+
+\begin{figure}[h]
+ \centering
+ \includegraphics[width=1\linewidth]{Figures/SpanishFLU.png} 
+ \caption{Pneumonia and Influenza mortality from 1910-1922 in London: note the three peaks due to the spanish flu.
+The variation of the diffusion rate R$_{0}$ is reported in the box.
+Source: \cite{he2011mechanistic} }
+ \label{SpanishFLU}
+\end{figure}
+
+\begin{figure}[h]
+ \centering
+ \includegraphics[width=0.6\linewidth]{Figures/Mortality_spanish_flu.png} 
+ \caption{Mortality range distribution for the two waves of Spanish flu: note that the second one was more severe for the young people. Source: \cite{taubenberger20061918}}
+ \label{Mortality_spanish_flu}
+\end{figure}
+
+\clearpage
+
+\paragraph{The Asian Influenza (H2N2)}
+
+In 1957 different cases another virus of influenza ( A/H2N2 ) spread in China.
+In this case, at difference with respect to the Spanish flu, the scholars were able to isolate the virus (the virus of the former was only isolated in 1933).
+It was proposed that the virus may be originated from mixing of avian and human influenza (see Fig.
+\ref{Mixing_Flu})
+
+\begin{figure}[h]
+ \centering
+ \includegraphics[width=0.8\linewidth]{Figures/Mixing_Flu.jpg} 
+ \caption{Wang-ShickRyu diagram for the explanantion of the Spanish, Asian flu and Honk-Kong flu . Image taken from \cite{RYU2017195}}
+ \label{Mixing_Flu}
+\end{figure}
 

### 2020-04-02 Documentation/article_4.tex
revisione 2 by Fabio

@@
  \includegraphics[width=0.8\linewidth]{Figures/Mixing_Flu.jpg} 
- \caption{Wang-ShickRyu diagram for the explanantion of the Spanish, Asian flu and Honk-Kong flu . Image taken from \cite{RYU2017195}}
+ \caption{Wang-ShickRyu diagram for the explanantion of the Spanish, Asian flu and Honk-Kong flu . Source: \cite{RYU2017195}}
  \label{Mixing_Flu}
@@
 \paragraph{Motivation} \label{Motivation}
-In order to have a better understanding about the spread and the effects on the population of the COVID19 disease, we propose a ShinyApp, called disCOVIDer19 , that gathers  the data , mostly provided by the Protezione Civile \cite{protezionecivile+git}, about the number of infections(current as well cumulative), hospitalized people as well in intensive care, deaths and recovered for Italy, its regions and provinces; more details will be discussed in sections \ref{Data Origin} and \ref{Panels}.
+In order to have a better understanding about the spread and the effects on the population of the COVID19 disease, we propose a ShinyApp, called disCOVIDer19, that gathers and present the data, chiefly coming from Protezione Civile \cite{protezionecivile+git}, of the numbers of new and cumulative infections, hospitalized and intensive care patients, deaths and recoveries in Italy, its regions and provinces; more details will be discussed in sections \ref{Data Origin} and \ref{Panels}.
 Furthermore we fitted the cumulative cases for Italy and all its regions and provinces with a logistic curve, in order to get a bird-of-eye view on the trend followed as well on the effectiveness of the restrictive measures imposed by the Italian government.

### 2020-04-03 inst/shiny/C19/server/2module/inspection.R
highchart age distribution

@@ -334,3 +334,33 @@ output$tamp_plot <- highcharter::renderHighchart(
                                yAxis = 1, highcharter::hcaes(x = date, y = share_infected_discovered),
                                name="share_infected_discovered", color="#383838")
 )
+
+
+# age distribution plot ---------------------------------------------------
+
+output$age_plot <- highcharter::renderHighchart(
+  highcharter::highchart() %>% 
+    # Data
+    highcharter::hc_add_series(age_df, "column",
+                               highcharter::hcaes(x = age_int, y = cases), name = "cases") %>%
+    highcharter::hc_add_series(age_df, "pie", 
+                               highcharter::hcaes(name = age_int, y = perc_cases), name = "% cases") %>%
+    # Optiosn for each type of series
+    highcharter::hc_plotOptions(
+      series = list(
+        showInLegend = FALSE,
+        pointFormat = "{point.y}%"
+      ),
+      column = list(
+        colorByPoint = TRUE
+      ),
+      pie = list(
+        colorByPoint = TRUE, center = c('30%', '10%'),
+        size = 120, dataLabels = list(enabled = FALSE)
+      )) %>%
+    # Axis
+    highcharter::hc_yAxis(
+      title = list(text = "cases")
+    ) %>%
+    highcharter::hc_xAxis(categories = age_df$age_int)
+)

### 2020-04-03 inst/shiny/C19/ui/tabs/2.tab_inspection.R

@@ -206,6 +206,12 @@ shinydashboard::tabItem(tabName = "tab_2",
                                                      width=12,
                                                      highcharter::highchartOutput("tamp_plot")%>%shinycssloaders::withSpinner( color="#dd4b39")
                                  )
+                          ),
+                          fluidRow(
+                            shinydashboard::box(
+                              title="Age Distribution", status="danger", solidHeader=T,
+                              highcharter::highchartOutput("age_plot")%>%shinycssloaders::withSpinner( color="#dd4b39")
+                            )
                           )
                         )
                         

### 2020-04-03 inst/shiny/C19/global.R
age distribution plots with selector

@@ -240,8 +240,20 @@ age_df <- age_df %>%
   dplyr::group_by(age_int) %>%
   dplyr::summarise(cases=sum(cases, na.rm=T)) %>%
   dplyr::ungroup() %>%
-  dplyr::mutate(perc_cases = round((cases/sum(cases))*100,2))
+  dplyr::mutate(perc_cases = round((cases/sum(cases))*100,2)) %>% 
+  dplyr::mutate(region="--- ALL ---")
 
+age_df_region <- purrr::map_df(names(age_cases), function(x) {
+  age_cases[[x]]$age_cases %>%
+    dplyr::mutate(region=x) %>% 
+    dplyr::filter(!age_int=="Not known") %>% 
+    dplyr::select(-perc_cases) %>%
+    dplyr::ungroup() %>% 
+    dplyr::mutate(perc_cases = round((cases/sum(cases))*100,2))
+})
+
+age_df_final <- age_df %>% 
+  dplyr::bind_rows(age_df_region)
 
 #================================
 

### 2020-04-03 inst/shiny/C19/server/2module/inspection.R

@@ -339,28 +339,29 @@ output$tamp_plot <- highcharter::renderHighchart(
 # age distribution plot ---------------------------------------------------
 
 output$age_plot <- highcharter::renderHighchart(
-  highcharter::highchart() %>% 
-    # Data
-    highcharter::hc_add_series(age_df, "column",
-                               highcharter::hcaes(x = age_int, y = cases), name = "cases") %>%
-    highcharter::hc_add_series(age_df, "pie", 
-                               highcharter::hcaes(name = age_int, y = perc_cases), name = "% cases") %>%
-    # Optiosn for each type of series
-    highcharter::hc_plotOptions(
-      series = list(
-        showInLegend = FALSE,
-        pointFormat = "{point.y}%"
-      ),
-      column = list(
-        colorByPoint = TRUE
-      ),
-      pie = list(
-        colorByPoint = TRUE, center = c('30%', '10%'),
-        size = 120, dataLabels = list(enabled = FALSE)
-      )) %>%
-    # Axis
-    highcharter::hc_yAxis(
-      title = list(text = "cases")
-    ) %>%
-    highcharter::hc_xAxis(categories = age_df$age_int)
-)
+    highcharter::highchart() %>% 
+      # Data
+      highcharter::hc_add_series(dplyr::filter(age_df_final,region==input$regiontab3), "column",
+                                 highcharter::hcaes(x = age_int, y = cases), name = "cases") %>%
+      highcharter::hc_add_series(dplyr::filter(age_df_final,region==input$regiontab3), "pie", 
+                                 highcharter::hcaes(name = age_int, y = perc_cases), name = "% cases") %>%
+      # Optiosn for each type of series
+      highcharter::hc_plotOptions(
+        series = list(
+          showInLegend = FALSE,
+          pointFormat = "{point.y}%"
+        ),
+        column = list(
+          colorByPoint = TRUE
+        ),
+        pie = list(
+          colorByPoint = TRUE, center = c('15%', '20%'),
+          size = 150, dataLabels = list(enabled = FALSE)
+        )) %>%
+      # Axis
+      highcharter::hc_yAxis(
+        title = list(text = "cases")
+      ) %>%
+      highcharter::hc_xAxis(categories = dplyr::filter(age_df_final,region==input$regiontab3)$age_int)
+  )
+

### 2020-04-03 inst/shiny/C19/ui/tabs/2.tab_inspection.R

@@ -70,6 +70,42 @@ shinydashboard::tabItem(tabName = "tab_2",
                                             )
                         ),
                         
+                        # age distribution
+                        shinydashboard::box(
+                          title="Age Distrbution", status = "danger", solidHeader = T,
+                          width = NULL,
+                          tags$head(tags$style(HTML('
+                                                     /* background */
+                                                     
+                                                     .nav-tabs-custom > .nav-tabs > li.active {
+                                                     border-top-color: red !important;
+                                                     }
+                                                     
+                                                     .btn-default {
+                                                     background-color: #dd4b39 !important;
+                                                     color: white !important;
+                                                     border-color: #dd4b39 !important;
+                                                     }
+                                                     '))),
+                          
+                          shiny::fluidRow(
+                            shiny::column(2,
+                                          shiny::selectInput(
+                                            inputId = "countrytab3", label = "Country",
+                                            choices = countryNames, selected = "Italy"),
+                                          
+                                          shiny::selectInput(
+                                            inputId = "regiontab3", label = "Region",
+                                            choices = unique(age_df_final$region), selected = "--- ALL ---")
+                            ),
+                            shiny::column(10,
+                                          highcharter::highchartOutput("age_plot")%>%shinycssloaders::withSpinner( color="#dd4b39")
+                                          
+                            )
+                          )
+                          
+                        ),
+                        
                         
                         
                         br(),
@@ -206,12 +242,6 @@ shinydashboard::tabItem(tabName = "tab_2",
                                                      width=12,
                                                      highcharter::highchartOutput("tamp_plot")%>%shinycssloaders::withSpinner( color="#dd4b39")
                                  )
-                          ),
-                          fluidRow(
-                            shinydashboard::box(
-                              title="Age Distribution", status="danger", solidHeader=T,
-                              highcharter::highchartOutput("age_plot")%>%shinycssloaders::withSpinner( color="#dd4b39")
-                            )
                           )
                         )
                         

### 2020-04-03 inst/shiny/C19/ui/tabs/2.tab_inspection.R
Infected age distribution box title change

@@ -72,7 +72,7 @@ shinydashboard::tabItem(tabName = "tab_2",
                         
                         # age distribution
                         shinydashboard::box(
-                          title="Age Distrbution", status = "danger", solidHeader = T,
+                          title="Infected Age Distribution", status = "danger", solidHeader = T,
                           width = NULL,
                           tags$head(tags$style(HTML('
                                                      /* background */

### 2020-04-03 Documentation/article_4.tex
update
updated motivation section

@@
  The first one, which happens to be the biggest human pandemic (in terms of spread just as much as deaths \cite{potter2001history} ), is the Spanish Flu: this name is quite misleading since it was not originated in Spain, though the Spanish press was the first one to talk about it.
-Indeed, at that time (1918-1920) almost every other country put censorships ........
-.
+Indeed, at that time (1917-1918) almost every other country was involved in the Great War and consequently the censorship was applied to the press .
 The origin of this disease is still under research: some scholars focused on the United States \cite{crosby_2003,barry2004site}, on France \cite{shanks2016no} and on Asia \cite{langford2005did}.
@@
 The variation of the diffusion rate R$_{0}$ is reported in the box.
-Source: \cite{he2011mechanistic} }
+Image source: Ref.
+\cite{he2011mechanistic} }
  \label{SpanishFLU}
@@
  \includegraphics[width=0.6\linewidth]{Figures/Mortality_spanish_flu.png} 
- \caption{Mortality range distribution for the two waves of Spanish flu: note that the second one was more severe for the young people. Source: \cite{taubenberger20061918}}
+ \caption{Mortality range distribution for the two waves of Spanish flu: note that the second one was more severe for the young people. Image source: Ref. \cite{taubenberger20061918}}
  \label{Mortality_spanish_flu}
@@
  \includegraphics[width=0.8\linewidth]{Figures/Mixing_Flu.jpg} 
- \caption{Wang-ShickRyu diagram for the explanantion of the Spanish, Asian flu and Honk-Kong flu . Source: \cite{RYU2017195}}
+ \caption{Wang-ShickRyu diagram for the explanantion of the Spanish, Asian flu and Honk-Kong flu .  Image source: Ref.  \cite{RYU2017195}}
  \label{Mixing_Flu}
@@
 \paragraph{Motivation} \label{Motivation}
-In order to have a better understanding about the spread and the effects on the population of the COVID19 disease, we propose a ShinyApp, called disCOVIDer19, that gathers and present the data, chiefly coming from Protezione Civile \cite{protezionecivile+git}, of the numbers of new and cumulative infections, hospitalized and intensive care patients, deaths and recoveries in Italy, its regions and provinces; more details will be discussed in sections \ref{Data Origin} and \ref{Panels}.
-Furthermore we fitted the cumulative cases for Italy and all its regions and provinces with a logistic curve, in order to get a bird-of-eye view on the trend followed as well on the effectiveness of the restrictive measures imposed by the Italian government.
-The theoretical foundations of this fit ,as well how to manage and use it, will be discussed in the section \ref{Theoretical Background}.
-Beside the logistic curve, we also considered a statistical approach, largely diffused among economists, that considers the number of cases in each day as a time series: on this basis we were able to make use of a particular tool, which will be briefly introduced in the section \ref{Theoretical Background}, that allowed us to  make a further forecast  about future cases.
+In order to have a thorough insight of the spread of the COVID-19 disease and its effects on italian population, we propose a ShinyApp, called disCOVIDer19, that gathers, presents and analyzes the numbers of new and cumulative infections, hospitalized and intensive care patients, deaths and recoveries, over the whole country and over its individual regions and provinces.
+A detailed guide to the functioning of the App is available in sections \ref{Data Origin} and \ref{Panels}.
+As regards the analysis panel, we fitted the cumulative cases with a logistic curve, in order to get a bird-of-eye view on the general and specific trends as well as on the effectiveness of the restrictive measures imposed by the Italian government.
+The theoretical foundations of this fit and how to manage and use it will be discussed in the section \ref{Theoretical Background} and \ref{Data Analysis}.
+Beside the logistic curve, we also considered a statistical approach, largely diffused among economists, that considers the number of total cases in each day as a time series: on this basis we were able to make use of a particular tool, which will be briefly introduced in the section \ref{Theoretical Background}, that allowed us to make a further forecast about future cases.
 This latter becomes useful in case of large deviations from logistic distribution.
+Finally, we must remark that our approach is informational: we do limit ourselves to list and comment our findings without claiming rather delicate conclusions, expecially about predictions, which many people hinge on nowadays and many lives too, unfortunately.
 

### 2020-04-03 Documentation/article_4.tex
updated par. 2.1

@@
 \section{Theoretical Background} \label{Theoretical Background}
-Different models were proposed in the literature for modelling the spread of a disease \cite{keeling2011modeling}, here we are going to consider a relative simple model taken from the growth dynamics of populations.
+Several models were proposed in the literature for modelling the spread of a disease \cite{keeling2011modeling}: here, we are going to consider a relative simple model taken from the growth dynamics of populations.
 
@@
 
-Where $x_{n}$ stands for the population at time and $r$ the growth rate. Thus the following discrete progression will be obtained:
+where $x_{n}$ stands for the population at time and $r$ the growth rate. Renaming $x$ to $P$ and applying recursive substitution, we get:
 
@@
 
-in which $P_{0}$ stands for the initial population. This discrete model can be reshaped in to a continuous one in the following way:
+in which $P_{0}$ represents the initial population. This discrete model can be reshaped in to a continuous one in the following way:
 
@@
 
-this is a relative simple differential equation with separable variable. Its integration gives:
+That's a first order ODE with separable variable. Integrating yields:
 
 \begin{equation}
-P(t)=P_{0}(t)
+P(t)=P_{0}e^{rt}
 \end{equation}
 
-Malthus, of course, did not believed that the population could grow ad infinitum with an exponential growth, but since he estimated that the resources growth follows a linear path, he argued that, at the intersection of the two curves, a consistent part of the population will not have access to the resources.
+Malthus, of course, did not believe that the population could grow ad infinitum with an exponential trend and, since he estimated that the resources growth follows a linear path, he argued that, at the intersection of the two curves, a consistent part of the population will not have access to the resources.
 Thus he expected that the exponential growth realizes only in the first part of the growth.
-Basing on these arguments, a Belgian mathematician Pierre-François Verhulst in 1838, proposed a model \cite{verhulst1838notice} in which the key point is the maximum number of people that the resources allows to live, usually called the carrying capacity here indicated as $K$ ($P$ and $r$ have the same meaning of the previous model):
+Basing on these arguments, in 1838 Belgian mathematician Pierre-François Verhulst developed a model \cite{verhulst1838notice} whose new feature and key point was the maximum size of the population allowed by resources, that is the natural upper limit of coexisting individuals in a particular environment, usually called the \emph{carrying capacity}, here indicated as $K$ ($P$ and $r$ have the same meaning of the previous model):
 
@@
 
-that have the following analytic solution
+that has the following analytic solution:
 
@@
 
-In Fig.
-\ref{Malthusian_growth_vs_logistic_growth} this solution is compared to the Malthusian one: as one can point out the logistic growth has a saturation effect near the asymptote of carrying capacity.
-This represent the limit capacity that Malthus proposed to indicate as the intersection between the linear and exponential growth.
-It is worth nothing that near the origin the two growth are identical, then the logistic curve becomes linear up to the saturation region.
+In Fig.\ref{Malthusian_growth_vs_logistic_growth} this solution is compared to the Malthusian one: as one can see, the logistic growth shows a saturation effect towards the asymptote of carrying capacity.
+This completes Malthus's intuition, as it represents population growth before and after the intersection between the linear resources curve and the natural exponential growth.
+It is worth noting that  in a neighbourhood of the origin the two curves are identical; later, the logistic curve bends down, still increasing, till it reaches the saturation region, in which it becomes almost flat.
 

### 2020-04-03 SEIR.R
Delete SEIR.R

@@ -1,159 +1,0 @@
-#R SEIR
-library(deSolve)
-library(covid19)
-library(ggplot2)
-
-#Data loading & global variables
-lomb <- get_regionTS()$Lombardia
-poplomb <- italy_pop$region[italy_pop$region$territorio == "Lombardia", "valore"]
-P <- lomb$totale_casi
-R <- lomb$totale_casi - lomb$totale_positivi
-
-IT <- c(0,0,0,0,0,1)
-TT <- c(1)
-IBST <- c(1)
-m = max(length(IT), length(TT), length(IBST))
-ITl = length(IT)
-TTl = length(TT)
-IBSTl = length(IBST)
-N = 10500000
-latent_time = IT%*%(1:length(IT)-1)
-
-
-
-#%% Functions
-refine_data <- function(P, R) {
-	
-	n_obs = length(P)
-	
-	S = integer(n_obs-m)
-	E = integer(n_obs-m)
-	I = integer(n_obs-m)
-	R = R[1:(n_obs-m)]
-	
-	for(t in (1:(n_obs-m))){
-		IT_temp = c(IT, integer(n_obs-ITl-t+1))
-		TT_temp = c(TT, integer(n_obs-TTl-t+1))
-		IBST_temp = c(IBST, integer(n_obs-IBSTl-t+1))
-		
-		E[t] <- ((IT_temp+TT_temp)/2) %*% P[t:n_obs]
-		I[t] <- ((IBST_temp+TT_temp)/2) %*% P[t:n_obs]
-	}
-	
-	S = N-E-I-R
-	
-	return(data.frame(S, E, I, R))
-	
-}
-
-SEIR_model <- function(t, state, parameters){
-	with(as.list(c(state, parameters)), {
-		dS <- alpha*R-beta*S*I
-		dE <- beta*S*I-sigma*E
-		dI <- sigma*E-gamma*I
-		dR <- gamma*I-alpha*R
-	
-		list(c(dS, dE, dI, dR))
-	})
-}
-
-solve_SEIR <- function(U0, eval_time=eval_time, parameters=parameters){
-	state <- c( S = U0[1], E=U0[2], I=U0[3], R=U0[4])
-	
-	
-	out <- ode(y=state, times=eval_time, func=SEIR_model, parms=parameters)
-	out <- data.frame(out)
-	return(out)
-}
-
[…]

### 2020-04-03 SEIR_2.py
Delete SEIR_2.py

@@ -1,167 +1,0 @@
-# -*- coding: utf-8 -*-
-"""
-Created on Wed Mar 25 16:07:41 2020
-
-@author: Io
-"""
-
-import numpy as np
-import pandas as pd
-
-from sklearn.linear_model import LinearRegression
-from scipy.integrate import odeint
-from scipy.optimize import curve_fit
-
-IT = np.array([0, 0, 0.05, 0.1, 0.15, 0.3, 0.5]) # Incubation time 
-TT = np.array([1]) # Test time
-IBST = np.array([1]) # IBST(tau) = P(infective at day of symptoms - tau)
-
-
-m = max(len(IT), len(TT), len(IBST))
-N = 60500000
-
-def import_italy():
-    
-    url = 'https://raw.githubusercontent.com/pcm-dpc/COVID-19/master/dati-andamento-nazionale/dpc-covid19-ita-andamento-nazionale.csv'
-    df = pd.read_csv(url)
-    
-    P = df.totale_casi.values   # tested positive cases
-    E = (P-P)[:-m]
-    I = (P-P)[:-m]
-    
-    for t in range(len(P)-m):
-        IT.resize(len(P)-t, refcheck=False)
-        TT.resize(len(P)-t, refcheck=False)
-        IBST.resize(len(P)-t, refcheck=False)
-        
-        E[t] = ((IT+TT)/2).dot(P[t:])
-        I[t] = ((IBST+TT)/2).dot(P[t:])
-    
-    R = df.dimessi_guariti.values + df.deceduti.values
-    R = R[:-m]
-    S = N-E-I-R
-    
-    refined_data = [S, E, I, R]
-    
-    return refined_data, P[:-m]
-    
-def SEIR_model(U, t, param):
-        S, E, I, R = U
-        alpha, beta, gamma, epsilon = param
-        return np.array([
-                -beta*S*I + alpha*R,    #susceptible
-                beta*S*I - epsilon*E,   #exposed
-                epsilon*E - gamma*I,    #infected
-                gamma*I - alpha*R])     #removed
-    
-def der(U, resolution=1):
-    dU = np.empty_like(U)
-    i = 0
-    
-    for X in U:
-        dU[i] = np.array([X[1]-X[0] if t == 0
-            else X[-1]-X[-2] if t == len(X)-1
-            else (X[t+1]-X[t-1])/2 
-            for t in range(len(X))])
-        dU[i] = dU[i]/resolution
-        i += 1
-    
-    
[…]

### 2020-04-03 SEIR_2_test.py
Delete SEIR_2_test.py

@@ -1,99 +1,0 @@
-# -*- coding: utf-8 -*-
-"""
-Created on Wed Mar 25 16:23:10 2020
-
-@author: Io
-"""
-
-from SEIR_2 import *
-import matplotlib.pyplot as plt
-from scipy.interpolate import *
-
-U, P = import_italy()
-
-k=1
-
-t = range(len(P))
-t_k = range(k*len(P))
-
-S, E, I, R = U
-
-# R0 = 2.2
-# alpha = 0
-# beta = R0/N
-# gamma = 0.05
-# epsilon = 1
-alpha, beta, gamma, epsilon = None, None, None, None
-
-
-sol, param, Rsq = solve_SEIR(U, k=k, alpha=alpha, beta=beta, gamma=gamma, epsilon=epsilon)
-
-S_, E_, I_, R_ = sol
-
-plt.plot(t_k, E_, 'r', label='SEIR Exposed')
-plt.plot(t_k, I_, 'g', label='SEIR Infected')
-plt.plot(t_k, R_, 'y', label='SEIR Removed')
-
-plt.plot(t, P, 'bo', label='Positive')    
-plt.plot(t, E, 'ro', label='Exposed')
-plt.plot(t, I, 'go', label='Infected')
-plt.plot(t, R, 'yo', label='Removed')
-plt.legend()
-
-# plt.show()
-
-dict = {0:'alpha', 1:'beta', 2:'gamma', 3:'epsilon', 4:'Rsq'}
-for i in range(len(param)+1):
-    if i<len(param):
-        print(dict.get(i), param[i], sep = '\t:\t')
-    else: 
-        print(dict.get(i), Rsq, sep = '\t:\t')
-        
-#%%    LOGISTIC, SPLINE: data interpolation. more resolution
-        
-# param_l, cov_l = logistic_fit(U)
-# sup, g_rate, cen = param_l
-
-# # dict_l = {0:'Superior', 1:'Growth Rate', 2:'Flex Point'}
-# # for i in range(len(param_l)):
-# #     print(dict_l.get(i), param_l[i], sep = '\t:\t')
-
-# z = [logistic(x, sup, g_rate, cen) for x in t_triple]
-
-# plt.plot(t_triple, z)
-# plt.plot(t, I, 'o')
-
-
-# #%% SEIR WITH DATA FROM LOGISTIC and interpolation
-
[…]

### 2020-04-03 Documentation/article_4.tex
revisioned par. 2.2

@@
 
-The dynamics of the two models described before can be used to shape the spread of an infection \cite{serfling1952historical,ma2014estimating} in this case the population $P$ is replaced with the total number of infected people, $P_{0}$ the number of first infected people, $r$ the spreading rate and the carrying capacity $K$ with the maximum number of people that can be infected.
-The latter is the key point for the modelling: in principle this parameter would be equal to the population number, in practice due to the restrictive measure a large part of population can be removed from this computation: this lowers the predicted number of infected people.
-Thus we can argue that the effectiveness of the restrictive measures can be inferred from the linear behaviour of the cumulative curve (except, of course, if the numer of infected people is close to the population size).
-It is worth nothing that if the quarantine fails for also one/few infected people, the carrying capacity is increased up to the people that can be infected by these subjects.
+The dynamics of the two models described before can be used to shape the spread of an infection \cite{serfling1952historical,ma2014estimating} in this case the population $P$ is replaced with the total number of infected people, $P_{0}$ with the number of initially infected people, $r$ with the spreading rate and the carrying capacity $K$ with the maximum number of people that can be infected.
+The latter is the key point for the modelling: in principle, this parameter would be equal to the population number; in practice, due to restrictive measures and, possibly, vaccines, a large share of the total population won't enter the computation, thus lowering the carrying capacity.
+So, we can argue that the effectiveness of the restrictive measures can be inferred from the behaviour of the cumulative curve (except, of course, if the number of infected people is close to the population size).
+It is worth noting that, for each non-quarantined infected, the carrying capacity is increased up according to the individuals that can be infected by these subjects.
 If this contribution is not negligible a new logistic growth may be found, with of course an exponential starting behaviour.
-Thus the prediction with the logistic curve should be taken as an evaluation of the effectiveness of restrictive measures and as the best scenario that can may realize.
+Thus the prediction with the logistic curve should be taken as an evaluation of the impact of restrictive measures and as the \emph{best scenario} that may occur.
 For this reason it is advisable to see the logistic curve as a local (in time) estimator.
-In principle one can model a differential set of equation were the carrying capacity is time dependent.
+In fact, one can model a differential set of equation were the carrying capacity is time dependent.
 This possibility may be considered  by the authors in a future version of this App.
-Furthermore it is worth nothing that other similar models such as Gopertz, Richards or Bertalanffy can be considered \cite{ma2014estimating} : the authors may consider to include them in  a future version of this App.
+Furthermore, we remind that other similar models such as Gopertz, Richards or Bertalanffy can be used to shape the COVID-19 outbreak\cite{ma2014estimating}: the authors may consider to include them in a future version of this App.
 

### 2020-04-03 SEIR_factotum.R
Add files via upload

@@ -1,0 +1,121 @@
+
+
+SEIR_factotum <- function(P, R, N, gamma=0.1, time_step=1, normalise=TRUE, IT=c(0,0,0,0,0,1), TT=c(1), IBST=c(1)) {
+	#### Functions ####
+	# Extract S, E, I, R from P, R. population is normalised to N=1
+	refine_data <- function(P, R, normalise=TRUE) {
+	
+		n_obs = length(P)
+		m = max(length(IT), length(TT), length(IBST))
+		ITl = length(IT)
+		TTl = length(TT)
+		IBSTl = length(IBST)
+		
+		S = integer(n_obs-m)
+		E = integer(n_obs-m)
+		I = integer(n_obs-m)
+		R = R[1:(n_obs-m)]
+		
+		for(t in (1:(n_obs-m))){
+			IT_temp = c(IT, integer(n_obs-ITl-t+1))
+			TT_temp = c(TT, integer(n_obs-TTl-t+1))
+			IBST_temp = c(IBST, integer(n_obs-IBSTl-t+1))
+			
+			E[t] <- ((IT_temp+TT_temp)/2) %*% P[t:n_obs]
+			I[t] <- ((IBST_temp+TT_temp)/2) %*% P[t:n_obs]
+		}
+		
+		if(normalise){
+			S = (N-E-I-R)/N
+			E = E/N
+			I = I/N
+			R = R/N
+		}
+		
+		return(data.frame(S, E, I, R))
+	
+	}
+	# Estimate R0 using I and its exponential properties
+	estimate_R0 <- function(sigma, gamma, I){
+		# Find the period of exponential growth for X
+		exp_period <- function(X){
+			start <- 1; end <- length(X)
+			period <- c(start, end)
+			return(period)
+		}
+		# Define the exponential approximation of I 
+		I_approx <- function(R0, sigma, gamma, I0, t) {
+	
+			out <- I0*exp(1/2*(sqrt(4*(R0-1)*sigma*gamma+(sigma+gamma)^2)-(sigma+gamma))*t)
+	
+			return(out) 
+	
+		}
+		# Define the target function, ie a least square estimation for log(I)
+		sqdist_log <- function(R0, sigma, gamma, I){
+	
+			out = 0
+			for(t in (1:length(I)))
+				out = out + (log(I_approx(R0, sigma, gamma, I[1],t-1))-log(I[t]))^2
+			
+			return(out)
+		}
[…]

### 2020-04-03 Documentation/article_4.tex
revised par. 4.1 (Home)

@@
 The data shown are synchronised with the civil protection database.
-Whenever the App is started,  a check for the updates is performed and their occurrence is indicated in the \textit{ most recent updates }section.
-The home panel consists of two main sections: the choropleth map and the summary statistics.The choropleth map (Fig \ref{Home_fig1}) is an interactive heat map with breakouts by region and by province tracking the number of Covid-19 cases.
+Whenever the App is started, a check for the updates is performed and their occurrence is indicated in the \textit{most recent updates }section.
+The home panel consists of two main sections: the choropleth map and the summary statistics.
+The choropleth map (Fig.
+\ref{Home_fig1}) is an interactive heat map with breakouts by region and by province tracking the number of Covid-19 cases.
 
@@
 \end{equation}
-This indicator ensures that the number of cases in more populated areas are not over-indexed compared to less populated ones.
-Similarly, it ensures less populated areas are not under-indexed.
-The density  indicator accounts for the territorial extension (in $km^2$) of regions/provinces expressed in parts per thousand (  \textperthousand ) and is calculated as follows:
+This indicator ensures that the number of cases in more populated areas are not over-indexed compared to less populated ones. Similarly, it ensures less populated areas are not under-indexed. 
+The density indicator accounts for the territorial extension (in $km^2$) of regions/provinces expressed in parts per thousand (\textperthousand ) and is calculated as follows:
 \begin{equation}
-density =  \dfrac{total~cases}{territorial~extension}\times 1000
+density~cases =  \dfrac{total~cases}{territorial~extension}\times 1000
 \end{equation}
-This indicator ensures that we are not underestimating small regions/provinces with fewer cases than larger ones but with more concentrated cases per $Km^2$.
+This indicator combines the population density with infection spread, offering a control over space. Indeed, it is directly proportional to population density and to percentage cases:
+\begin{equation}
+density~cases =  \dfrac{total~cases}{local~population} \cdot \dfrac{local~population}{territorial~extension} \times 1000 = percentage~cases \cdot density~population \times 1000
+\end{equation}
+It is enhanced by densely populated territories, which, by the way, may in turn enhance spread infection too.
 Finally, the summary statistics section (Fig.

### 2020-04-03 Documentation/article_4.tex
Update article_4.tex

@@
 The \textit{Data Inspection} panel provides a data visualisation of the general information and a deeper overview about the hospital occupancy, growth monitoring and test tracking for the COVID19 in Italy.
-It is divided in main two section: introduction and deeper inspection.
-In first chart (Fig.
-\ref{fig:Inspection_general_info}) of the introduction are presented the general information about the total cases, total recovered, total hospitalised and intensive care occupancy.
+It is divided in two main section: introduction and deeper inspection.
+In the first chart (Fig.\ref{fig:Inspection_general_info}) of the introduction section, general data of the total cases, total recovered, total hospitalised and intensive care occupancy is presented.
 It is possible to select the entire Italian country or a particular region or province.

### 2020-04-03 Documentation/article_4.tex
pub

--- 2020-04-03 Documentation/article_4.tex
@@
 
-\date{} % An optional date to appear under the author(s)
+\date{\today} % An optional date to appear under the author(s)
 
@@
 
-In 1957 different cases another virus of influenza ( A/H2N2 ) spread in China.
+In 1957 different cases another virus of influenza ( A/H2N2 ) spread in China starting from Guizhou \cite{strahan1994oriental} .
 In this case, at difference with respect to the Spanish flu, the scholars were able to isolate the virus (the virus of the former was only isolated in 1933).
 It was proposed that the virus may be originated from mixing of avian and human influenza (see Fig.
-\ref{Mixing_Flu})
+\ref{Mixing_Flu}).
+The virus diffused in Hong Kong in April then in Singapore, Taiwan and Japan and then globally \cite{saunders2016reviewing} .
+While for the US a crucial role for the spread was supposed to be played by naval  bases \cite{henderson2016development,saunders2016reviewing}, for Europe the same role was played by the  land route trough Asia \cite{langmuir1961epidemiology,saunders2016reviewing}.
+Scholar also pointed out the role of a church conference in Iowa \cite{payne1958some} were almost 1600 people from all countries met.
+Furthermore the open of school in fall increased abruptly the spread of the virus (40-60 \% clinical attack rate) \cite{henderson2009public,saunders2016reviewing}.
+Also in this case it was reported a second wave (or perhaps others one, see Fig.
+\ref{Asiatic_HK_flu}) more lethal than the first one \cite{ebasianflu} .
+In this case, contrarily to the Spanish Flu a vaccine was developed although \cite{henderson2009public}, it was argued that its diffusion was not effective to diminish the effect of the pandemics \cite{jensen1958influenza}.
+It was nothing that according to scholars \cite{trotter1959asian} no effective measures of quarantine or other non-pharmacological one were undertaken by the authorities \cite{trotter1959asian} due to the mild symptoms \cite{saunders2016reviewing}.
+The total toll for this pandemic was from 1 to 1.4 million deaths according to the WHO \cite{who_influenza}: from our point of view this pandemic is a counterfactual teaching lesson about what happens when a pandemics is underestimated (and so no restrictive measure is taken)  due to its relative  mild synthoms
 
@@
  \includegraphics[width=0.8\linewidth]{Figures/Mixing_Flu.jpg} 
- \caption{Wang-ShickRyu diagram for the explanantion of the Spanish, Asian flu and Honk-Kong flu .  Image source: Ref.  \cite{RYU2017195}}
+ \caption{Wang-ShickRyu diagram for the explanation of the Spanish, Asian flu and Honk-Kong flu .  Image source: Ref.  \cite{RYU2017195}}
  \label{Mixing_Flu}
 \end{figure}
+
+\begin{figure}[h]
+ \centering
+ \includegraphics[width=1\linewidth]{Figures/Asiatic_HK_flu.png} 
+ \caption{ Excess mortality due to respiratory disease from 1950 to 1970 as reported by Viboud et al.
+The dashed lines marks the timing of the Hong Kong flue.
+Image source: Ref.
+\cite{viboud2016global}}
+ \label{Asiatic_HK_flu}
+\end{figure}
+
+\clearpage
 
@@
 As regards the analysis panel, we fitted the cumulative cases with a logistic curve, in order to get a bird-of-eye view on the general and specific trends as well as on the effectiveness of the restrictive measures imposed by the Italian government.
-The theoretical foundations of this fit and how to manage and use it will be discussed in the section \ref{Theoretical Background} and \ref{Data Analysis}.
+The theoretical foundations of this fit and how to manage and use it will be discussed in the section \ref{Theoretical Background} and \ref{sub:Data Analysis}.
 Beside the logistic curve, we also considered a statistical approach, largely diffused among economists, that considers the number of total cases in each day as a time series: on this basis we were able to make use of a particular tool, which will be briefly introduced in the section \ref{Theoretical Background}, that allowed us to make a further forecast about future cases.
@@
 
-\subsection{Data Analysis}
+\subsection{Data Analysis}\label{sub:Data Analysis}
 

### 2020-04-03 Documentation/article_4.tex
pub

@@
 \begin{equation}
-density~cases =  \dfrac{total~cases}{local~population} \cdot \dfrac{local~population}{territorial~extension} \times 1000 = percentage~cases \cdot density~population \times 1000
+\begin{split}
+density~cases & =   \dfrac{total~cases}{local~population} \cdot \dfrac{local~population}{territorial~extension} \times 1000 \\  & = percentage~cases \cdot density~population \times 1000
+\end{split}
 \end{equation}

### 2020-04-04 Documentation/article_4.tex
pub

@@
 
-The \textit{Data Inspection} panel provides a data visualisation of the general information and a deeper overview about the hospital occupancy, growth monitoring and test tracking for the COVID19 in Italy.
-It is divided in main two section: introduction and deeper inspection.
-In first chart (Fig.
-\ref{fig:Inspection_general_info}) of the introduction are presented the general information about the total cases, total recovered, total hospitalised and intensive care occupancy.
-It is possible to select the entire Italian country or a particular region or province.
-Moreover, it is possible to visualise and filter the raw data for country, region and province in the panel Raw Data (Fig.
-\ref{fig:Inspection_rawdata}).
-In the deeper inspection panel there are four charts: the first two charts are in two different panels of a box and visualise information about intensive care occupancy in the Italian hospital in different regions, while the others represent the growth monitoring of total cases and test tracking.
-It is worth nothing that the occupation may be higher than their total number, due to the fact that the places in intensive therapy that we considered may be upgraded (see the section \ref{Data Origin}).
-In the first chart \ref{fig:Inspection_perc_occupancy} is represented the percentage hospital occupancy in the selected day divided by capacity with respect to the initial intensive care capacity at the start of the pandemic.
+The \textit{Data Inspection} panel provides an exploratory analysis of the COVID19 spread in Italy and it consists of 2 main sections: an introductory inspection followed by a deeper inspection.
+
+The first section introduces the user to a higher-level time series with the daily frequency of the number of Covid-19 cases broken down into symptomatic, recovered, deaths, and intensive care (Fig.
+\ref{fig:Inspection_general_info} in the \textit{General info} box).
+The cumulative cases add up each day while the daily cases are obtained by taking the first difference of the cumulative series ( $y_t - y_{t-1}$ ).
+When these time series capture all the stages of the pandemic evolution, the daily data are expected to assume a bell-shaped curve with a peak while the cumulative series a shape that is more similar to a logistic curve.
+The user can select different options spanning from the macro-level country view to the more granular province-level.
+In the same \textit{General Info Box}, the second tab provides direct access to the raw data (Fig.
+\ref{fig:Inspection_rawdata}) which is behind most of the visualisation used in the \textit{Home} and \textit{Inspection} sections of the dashboard.
+Through some input selectors and buttons, the dataset can be queried to show the information that is of most interest for the user.
+The \textit{Infected Age Distribution} box
+
+The next chart shows the age distribution of infected patients (Fig.
+\cite{fig:Age_distribution}) with a region selector that offers the possibility to further investigate local-level dynamics and spot regional differences.
+The data was retrieved from Istituto Superiore di Sanità (ISS) \cite{epicentro} which is the National Institute of health in Italy.
+The information dates back to the 30$^{th}$ of March 2020.
+The chart shows the age distribution of the population which is divided into 10 age bins each one representing a 10-year interval.
+In particular, the bar chart indicates the absolute number of infected people in the respective age category while the pie chart shows the percentage breakdown of cases.
+
+The \textit{Deeper Inspection } section consists of 4 different types of visualisations.
+The first two are located in the two tabs of the box called \textit{Intensive Care Information}.
+As the title suggests, these charts provide a comparative visualisation across regions of intensive care occupancy in hospitals.
+While one chart shows the occupancy/capacity index in percentage terms (Fig.
+\ref{fig:Inspection_perc_occupancy}), the other shows absolute figures of capacity and occupancy by region (Fig.
+\ref{fig:Inspection_occupancy2}).
+In particular, the occupancy index is calculated by dividing the percentage occupancy in the selected day by the hospital initial intensive care capacity at the start of the pandemic.
+
 \begin{equation}
@@
 \end{equation}
-The second chart (Fig.
-\ref{fig:Inspection_occupancy2}) visualises a bar chart of the hospital occupancy and capacity of intensive care in different region of Italy at the selected day.The third chart (Fig.
-\ref{fig:Inspection_growth_monitoring}) represents the percentage growth and growth change of total cases day by day.
-The fifth chart (Fig.
-\ref{fig:Inspection_test_tracking}) visualises the daily cases with respect to the daily tests.
+It is worth noting that occupancy may be higher than the total number of available spots due to the fact that the beds in intensive care may have been more recently upgraded (see Sec.
+\ref{Data Origin}).
+The date selector, allows the user to explore the time evolution and variations of such quantities.
+The third chart (Fig.\ref{fig:Inspection_growth_monitoring})  represents the percentage growth and growth.
+
+\begin{equation}
+growth = \frac{cases_t - cases_{t-1}}{cases_t} ~~~~ growth \Delta = growth_t - growth_{t-1}
+\end{equation}
+
+Despite eventual growth in the number of infected people, a persistent negative $\Delta$ is a useful indicator to signal if the pandemic spread is slowing down.
+The fifth chart (\ref{fig:Inspection_test_tracking}) visualises the daily cases with respect to the daily tests.
+It is useful to discern if an increased number of cases is due to an endogenous increase in the number of tests or due to an exogenous increase in cases due to new outbreaks.
 
 \begin{figure}[H]
-\begin{subfigure}{.5\textwidth}
+  \centering
+\begin{subfigure}{.8\textwidth}
   \centering
@@
 \end{subfigure} 
-\begin{subfigure}{.5\textwidth}
+\begin{subfigure}{.8\textwidth}
   \centering
@@
 \begin{figure}[H]
-\begin{subfigure}{0.5\textwidth}
+  \centering
+\begin{subfigure}{0.8\textwidth}
   \centering
@@
 \end{subfigure} 
-\begin{subfigure}{0.5\textwidth}
+\begin{subfigure}{0.8\textwidth}
   \centering
@@
 \end{figure}
+
+\clearpage
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.8\linewidth]{Figures/Age_distribution.jpg} 
+  \caption{The age distribution for the recovered people}
+  \label{fig:Age_distribution}
+\end{figure}\hspace{0.3\textwidth}
 
@@
 \end{subfigure}
-\caption{ }
+\caption{Subpanels of Data Analysis panel }
 \label{fig:logistic_plots_set}
@@
 \end{subfigure}
-\caption{ }
+\caption{Subpanels of Data Analysis panel}
 \label{fig:logistic_plots_set2}

### 2020-04-04 Documentation/article_4.tex
pub

@@
 \section*{Abstract} % This section will not appear in the table of contents due to the star (\section*)
+We developed a user-friendly environment, such a dashboard application,in order to monitor the development of the COVID19 case in Italy.
+We provided a bird of eye view about the total cases for Italy, its Region and Provinces; the data about the symptomatic,recovered and the deaths are also considered.
+Furthermore we provided a detailed overview about the intensive care capacity for each region as well as the overall trend of the test.
+In addition the data are fitted with the logistic curve and with a ARIMA model (in this case the log of the cumulative number is considered)
 
@@
 
-The dynamics of the two models described before can be used to shape the spread of an infection \cite{serfling1952historical,ma2014estimating} in this case the population $P$ is replaced with the total number of infected people, $P_{0}$ with the number of initially infected people, $r$ with the spreading rate and the carrying capacity $K$ with the maximum number of people that can be infected.
+The dynamics of the model described before can be used to shape the spread of an infection  \cite{serfling1952historical,ma2014estimating}.
+In this case the population $P$ is replaced with the total number of infected people, $P_{0}$ with the number of initially infected people, $r$ with the spreading rate and the carrying capacity $K$ with the maximum number of people that can be infected.
 The latter is the key point for the modelling: in principle, this parameter would be equal to the population number; in practice, due to restrictive measures and, possibly, vaccines, a large share of the total population won't enter the computation, thus lowering the carrying capacity.
@@
 The next chart shows the age distribution of infected patients (Fig.
-\cite{fig:Age_distribution}) with a region selector that offers the possibility to further investigate local-level dynamics and spot regional differences.
+\ref{fig:Age_distribution}) with a region selector that offers the possibility to further investigate local-level dynamics and spot regional differences.
 The data was retrieved from Istituto Superiore di Sanità (ISS) \cite{epicentro} which is the National Institute of health in Italy.
@@
 \begin{equation}
-growth = \frac{cases_t - cases_{t-1}}{cases_t} ~~~~ growth \Delta = growth_t - growth_{t-1}
+growth = \frac{cases_t - cases_{t-1}}{cases_t} ~~~~~ \Delta growth  = growth_t - growth_{t-1}
 \end{equation}
@@
 \end{subfigure} 
-\caption{•}
+\caption{Subpanels of Data Inspection panel }
 \end{figure}
@@
 \end{subfigure}\hspace{0.3\textwidth}
-\caption{ }
 \label{fig:Inspection_panels_2}
+\caption{Subpanels of Data Inspection panel }
 \end{figure}
@@
 
-Panel \textit{Analysis} explores different mathematical models for shaping the infection\textquotesingle s time series, namely the ones introduced in Section \ref{Theoretical Background}.
+Panel \textit{Analysis} explores  several mathematical models for shaping the infection\textquotesingle s time series, namely the ones introduced in Section \ref{Theoretical Background}.
 As previously argued, each of these model\textquotesingle s predictive power hinges on uncertain actual scenarios, such as the effectiveness of restrictive measures, the presence of several independent and spread outbreaks, and so on.
@@
   \includegraphics[width=1\linewidth]{Figures/logistic_input.png} 
-  \caption{Set of useful tools for navigating into plots, by plotly.}
+  \caption{Input box of the logistic curve section.}
   \label{fig:logistic_input}
@@
   \includegraphics[width=1\linewidth]{Figures/logistic_plot1.png} 
-  \caption{Summary output box of logistic section.}
+  \caption{ Plot of fitted logistic curve over sample cumulative cases of infection, plus a 95 \% confidence band.
+Green points are the observed data used for fitting, red ones are the remaining.
+Hovering over any point or curve displays info such as x and y coordinates.}
   \label{fig:logistic_plot1}
@@
   \includegraphics[width=1\linewidth]{Figures/logistic_plot2.png} 
-  \caption{One plot of nls residuals: square root of absolute residuals vs. fitted values.}
+  \caption{Plot of estimated logistic distribution over sample new cases of infection.}
   \label{fig:logistic_plot2}
 \end{subfigure}
+\label{fig:logistic_plots_set}
 \caption{Subpanels of Data Analysis panel }
-\label{fig:logistic_plots_set}
 \end{figure}
@@
   \includegraphics[width=1\linewidth]{Figures/logistic_plot3.png} 
-  \caption{Set of useful tools for navigating into plots, by plotly.}
+  \caption{SSet of useful tools for navigating into plots, by plotly.}
   \label{fig:logistic_plot3}
@@
   \includegraphics[width=1\linewidth]{Figures/logistic_res1.png} 
-  \caption{Summary output box of logistic section.}
+  \caption{One plot of nls residuals: square root of absolute residuals vs. fitted values.}
   \label{fig:logistic_res1}
@@
   \includegraphics[width=1\linewidth]{Figures/logistic_smry1.png} 
-  \caption{One plot of nls residuals: square root of absolute residuals vs. fitted values.}
+  \caption{Summary output box of logistic section.}
   \label{fig:logistic_smry1}
 \end{subfigure}
-\caption{Subpanels of Data Analysis panel}
 \label{fig:logistic_plots_set2}
+\caption{Subpanels of Data Analysis panel }
 \end{figure}

### 2020-04-04 inst/shiny/C19/server/4module/documentation.R
Update documentation.R

@@ -1,9 +1,9 @@
 output$downloadDocument <- downloadHandler(
-  filename <- "documentation_29032020.pdf",
+  filename <- "disCOVIDer19_documentation.pdf",
   
   
   content <- function(file) {
-    file.copy("www/documentation_29032020.pdf", file)
+    file.copy(paste0("www/",filename), file)
   },
 )
 
@@ -14,5 +14,5 @@ output$downloadDocument <- downloadHandler(
 tags$iframe(style="height:600px; width:100%", src="")
 
 output$pdfview <- renderText({
-  return(paste('<iframe style="height:600px; width:100%" src="documentation_29032020.pdf"></iframe>', sep = ""))
+  return(paste('<iframe style="height:600px; width:100%" src="',filename,'"></iframe>', sep = ""))
 })

### 2020-04-04 inst/shiny/C19/ui/tabs/4.tab_conclusion.R
Update 4.tab_conclusion.R

@@ -124,6 +124,7 @@ shinydashboard::tabItem(tabName = "tab_4",
                                                             <li><h4>  <a href="http://dati.istat.it/Index.aspx?DataSetCode=DCIS_POPRES1">  Istat - Population  </a> </h4> </li>                                                                                                                           
                                                             <li><h4>  <a href="https://www.istat.it/it/files//2015/04/Superfici-delle-unit%C3%A0-amministrative-Dati-comunali-e-provinciali.zip">  Istat - Territory </a> </h4> </li>
                                                             <li><h4>  <a href="https://www.corriere.it/cronache/20_marzo_16/coronavirus-quanti-posti-terapia-intensiva-ci-sono-italia-quanti-ne-arriveranno-0fbafa76-678a-11ea-93a4-da8ab3a8afb1.shtml" rel="nofollow noreferrer">   Corriere.it - Hospital intensive care capacity </a>  </h4></li>
+                                                            <li><h4>  <a href="https://www.epicentro.iss.it"> Istituto superiore di Sanita\' - Age distribution report </a> </h4> </li>
 
                                                             </ul>
                                                                  ')

### 2020-09-16 inst/shiny/C19/global.R
added weekly aggregation to tests tracking

@@ -9,7 +9,7 @@ intensivecare_capacity = covid19:::get_intensivecare_cap(regionTS)
 age_cases = covid19:::get_agecases(as.character(sort(names(regionTS))))
 decrees = covid19:::get_decrees()
 
-N <- nrow(countryTS)
+N <- nrow(countryTS$Italy)
 
 
 #===  Global function to check error 
@@ -42,6 +42,19 @@ toLog10<- function(num)
 UTSdate <- function(date) {
   return(as.integer(date) * 86400000)
 }
+
+
+# Aggregates a time series weekly
+aggr_wly <- function(v,avg=F) {
+  out <- unlist( lapply( split(v,ceiling(seq_along(v)/7)) ,sum) )
+  if(avg)
+    out <- out/7
+  return(unname(out))
+}
+
+init_date <- min(countryTS$Italy$data)
+
+fin_date <- max(countryTS$Italy$data)
 
 #================================
 
@@ -272,10 +285,24 @@ tamp_country <- tibble::tibble(
   region = "--- ALL ---"
 )
 
+tamp_country_wly <- tibble::tibble(
+  data=countryTS$Italy$data[seq(1,N,7)],
+  tamponi_giornalieri=aggr_wly(diff(c(0,countryTS$Italy$tamponi))),
+  casi_giornalieri=aggr_wly(diff(c(0,countryTS$Italy$totale_casi))),
+  region = "--- ALL ---"
+)
+
 tamp_region <- purrr::map_df(names(regionTS), function(x){
   regionTS[[x]] %>%
     dplyr::select(data,tamponi,totale_casi) %>%
     dplyr::mutate(region=x)
+})
+
+tamp_region_wly <- purrr::map_df(names(regionTS), function(x){
+  data.frame("data"=regionTS[[x]]$data[seq(1,N,7)],
+             "tamponi_giornalieri"=aggr_wly(diff(c(0,regionTS[[x]]$tamponi))),
+             "casi_giornalieri"=aggr_wly(diff(c(0,regionTS[[x]]$totale_casi))),
+             "region"=x)
 })
 
 tamp_creg <- tamp_country %>% 
@@ -283,15 +310,27 @@ tamp_creg <- tamp_country %>%
   dplyr::ungroup() %>%
   dplyr::group_by(region) %>%
   dplyr::mutate(casi_giornalieri=totale_casi-dplyr::lag(totale_casi)) %>% 
[…]

### 2020-09-16 inst/shiny/C19/server/2module/inspection.R

@@ -2,6 +2,7 @@
 #Dataset and plot reactive
 reac_dataset <- shiny::reactiveValues()
 reac_delay <- shiny::reactiveValues()
+reac_test <- shiny::reactiveValues()
 
 disc_NAfind <- function(v) {
   n <- length(v)
@@ -476,11 +477,23 @@ highcharter::highchart(type = "stock") %>%
 
 )
 
-# tamponi graph -----------------------------------------------------------
+# Tests tracking -----------------------------------------------------------
+
+shiny::observe({
+  if(input$test_aggr) {
+    reac_test$tamp_creg <- tamp_creg_wly
+    reac_test$tamp_creg_1 <- tamp_creg_1_wly
+  } else {
+    reac_test$tamp_creg <- tamp_creg
+    reac_test$tamp_creg_1 <- tamp_creg_1
+  }
+  
+})
 
 
 output$tamp_plot <- highcharter::renderHighchart(
-  highcharter::hchart(dplyr::filter(tamp_creg_1,region==input$test_region), "column", highcharter::hcaes(x = date, y = value, group = key), color=c("red","#888888")) %>% 
+  highcharter::hchart(dplyr::filter(reac_test$tamp_creg_1,region==input$test_region), "column", 
+                      highcharter::hcaes(x = date, y = value, group = key), color=c("red","#888888")) %>% 
     # BUGGED
     #highcharter::hc_chart(zoomType = "xy", scrollablePlotArea = list(minWidth = 1000, scrollPositionX = 1)) %>%
     highcharter::hc_chart(zoomType = "xy") %>%
@@ -488,8 +501,8 @@ output$tamp_plot <- highcharter::renderHighchart(
       list(lineWidth = 3, title = list(text  =  '')),
       list(showLastLabel = FALSE, opposite = TRUE, title = list(text  =  ''))
     ) %>%
-    highcharter::hc_add_series(data = dplyr::filter(tamp_creg,region==input$test_region), type = "spline", 
-                               yAxis = 1, highcharter::hcaes(x = date, y = share_infected_discovered),
+    highcharter::hc_add_series(data = dplyr::filter(reac_test$tamp_creg,region==input$test_region), 
+                               type = "spline", yAxis = 1, highcharter::hcaes(x = date, y = share_infected_discovered),
                                name="share_infected_discovered", color="#383838")
 )
 
@@ -589,7 +602,7 @@ output$growth_NAlog <- renderUI({
 
[…]

### 2020-09-16 inst/shiny/C19/ui/tabs/2.tab_inspection.R

@@ -261,6 +261,10 @@ shinydashboard::tabItem(tabName = "tab_2",
                                                                 inputId = "test_region", label = "Region",
                                                                 choices = c(names(regionTS), "--- ALL ---"), selected = "--- ALL ---"),
                                                               
+                                                              shiny::checkboxInput(
+                                                                inputId = "test_aggr", label = "Aggregate by week",
+                                                                value = FALSE
+                                                              ),
                                                               shiny::uiOutput("test_NAlog")
                                                               ),
                                                        column(10,

### 2020-09-16 Documentation/article_4.tex
original doc

--- 2020-04-03 Documentation/article_4.tex
@@
 
-\date{\today} % An optional date to appear under the author(s)
+\date{} % An optional date to appear under the author(s)
 
@@
 \ref{coronavirus_picture} The origin of this virus is still subject of scientific debate between the scientific community, however one of the most common opinion is that this virus comes from bats, in particular the genus Rhinolophus \cite{zhou2020pneumonia}.
-The most compelling feature of this virus is its ability to spread via coughing and sneezing \cite{ghinai2020first}, and also by touching infected surfaces \cite{chang2020protecting}.
+The most compelling feature of this virus is that is its ability to spread also via coughing and sneezing \cite{ghinai2020first}, and also by touching infected surfaces \cite{chang2020protecting}.
 Differently with respect to the SARS-CoV the virus seems to have a lower mortality rate \cite{sorensen2006severe,weiss2020clinical}.
@@
 In January the Chinese government imposed the quarantine for the city of Wuhan (almost 11M people); the quarantine was later expanded to the full province of Hubei (60M people) and then to the neighbour provinces Huanggang, Ezhou and Xianning.
-The virus then spread in Canada, Germany, Thailand and Japan and then in other different countries including Italy \cite{timeline+web}.
-In Italy the first cases, two Chinese tourist from Hubei, were reported in Rome \cite{corr+roma}; then  other cases were reported in Codogno (Lodi, Lombardy).
+The virus then spread in Canada, Germany, Thailand and Japan and then in other different countries included Italy \cite{timeline+web}.
+In Italy the first cases, two Chinese tourist from Hubei, were reported in Rome \cite{corr+roma}; then  other cases were reported in Codogno (Lodi,Lombardy).
 Later the virus spread almost in all regions of Italy with a higher density in Lombardy, Veneto, Emilia-Romagna, Piemonte and Marche.
-Starting from 22nd of February, the Italian government started to impose the quarantine (red-zone) for eleven different municipalities, in particular Codogno, Casalpusterlengo, Lodi (included the neighbour municipalities) and Vo\textquotesingle.
-From this date onward, different restrictive measures were imposed starting from regions with the highest number of cases: public entrainment were almost suspended as well as schools and universities.
+Starting from 22 of February, the Italian government started to impose the quarantine (red-zone) for eleven different municipalities, in particular Codogno, Casalpusterlengo, Lodi (included the neighbour municipalities) and Vo' .
+Starting from this date different restrictive measures were imposted starting from the regions with the highest number of cases: public entrainment were almost suspended as well as schools and universities.
 Workers (in public and private sectors) were allowed, when possibile, to work at home (smart-working).
-Such measures culminated with a decree approved by the Italian government that divided Italy in three areas: the red zones, in which the municipalities with highest number of cases in which all the population was subjected to quarantine; the yellow zones (Lombardia, Veneto, Emilia Romagna) where schools, universities and public events, sports as well cinemas were suspended; the rest of Italy, in which no restrictive measures were adopted \cite{rep+dec1}; however 3 days later all schools and universities were suspended \cite{guard+01}.
-The restrictive measures were also extended up to all Italy with a further decree \cite{11marzo}.
+Such measures culminated with a decree approved by the Italian government that divided Italy in three areas: the red zones in which the municipalities with highest number of cases in which all the population was subjected to quarantine, the yellow one (Lombardia, Veneto, Emilia Romagna) in which schools, universities and public events ,sports  as well cinemas were suspended and the rest of Italy in which no restrictive measures were adopted \cite{rep+dec1}; however 3 days later all schools and universities were suspended \cite{guard+01}.
+The restrictive measures were also extend up to all Italy with a further decree \cite{11marzo}.
 
@@
 
-In the last centuries several cases of pandemic diseases were recorded: from our point of view, their dynamic is useful to understand and perhaps to model that further peaks may succeed the first one even with a highest death rate.
+In the last centuries different cases of pandemic diseases were recorded: from our point of view the dynamics of them it is useful to understand, ad perhaps to model,  that further peaks may succeed the first one also with a highest death rate.
 
@@
 
- The first one, which happens to be the biggest human pandemic (in terms of spread just as much as deaths \cite{potter2001history} ), is the Spanish Flu: this name is quite misleading since it was not originated in Spain, though the Spanish press was the first one to talk about it.
-Indeed, at that time (1917-1918) almost every other country was involved in the Great War and consequently the censorship was applied to the press .
-The origin of this disease is still under research: some scholars focused on the United States \cite{crosby_2003,barry2004site}, on France \cite{shanks2016no} and on Asia \cite{langford2005did}.
-The disease was ascribed to the virus A/H1N1, a subtype of influenza A Virus.
-The spread of the disease was strongly amplified by the fact that many people lived in very bad hygienic conditions.
-As one can see from Fig.\ref{SpanishFLU} there were basically three waves: the initial spread, the second wave and the third wave.
-The second wave, that was the most lethal one, coincides with the end of the war and with the coming back of the troops: the close contact inside the trains, and the diffusion of the troops inside their home villages/cities increased abruptly the contagion rate and then the death rate.
-Furthermore, this effect was amplified by the fact that a more deadly mutation of the virus became widespread \cite{barry2008cross}.
-As pointed out by different scholars, such effect was enhanced by the fact thst sick people, which infection potential was higher, were transferred by train to the hospitals.
-Immediately, the virus spread with a higher rate \cite{wever2014death}.
-As it is possible to note from Fig.
+ The first one, the biggest one known for humans (in terms of spread as well for deaths \cite{potter2001history} ), is the Spanish Flu: this name is quite misleading since it was not originated in Spain, but the Spanish newspaper were the first one to talk about because large part of the involved countries were involved in the Great War (1914-1918).
+The origin of this disease is still under research: some scholars focused on the  United States \cite{crosby_2003,barry2004site}on France \cite{shanks2016no} and on Asia \cite{langford2005did}.
+The disease was ascribed to the virus A/H1N1 a subtype of influenza A Virus.
+The spread of the disease was strongly amplified by the fact that many people lived very close in to a very difficult hygienic conditions.
+As one can see from Fig.
+\ref{SpanishFLU} there are basically three waves: the initial spread, the second wave and the third wave.
+The second wave, that is the most lethal one, coincides with the end of the war and with the coming back of the troops: the close contact inside the trains, and the diffusion of the troops inside their home villages/cities increased abruptly the contagion rate and then the death rate.
+Furthermore this effect was amplified by the fact that a more deadly mutation of the virus diffused  \cite{barry2008cross}.
+As pointed out by different scholars such effect was enhanced since, the infected people, for which the severity of the ill was higher were transferred by train  into the hospitals and so  the virus spread with a higher rate \cite{wever2014death}.As it is possible to note from Fig.
 \ref{Mortality_spanish_flu} this second wave was largely more deadly for young people with respect the first wave.
-Finally, looking at plot \ref{SpanishFLU}, a third wave also manifested itself in the beginning of 1919: here, the mortality was higher with respect to the first one, but lower to the second one.
-In the end, this pandemic infected 500 million people and killed something between 17 and 50 million \cite{spreeuwenberg2018reassessing}.
-It is worth nothing that, just to compare, the casualties in the Great War were from 8.8 to 10.7 million among soldiers and 11 million among civilian \cite{haythornthwaite1996world}
+Finally, as can be noted in the plot \ref{SpanishFLU}, a third wave also manifested in the beginning of 1919: in this case the mortality was higher with respect to the first one but lower to the second one.
+In the end this pandemic infected 500 million people and a large part of them , from 17 to 50 million, died \cite{spreeuwenberg2018reassessing}.
+It is worth nothing that, just to give a comparison, the causalities in the Great War were from 8.8 to 10.7 million among soldiers and 11 million among civilian \cite{haythornthwaite1996world}
 
@@
  \includegraphics[width=1\linewidth]{Figures/SpanishFLU.png} 
- \caption{Pneumonia and Influenza mortality from 1910-1922 in London: note the three peaks due to the spanish flu.
-The variation of the diffusion rate R$_{0}$ is reported in the box.
-Image source: Ref.
-\cite{he2011mechanistic} }
+ \caption{Pneumonia and Influenza mortality from 1910-1922 in London: note the tree peaks due to the spanish flu.
+In the in box the variation of the diffusion rate R$_{0}$ is reported.
+Image taken from \cite{he2011mechanistic} }
  \label{SpanishFLU}
@@
  \includegraphics[width=0.6\linewidth]{Figures/Mortality_spanish_flu.png} 
- \caption{Mortality range distribution for the two waves of Spanish flu: note that the second one was more severe for the young people. Image source: Ref. \cite{taubenberger20061918}}
+ \caption{Mortality range distribution for the two waves of Spanish flu: note that the second one was more severe for the young people. Image taken from \cite{taubenberger20061918}}
  \label{Mortality_spanish_flu}
@@
 
-In 1957 different cases another virus of influenza ( A/H2N2 ) spread in China starting from Guizhou \cite{strahan1994oriental} .
+In 1957 different cases another virus of influenza ( A/H2N2 ) spread in China.
 In this case, at difference with respect to the Spanish flu, the scholars were able to isolate the virus (the virus of the former was only isolated in 1933).
 It was proposed that the virus may be originated from mixing of avian and human influenza (see Fig.
-\ref{Mixing_Flu}).
-The virus diffused in Hong Kong in April then in Singapore, Taiwan and Japan and then globally \cite{saunders2016reviewing} .
-While for the US a crucial role for the spread was supposed to be played by naval  bases \cite{henderson2016development,saunders2016reviewing}, for Europe the same role was played by the  land route trough Asia \cite{langmuir1961epidemiology,saunders2016reviewing}.
-Scholar also pointed out the role of a church conference in Iowa \cite{payne1958some} were almost 1600 people from all countries met.
-Furthermore the open of school in fall increased abruptly the spread of the virus (40-60 \% clinical attack rate) \cite{henderson2009public,saunders2016reviewing}.
-Also in this case it was reported a second wave (or perhaps others one, see Fig.
-\ref{Asiatic_HK_flu}) more lethal than the first one \cite{ebasianflu} .
-In this case, contrarily to the Spanish Flu a vaccine was developed although \cite{henderson2009public}, it was argued that its diffusion was not effective to diminish the effect of the pandemics \cite{jensen1958influenza}.
-It was nothing that according to scholars \cite{trotter1959asian} no effective measures of quarantine or other non-pharmacological one were undertaken by the authorities \cite{trotter1959asian} due to the mild symptoms \cite{saunders2016reviewing}.
-The total toll for this pandemic was from 1 to 1.4 million deaths according to the WHO \cite{who_influenza}: from our point of view this pandemic is a counterfactual teaching lesson about what happens when a pandemics is underestimated (and so no restrictive measure is taken)  due to its relative  mild synthoms
+\ref{Mixing_Flu})
 
@@
  \includegraphics[width=0.8\linewidth]{Figures/Mixing_Flu.jpg} 
- \caption{Wang-ShickRyu diagram for the explanation of the Spanish, Asian flu and Honk-Kong flu .  Image source: Ref.  \cite{RYU2017195}}
+ \caption{Wang-ShickRyu diagram for the explanantion of the Spanish, Asian flu and Honk-Kong flu . Image taken from \cite{RYU2017195}}
  \label{Mixing_Flu}
@@
 
-\begin{figure}[h]
- \centering
- \includegraphics[width=1\linewidth]{Figures/Asiatic_HK_flu.png} 
- \caption{ Excess mortality due to respiratory disease from 1950 to 1970 as reported by Viboud et al.
-The dashed lines marks the timing of the Hong Kong flue.
-Image source: Ref.
-\cite{viboud2016global}}
- \label{Asiatic_HK_flu}
-\end{figure}
-
-\clearpage
-
 \paragraph{Motivation} \label{Motivation}
-In order to have a thorough insight of the spread of the COVID-19 disease and its effects on italian population, we propose a ShinyApp, called disCOVIDer19, that gathers, presents and analyzes the numbers of new and cumulative infections, hospitalized and intensive care patients, deaths and recoveries, over the whole country and over its individual regions and provinces.
-A detailed guide to the functioning of the App is available in sections \ref{Data Origin} and \ref{Panels}.
-As regards the analysis panel, we fitted the cumulative cases with a logistic curve, in order to get a bird-of-eye view on the general and specific trends as well as on the effectiveness of the restrictive measures imposed by the Italian government.
-The theoretical foundations of this fit and how to manage and use it will be discussed in the section \ref{Theoretical Background} and \ref{sub:Data Analysis}.
-Beside the logistic curve, we also considered a statistical approach, largely diffused among economists, that considers the number of total cases in each day as a time series: on this basis we were able to make use of a particular tool, which will be briefly introduced in the section \ref{Theoretical Background}, that allowed us to make a further forecast about future cases.
+In order to have a better understanding about the spread and the effects on the population of the COVID19 disease, we propose a ShinyApp, called disCOVIDer19 , that gathers  the data , mostly provided by the Protezione Civile \cite{protezionecivile+git}, about the number of infections(current as well cumulative), hospitalized people as well in intensive care, deaths and recovered for Italy, its regions and provinces; more details will be discussed in sections \ref{Data Origin} and \ref{Panels}.
+Furthermore we fitted the cumulative cases for Italy and all its regions and provinces with a logistic curve, in order to get a bird-of-eye view on the trend followed as well on the effectiveness of the restrictive measures imposed by the Italian government.
+The theoretical foundations of this fit ,as well how to manage and use it, will be discussed in the section \ref{Theoretical Background}.
+Beside the logistic curve, we also considered a statistical approach, largely diffused among economists, that considers the number of cases in each day as a time series: on this basis we were able to make use of a particular tool, which will be briefly introduced in the section \ref{Theoretical Background}, that allowed us to  make a further forecast  about future cases.
 This latter becomes useful in case of large deviations from logistic distribution.
-Finally, we must remark that our approach is informational: we do limit ourselves to list and comment our findings without claiming rather delicate conclusions, expecially about predictions, which many people hinge on nowadays and many lives too, unfortunately.
 
@@
 \section{Theoretical Background} \label{Theoretical Background}
-Several models were proposed in the literature for modelling the spread of a disease \cite{keeling2011modeling}: here, we are going to consider a relative simple model taken from the growth dynamics of populations.
+Different models were proposed in the literature for modelling the spread of a disease \cite{keeling2011modeling}, here we are going to consider a relative simple model taken from the growth dynamics of populations.
 
@@
 
-where $x_{n}$ stands for the population at time and $r$ the growth rate. Renaming $x$ to $P$ and applying recursive substitution, we get:
+Where $x_{n}$ stands for the population at time and $r$ the growth rate. Thus the following discrete progression will be obtained:
 
@@
 
-in which $P_{0}$ represents the initial population. This discrete model can be reshaped in to a continuous one in the following way:
+in which $P_{0}$ stands for the initial population. This discrete model can be reshaped in to a continuous one in the following way:
 
@@
 
-That's a first order ODE with separable variable. Integrating yields:
+this is a relative simple differential equation with separable variable. Its integration gives:
 
 \begin{equation}
-P(t)=P_{0}e^{rt}
+P(t)=P_{0}(t)
 \end{equation}
 
-Malthus, of course, did not believe that the population could grow ad infinitum with an exponential trend and, since he estimated that the resources growth follows a linear path, he argued that, at the intersection of the two curves, a consistent part of the population will not have access to the resources.
+Malthus, of course, did not believed that the population could grow ad infinitum with an exponential growth, but since he estimated that the resources growth follows a linear path, he argued that, at the intersection of the two curves, a consistent part of the population will not have access to the resources.
 Thus  he expected that the exponential growth realizes only in the first part of the growth.
-Basing on these arguments, in 1838 Belgian mathematician Pierre-François Verhulst developed a model \cite{verhulst1838notice} whose new feature and key point was the maximum size of the population allowed by resources, that is the natural upper limit of coexisting individuals in a particular environment, usually called the \emph{carrying capacity}, here indicated as $K$ ($P$ and $r$ have the same meaning of the previous model):
+Basing on these arguments, a Belgian mathematician Pierre-François Verhulst in 1838, proposed a model \cite{verhulst1838notice} in which the key point is the maximum number of people that the resources allows to live, usually called the carrying capacity here indicated as $K$ ($P$ and $r$ have the same meaning of the previous model):
 
@@
 
-that has the following analytic solution:
+that have the following analytic solution
 
@@
 
-In Fig.\ref{Malthusian_growth_vs_logistic_growth} this solution is compared to the Malthusian one: as one can see, the logistic growth shows a saturation effect towards the asymptote of carrying capacity.
-This completes Malthus's intuition, as it represents population growth before and after the intersection between the linear resources curve and the natural exponential growth.
-It is worth noting that  in a neighbourhood of the origin the two curves are identical; later, the logistic curve bends down, still increasing, till it reaches the saturation region, in which it becomes almost flat.
+In Fig.
+\ref{Malthusian_growth_vs_logistic_growth} this solution is compared to the Malthusian one: as one can point out the logistic growth has a saturation effect near the asymptote of carrying capacity.
+This represent the limit capacity that Malthus proposed to indicate as the intersection between the linear and exponential growth.
+It is worth nothing that near the origin the two growth are identical, then the logistic curve becomes linear up to the saturation region.
 
@@
 
-The dynamics of the two models described before can be used to shape the spread of an infection \cite{serfling1952historical,ma2014estimating} in this case the population $P$ is replaced with the total number of infected people, $P_{0}$ with the number of initially infected people, $r$ with the spreading rate and the carrying capacity $K$ with the maximum number of people that can be infected.
-The latter is the key point for the modelling: in principle, this parameter would be equal to the population number; in practice, due to restrictive measures and, possibly, vaccines, a large share of the total population won't enter the computation, thus lowering the carrying capacity.
-So, we can argue that the effectiveness of the restrictive measures can be inferred from the behaviour of the cumulative curve (except, of course, if the number of infected people is close to the population size).
-It is worth noting that, for each non-quarantined infected, the carrying capacity is increased up according to the individuals that can be infected by these subjects.
+The dynamics of the two models described before can be used to shape the spread of an infection \cite{serfling1952historical,ma2014estimating} in this case the population $P$ is replaced with the total number of infected people, $P_{0}$ the number of first infected people, $r$ the spreading rate and the carrying capacity $K$ with the maximum number of people that can be infected.
+The latter is the key point for the modelling: in principle this parameter would be equal to the population number, in practice due to the restrictive measure a large part of population can be removed from this computation: this lowers the predicted number of infected people.
+Thus we can argue that the effectiveness of the restrictive measures can be inferred from the linear behaviour of the cumulative curve (except, of course, if the numer of infected people is close to the population size).
+It is worth nothing that if the quarantine fails for also one/few infected people, the carrying capacity is increased up to the people that can be infected by these subjects.
 If this contribution is not negligible a new logistic growth may be found, with of course an exponential starting behaviour.
-Thus the prediction with the logistic curve should be taken as an evaluation of the impact of restrictive measures and as the \emph{best scenario} that may occur.
+Thus the prediction with the logistic curve should be taken as an evaluation of the effectiveness of restrictive measures and as the best scenario that can may realize.
 For this reason it is advisable to see the logistic curve as a local (in time) estimator.
-In fact, one can model a differential set of equation were the carrying capacity is time dependent.
+In principle one can model a differential set of equation were the carrying capacity is time dependent.
 This possibility may be considered  by the authors in a future version of this App.
-Furthermore, we remind that other similar models such as Gopertz, Richards or Bertalanffy can be used to shape the COVID-19 outbreak\cite{ma2014estimating}: the authors may consider to include them in a future version of this App.
+Furthermore it is worth nothing that other similar models such as Gopertz, Richards or Bertalanffy can be considered \cite{ma2014estimating} : the authors may consider to include them in  a future version of this App.
+
+\subsection{Entropy}
+
+In order to evaluate the effectiveness of the containment of the virus, we considered to use the Shannon Entropy:
+
+\begin{equation}
+\textbf{H}=-\sum^{n}_{i=1}\textbf{P}(x_{i})log\textbf{P}(x_{i})
+\end{equation}
+
+where, in our case, we considered $\textbf{P}(x_{i})$ as the number of currently infected people for a single cluster $i$ (the sum runs on the overall number of clusters).
+If for each cluster the ratio is the same, the containment of the virus between the clusters is failed: in this case a large value of entropy will be found; on the other hand if a strong dissimilarity between the clusters is present (e.g.
+the cases are located only in to a single one), the entropy will be low.
+The situation can be compared with the expansion of a gas: suppose that a toxic gas spreads in to building that has different rooms; the gas, due to the thermal noise, tends to expand uniformly into the whole building so that an homogeneous concentration will be obtained at least.
+This natural spread can be stopped with an external work, for instance by closing some doors or by building walls: in this case the spread will be stopped.
+It is worth nothing that this analogy, that motivated our use of the entropy index is valid as long as the virus is in its expansive phase (more precisely if the overall R(t) is $\geq $ 1): otherwise, differently with respect to the molecules of the gas that are constant over time, the active case reduces as the time increase, thus the analogy and so the index have no further meaning.
+
+\subsection{Fourier Analysis}
+
+Since we were also interested in the evaluation of the cases seasonality, we included in our App the Fourier analysis for different data.
+The idea of such analysis is to apply the following transformation:
+
+ \begin{equation}
+S(f)=\int^{+\infty}_{-\infty}s(t)e^{-2i\pi}\textit{dt}
+\end{equation}
+in order to change the basis of the function $s(t)$ from the domain of time $(t)$ to the domain of frequency $(f)$.
+For our analysis the function s(t) is the number of cases as function of time (days).
+It is worth nothing that in reality the algorithm does not evaluate the integral but instead the following expression, know as Fast Fourier Transformation:
+ \begin{equation}
+X_{q}=\sum^{N-1}_{k=0}x_{k}e^{-i\dfrac{2\pi}{N}kq}
+\end{equation}
 
@@
 The data shown are synchronised with the civil protection database.
-Whenever the App is started, a check for the updates is performed and their occurrence is indicated in the \textit{most recent updates }section.
-The home panel consists of two main sections: the choropleth map and the summary statistics.
-The choropleth map (Fig.
-\ref{Home_fig1}) is an interactive heat map with breakouts by region and by province tracking the number of Covid-19 cases.
+Whenever the App is started,  a check for the updates is performed and their occurrence is indicated in the \textit{ most recent updates }section.
+The home panel consists of two main sections: the choropleth map and the summary statistics.The choropleth map (Fig \ref{Home_fig1}) is an interactive heat map with breakouts by region and by province tracking the number of Covid-19 cases.
 
@@
 \end{equation}
-This indicator ensures that the number of cases in more populated areas are not over-indexed compared to less populated ones. Similarly, it ensures less populated areas are not under-indexed. 
-The density indicator accounts for the territorial extension (in $km^2$) of regions/provinces expressed in parts per thousand (\textperthousand ) and is calculated as follows:
+This indicator ensures that the number of cases in more populated areas are not over-indexed compared to less populated ones.
+Similarly, it ensures less populated areas are not under-indexed.
+The density  indicator accounts for the territorial extension (in $km^2$) of regions/provinces expressed in parts per thousand (  \textperthousand ) and is calculated as follows:
 \begin{equation}
-density~cases =  \dfrac{total~cases}{territorial~extension}\times 1000
+density =  \dfrac{total~cases}{territorial~extension}\times 1000
 \end{equation}
-This indicator combines the population density with infection spread, offering a control over space. Indeed, it is directly proportional to population density and to percentage cases:
-\begin{equation}
-\begin{split}
-density~cases & =   \dfrac{total~cases}{local~population} \cdot \dfrac{local~population}{territorial~extension} \times 1000 \\  & = percentage~cases \cdot density~population \times 1000
-\end{split}
-\end{equation}
-It is enhanced by densely populated territories, which, by the way, may in turn enhance spread infection too.
+This indicator ensures that we are not underestimating small regions/provinces with fewer cases than larger ones but with more concentrated cases per $Km^2$.
 Finally, the summary statistics section (Fig.
@@
 
-\subsection{Data Analysis}\label{sub:Data Analysis}
+\subsection{Data Analysis}
 

### 2020-09-17 Documentation/article_4.tex
PUB

@@
 
-In order to evaluate the effectiveness of the containment of the virus, we considered to use the Shannon Entropy:
+In order to evaluate the effectiveness of the containment of the virus, we considered to use the Gibbs formulation of the entropy:
 
@@
 
-where, in our case, we considered $\textbf{P}(x_{i})$ as the number of currently infected people for a single cluster $i$ (the sum runs on the overall number of clusters).
+where, in our case, we considered $\textbf{P}(x_{i})$ as the number of currently infected people for the single cluster $i$ (the sum runs on the overall number of clusters) and the the Boltzmann's constant was set to 1.
+Note that, in a different context, this expression is the Shannon Entropy where $\textbf{P}(x_{i})$ is the probability of the outcome $x_{i}$.
 If for each cluster the ratio is the same, the containment of the virus between the clusters is failed: in this case a large value of entropy will be found; on the other hand if a strong dissimilarity between the clusters is present (e.g.
@@
 The situation can be compared with the expansion of a gas: suppose that a toxic gas spreads in to building that has different rooms; the gas, due to the thermal noise, tends to expand uniformly into the whole building so that an homogeneous concentration will be obtained at least.
+This behaviour is due to the fact that, since the atoms are indistinguishable, the configurations in which all the atoms are distributed isotropically in to the space are much more than the configurations in which the atoms are localized \cite{kittel1998thermal} .
 This natural spread can be stopped with an external work, for instance by closing some doors or by building walls: in this case the spread will be stopped.

### 2020-09-17 Documentation/article_4.tex
pub

@@
 This natural spread can be stopped with an external work, for instance by closing some doors or by building walls: in this case the spread will be stopped.
-It is worth nothing that this analogy, that motivated our use of the entropy index is valid as long as the virus is in its expansive phase (more precisely if the overall R(t) is $\geq $ 1): otherwise, differently with respect to the molecules of the gas that are constant over time, the active case reduces as the time increase, thus the analogy and so the index have no further meaning.
+It is worth nothing that this analogy, that motivated our use of the entropy index is valid as long as the virus is in its expansive phase (more precisely if the overall R(t) is $\geq $ 1): otherwise, differently with respect to the molecules of the gas that are constant over time, the active case reduces as the time increase, thus the analogy and so the index have no further meaning since the overall entropy can spontaneously decrease.
 
